### 2021-02-17 models.Rmd
Sketch Models section

# Measurement models {#models}

Deterministic measurement error models that extends the @McLaren2019 model of metagenomics measurement to 1) describe absolute as well as relative abundance, 2) include spike-in taxa, and 3) include supplementary measurements of bulk and targeted (absolute) abundance.

<!-- TODO: go through org notes and make sure all incorporated. -->

## Actual abundances and general notation

(For now) Number of samples = J; Number of taxa = I.

$A_{ij}$ is the actual absolute abundance of taxon $i$ in sample $j$, in the target units.
Unless specified otherwise, $A$ always refers to abundance in the original sample material, not in extracted DNA.
I will assume the target units are concentration of cells per unit mass. However, other units (such as biomass per volume) may be relevant.

Let $A_{Tj}$ (or $A_{\cdot j}$??) be the total abundance in sample $j$, $A_{Tj} = \sum_i A_{ij}$.

For variables that hold matrixes or vectors, capital letters denote absolute abundances and lower case letters denote proportions.
Thus $a_{ij}$ is the matrix of actual relative abundances in terms of cells as a proportion of the total amount of cells in the sample; $a_{ij} = A_{ij} / A_{Tj}$.

Bias vectors $B^{(P)}$ are vectors of length $I$ indicating the taxon-specific bias associated with a measurement protocol $P$.
I generally assume these to be sample-independent.
An overbar denotes the average value over taxa weighted by actual abundance; 
i.e., $\bar B_j = I^{-1} \sum_i B_i a_{ij}$.

<!-- Table  -->

Measurement type        Units                              Abundance matrix                 Bias vector
----------------        -----                              ----------------                 -----------
**A**ctual abundance    [cells]                            $A$, $a$ ($I\times J$)           -
**M**etagenomics        counts                             $M$, $m$ ($I\times J$)           $B^{(M)}$
Bul**k** abundance      [cells], [DNA], or [gene copies]   $K$ ($I\times 1$ column vector)  $B^{(K)}$
**T**argeted abundance  [cells], [DNA], or [gene copies]   $T$ ($|I_T| \times J$)           $B^{(T)}$
**S**pike-in abundance  [cells], [DNA], or [gene copies]   $S$ ($|I_S| \times J$)           $B^{(S)}$

Let $I_T$ and $I_S$ be the sets of taxa with targeted abundance measurements or known spike-in abundances, respectively.
Question: Will it be more convenient for $S$ and $T$ to have the standard shape $I\times J$?

## Metagenomics measurement

$M$ is a X BY X matrix representing the (marker-gene or) metagenomics measurement; $M_{ij}$ is the sequencing count associated with taxon $i$ in sample $j$, such that $M_{ij} = 10$ indicates that 10 fragments of DNA were sequenced and assigned to taxon $i$ and sample $j$.
In the case of paired-end sequencing, a single fragment of DNA yields two reads, which may or may not be counted jointly or independently by different algorithms; here I assume that read-pairs are assigned jointly, and will use "read" and "fragment" interchangeably.
For simplicity, I ignore the possibility for index switching or other sources of cross-contamination and assume that reads assigned to sample $j$ truly arose from sample $j$.

Following the notation defined above, I let $M_{Tj} = \sum_i M_{ij}$ is the total count for sample $j$; I will sometimes refer to this number as the *sequencing depth* or *read depth* of the sample, though in reality the total count will be less than the sample's sequencing depth since often a significant fraction of reads are lost by filtering steps prior to producing the final count matrix.
I use $m_{ij} = M_{ij} / M_{Tj}$ to denote the proportion of reads assigned to taxon $i$ in sample $j$.

I assume that the composition (relative abundances) is given by the deterministic error model presented in @McLaren2019.
In this model, the bias vector $B^{(M)}$ represents the *relative* efficiencies with which various taxa are measured, and so has the same effect on measurement if we rescale it $B \to cB$ for any $c > 0$.
To remove this degree of freedom, I will suppose that $B^{(M)}_{1} = 1$, so that $B^{(M)}_i$ is the efficiency of taxon $i$ relative to the first taxon.
We may take the definition of the MWC model as being that, conditional on the total count $M_{Tj}$, the observed read count of taxon $i$ is proportional to its actual abundance times its efficiency times a taxon-independent factor $C_j$ chosen to make sample $j$'s counts sum to $M_{Tj}$,
\begin{align}
  (\#eq:M)
  M_{ij} = A_{ij} B^{(M)}_{i} C_j, && C_j = \frac{M_{Tj}}{\sum_i A_{ij} B^{(M)}_{i}}.
\end{align}
In this model, the ratios among taxa are distorted by constant factors; for two taxa $i$ and $i'$,
\begin{align}
  (\#eq:M-ratio)
  \frac{M_{ij}}{M_{i'j}} 
    = \frac{m_{ij}}{m_{i'j}} 
    = \frac{A_{ij}}{A_{i'j}} \cdot \frac{B^{(M)}_{i}}{B^{(M)}_{i'}}
\end{align}
for all samples $j$.
If one or both taxa are absent ($A_i = 0$) or undetectable ($B_i = 0$), both sides of the equation are either $0$, $\infty$, or $0/0$ (undefined).
The observed proportion of a taxon equals its actual proportion multiplied by its efficiency relative to the sample mean efficiency,
\begin{align}
  (\#eq:m)
  m_{ij} = a_{ij} \cdot \frac{B_i}{\bar B^{(M)}_j}.
\end{align}

## Spike-ins

Let the set $I_S$ denote the indexes of the spiked taxa, which I assume are also measured by the metagenomics experiment and so also correspond to rows in $M$.
The $|I_S|\times J$ matrix $S$ describes the nominal abundances with which spike-in taxa were added to each sample.
Spike-in experiments are often designed so that a taxon $i$ is added in the same concentration to all samples, which we can represent by supposing the columns of $S$ to be identical.

The bias $B^{(S)}$ should be interpreted as error in the quantification of how much spike-in was added, such that
\begin{align}
  S_{ij} = A_{ij} B^{(S)}_i.
\end{align}
I assume that this error is constant across samples.
The motivation for this error model is as follows: Suppose that we misestimated the proportion of spike-in taxon $s\in I_S$ in our original spike-in stock, by a factor $B^{(S)}_s$; now its true abundance will be off by that factor in any spike-in derived from the starting stock.
Unlike the metagenomics efficiencies, these error factors are absolute numbers.

Taxa can be added as cells (pre-extraction) or as DNA (post-extraction).
In an application where DNA spike-ins are used to make inferences about changes in absolute cell abundances, we can include the difference between the nominal abundances $S$ (the spiked DNA concentration) and the actual abundances $A$ that are expected due to factors such as DNA extraction and copy-number variation in the bias term $B^{(S)}$, so long as DNA extraction is linear (see section on saturation effects below).

## Bulk absolute abundance measurement

Consider estimating the total absolute abundance $A_{T}$ via an aggregate or broad-range measurement, such as cell counting, flow cytometry, universal 16S qPCR, or total DNA concentration (e.g. Qubit).
I let $K_j$ donote the bulk absolute abundance measurement for sample $j$ (mneumonic: "K" for the last letter in "bulk").

Bulk measurements are affected by taxon-specific bias $B^{(K)}$, where $B^{(K)}_i$ indicates how efficiently cells of taxon $i$ contribute to the bulk measurement $K$.
Unlike the metagenomics efficiencies, these bulk measurement efficiencies are absolute numbers.
For example, cells of one taxon may be more reliably counted than another; and for broad-range 16S qPCR measurement, we expect taxa to contribute in proportion to how reliably they are lysed and to their 16S copy number.
Let $B^{(K)}_i$ denote the absolute efficiency with which a cell of taxon $i$ contributes to the measurement in sample $s$, which I assume is sample-independent.
We can write the estimate as
\begin{align}
  K_j = \sum_i A_{ij} B^{(K)}_i = A_{Tj} \bar B^{(K)}_j,
\end{align}
where $bar B^{(K)}_j$ is the mean bulk-measurement efficiency in the sample.

## Targeted absolute abundance measurement

Targeted measurement of the absolute abundance of specific taxa can be made via a method such as qPCR, ddPCR, or counting CFUs selective media.
These measurements too may be subject to bias; 
I suppose that the measurement of taxon $i$ is given by
\begin{align}
  T_{ij} = A_{ij} B^{(T)}_i,
\end{align}
where $B^{(T)}_i$ is the sample-independent, taxon-specific efficiency of the targeted measurement for taxon $i$.

## Complications

These models are the simplest for these types of experiments that have taxon-specific bias.
They therefore are idealizations that may be violated by real experiments in ways that affect our conclusions.
Here I list some relevant complications that we will return to in our analysis.
For more considerations, see the Discussion of @McLaren2019.

### Abundances units and protocol stages

*Still need to find the right way to synthesize these issues.*

qPCR measurements directly estimating the concentration of a marker gene in the extracted DNA.
Efficiency factors can translate the units from DNA to cell concentration *and* account for the bias imposed by extraction and marker-gene CNV (copy-number variation).
Sometimes we may want to refer to the DNA concentrations in the sample after extraction, and in that case I will use the variable $D$, and (perhaps) use $B^{(T/D)}$ to refer to the bias of the targeted measurement vs the DNA concentrations it was applied to.
This can be useful when we are considering qPCR and also DNA spike-ins.

### Saturation in DNA extraction yields

The equation for bulk measurement reflects an idealized form of bulk measurement in which, for a fixed taxonomic composition, $K_j$ is directly proportional to $A_T$.
However, a more realistic model for qPCR may be that $K_j$ is a saturating function of $A_T$, due to factors such as enzyme consumption or saturation in the elution step creating saturation in DNA yield during extraction.
However, it is not clear what this function should depend on (input biomass, cell concentration, total DNA?) and it is not clear what bias components should be included in determining the saturation, and these are not simply $\bar B$ since that includes 16S copy-number which isn't relevant here.
Rather than try to accurately model the relationship between $K$ and $A$ in this more complicated scenario, I will use the models $K_j = f(A_{Tj})$ and/or $K_j = f(A_{Tj}\bar B^{(K)}_j)$ to get some intuition for the effect that a strong saturation effect has on our inferences when it is not accounted for.

Saturation during extraction also implications for targeted PCR measurements but not for measurements that work directly on cells; and for DNA spike-ins, but not cellular spike-ins.

### Taxonomic resolution of measurements

I suppose that the "taxa" we consider are well-defined in a special sense, such that they can be equated both across samples within a measurement type and across different measurement types.
In other words, I assume that there is a real group of organisms corresponding to "taxon $i$" with abundance $A_{i\cdot}$ that the various "taxon $i$" measurements are really measuring, such that we can, for example, directly equate the taxonomic source of the counts $M_{i\cdot}$ and the targeted measurement $T_{i\cdot}$.
Achieving this situation in practice requires careful attention to things like primer design and how the metagenomic taxonomic assignment is done, and even then it may not always be possible.

One important special case where this assumption breaks down is when measurements aggregate one or more of the *atomic taxa* (corresponding to the rows in $A$), possibly in different ways by different measurement types.
This raises at least two sorts of complications.
First, constant multplicative bias at the level of atomic taxa does not translate into constant bias at the level of aggregate taxa [@McLaren2019].
Second, measurements may target different aggregates.
For example, if our metagenomics protocol provides ASV or species-level estimates but our PCR-based targeted measurements may quantify a genus or family.
Often through sufficiently careful bioinformatics we can at least nest the metagenomics taxa within the targeted taxon, but complications remain if efficiency varies within the aggregates.

Maybe:

* Give example of effect of aggregation when efficiencies vary among the aggregated taxa; understand how in terms of the average efficiency within the aggregate.
* Write out algebraically some of the cases I'll consider later.