### 2021-07-17 article.Rmd
Prep for revision
- Edit/add notes for the revision
- Switch from rendering the vaginal example figure in article.Rmd to
  pulling the figure that is created in the notebook post
  2021-05-10-vaginal-microbiome-example

# Introduction

Most microbiome research aims to go beyond qualitative descriptions to make quantitative conclusions about associations between specific microbes or community states and key host or environmental properties.
But there is serious disagreement about whether the microbiome measurements made by marker-gene and metagenomic sequencing (MGS) are or can ever be truly quantitative. A primary reason is that taxa vary in how they respond to each step in an MGS protocol, from sample collection to bioinformatic classification.
As a result taxa can differ dramatically (e.g. 10-1000X) in how efficiently they are measured—that is, converted from cells into taxonomically classified sequencing reads—making the relative abundances we observe by MGS inaccurate representations of the actual sample compositions.
Although often associated with variation in primer binding and amplification rates and marker-gene copy-number, large variation in DNA extraction efficiency and in the ability to correctly classify reads among taxa make this taxonomic bias a feature of shotgun metagenomic measurements as much as marker-gene measurements.
Taxonomic bias is protocol specific (@mclaren2019cons) and can even vary among batches within an experiment (@yeh2018taxo); it thus not only makes MGS measurements inaccurate, but also incomparable between studies.
Taxonomic bias thereby poses a major challenge for any seeking to draw quantitative conclusions from MGS studies.

- On whether MGS can be quantitative: Is there a reference that explains the disagreement? Perhaps this question hasn't been explicitly addressed in the literature. I was gesturing towards a tension where on one hand there are many papers on the problems with community sequencing and on the other hand many papers (often by the same authors, and including us) are using it to make quantitative inferences. KL suggests Mallick et al., 2017 DOI 10.1186/s13059-017-1359-z; Or Barlow et al. 2020 https://doi.org/10.1038/s41467-020-16224-6 as possible references.

The field's current answer to this challenge is methods standardization, prioritizing consistency over accuracy.
By treating each sample within a study in exactly the same manner, we can ensure that observed differences between samples are not simply due to methodological differences.
Yet standardization does not guarantee that these observed differences reflect the actual differences in taxonomic composition among samples.
Consider what might be the simplest and most common inference made from MGS: How does the proportion of a taxon vary between samples? An observation that *Escherichia* coli makes up 10% of the sequencing reads in a fecal sample might, thanks to bias, have arisen from a sample in which the proportion of *E. coli* is actually 50% or 1%.
Yet many hope that if we observe the proportion of *E. coli* to increase from 10% to 20% between two samples, its proportion must have actually doubled—by fixing the taxonomic bias of our measurements through standardization, we have (so the argument goes) fixed the measurement error in both samples so that it will cancel when estimate the fold change.
Unfortunately, the results of @mclaren2019cons show that reality is not so neat.
Because the taxa compete for sequencing effort within a sample, the effect that taxonomic bias has on the observed proportions depends on the taxonomic composition of the sample.
As a result, MGS measurements can lead to spurious inferences about changes across samples even when taxonomic bias is constant for each taxon across samples.

- 'many hope' - consider make small tweaks to phrases like this to sound less critical

This problem posed by taxonomic bias intersects with a second, growing concern within the microbiome field: that the relative abundances measured by MGS can present a misleading view of absolute-abundance dynamics.
Without making further assumptions (which are often difficult or impossible to verify) one cannot tell whether a doubling in the proportion of a taxon reflects an increase in its abundance or a decrease in the abundances of other taxa.
This concern has motivated the development of various experimental methods to enable researchers to convert the relative abundances measured by MGS into absolute abundances (methods to be reviewed in the Appendix in a future version).
An increasingly popular approach in the study of both environmental and host-associated microbiomes involves making an independent measurement of total microbial abundance in each sample (@props2017abso,@kevorkian2018esti,@lloyd2020evid,@tettamantiboshier2020comp,@contijoch2019gutm,@vandeputte2017quan,@vieirasilva2019quan).
This estimate is then multiplied by the MGS proportions to convert them to estimates of absolute abundance.
This simple calibration procedure directly transfers the error in the MGS proportions due to taxonomic bias to the estimated absolute abundances, leading to estimates that are not only individually inaccurate but can imply inaccurate variation in abundance across samples.

- Amy points to rao2021mult as another good ref for the above

Here we summarize recent advances in our understanding of how taxonomic bias affects our ability to accurately determine how taxa vary in relative and absolute abundance across samples from different conditions—the inference problem known as *differential abundance (DA) analysis*.
We first explain how taxonomic bias can lead to spurious DA results and consider the biological conditions in which such spurious results are likely.
But we then describe several experimental and computational methods that can be used to avoid or correct these errors.
Collectively, these methods may provide practical solutions to mitigating the effects of taxonomic bias on DA analysis for a majority of microbiome studies.

## Notes for revision

- introduce differential abundance as the focus earlier
- improve introduction of absolute abundance vs. relative abundance, and methods for obtaining absolute abundances
    + might include computational methods in scope, and use the term "calibration"; and ref to the future appendix that surveys the various methods
    + consider giving general explanation and/or specific examples of why absolute abundances are important

# Bias can cause spurious results in proportion-based DA analysis

Comments on this version:

- This version skips over the effect of bias on MGS measurement, leaving that for the intro, and goes right to the effect on FCs. I think this is a mistake; I'd rather walk the reader through the logic starting with the effect on one sample.
- To accompany this, I'd like a new first figure, that illustrates the effect of competition. Essentially like Figure 2 of the eLife paper; perhaps using the vaginal mb example.

```{r vaginal-example, cache = TRUE, fig.cap = '(ref:vaginal-example)'}
svg_path <- here::here(
  "notebook/_posts/2021-05-10-vaginal-microbiome-example",
  "vaginal-microbiome-example_files/figure-html5",
  "vaginal-example-proportions-timepoints-1.svg"
)
if (knitr::is_html_output()) {
  knitr::include_graphics(svg_path)
} else if (knitr::is_latex_output()) {
  cowplot::ggdraw() + cowplot::draw_image(svg_path)
}
rm(svg_path)
```

(ref:vaginal-example) **Taxonomic bias can distort differential abundance results even when it is consistent for each taxon across samples.** Panel A shows the actual and observed proportions for hypothetical community samples from two time points, which differ in their relative abundance of three taxa. Panel B shows taxonomic bias in terms of the relative efficiencies of the three taxa against the mean efficiency of each sample; the difference between the taxon's efficiency and the sample's mean (vertical arrows) determines the fold error seen in Panel A. Panel C rearranges the plot from Panel A to show the actual and observed fold changes between time points. The efficiencies of individual taxa were estimated by @mclaren2019cons from mock community data from @brooks2015thet. The abundances are hypothetical but inspired by observations from the human vaginal microbiome; see main text.

- TODO: Add main-message sentence; see Ben's note below

We illustrate the problems posed by taxonomic bias for differential abundance using an example where the bias of an MGS protocol was explicitly measured using mock communities.
To analyze the effects of taxonomic bias on measurements of the human vaginal microbiome, @brooks2015thet performed 16S rRNA gene sequencing of mock communities of seven key vaginal species using a similar protocol to that of the Vaginal Human Microbiome Project (VaHMP).
@mclaren2019cons used this data to estimate the relative measurement efficiency of the seven species.
They found that the most efficiently measured species, *Lactobacillus iners*, had an efficiency that was 29X that of the least efficiently measured species, *Gardnerella vaginalis*, primarily due to a greater DNA extraction efficiency and 16S copy number.
A third species, *Atopobium vaginae*, had an efficiency 1.8X that of *G. vaginalis*.
A common aim of human vaginal microbiome studies, including the VaHMP, is to understand the causes and consequences of shifts in communities from being dominated by a particular *Lactobacillus* species to dominance by a non-*Lactobacillus* species such as *G. vaginalis*, which have been associated with deleterious health outcomes including an increased risk of preterm births in pregnant women.
It is therefore important to know whether the efficiency variation found by @mclaren2019cons affects the observed differential abundances between *Lactobacillus*-dominated and *Gardnerella*-dominated communities.

- Ben flagged that we want to make sure that it is clear from the text that this example is not just a one-off case of bias mattering, but is an illustration of a general/universal problem. The later discussion helps with this already; but what will help more is to start this section with a main-message sentence stating the general problem first.

Figure \@ref(fig:vaginal-example) illustrates the effects of taxonomic bias on the measured proportions for hypothetical vaginal samples from a single woman at two points in time, during which the community shifts from being dominated by *Lactobacillus* to being dominated by *Gardnerella*.
In each case, we observe a greater proportion of *Lactobacillus* than is truly present, and less of the other two species (Figure \@ref(fig:vaginal-example)A).
This positive multiplicative error^[what is the right terminology for multiplicative error that is greater than 1?] arises because *Lactobacillus* is more efficiently measured than the average species in each sample.
@mclaren2019cons showed that the observed and actual proportions of a taxon $i$ in a sample $s$ are related by the equation
\begin{align}
  (\#eq:error)
  \text{observed}_{i}(s) = \text{actual}_{i}(s) \cdot \frac{\text{efficiency}_{i}}{\text{mean efficiency}(s)}
\end{align}
where $\text{mean efficiency}(s)=\sum_{j}\text{actual}_j(s)\cdot \text{efficiency}_j$ in the denominator is the average efficiency of all taxa in that sample.
Intuitively, Equation \@ref(eq:error) says that the taxon will be over-represented by the extent to which its measurement efficiency is greater than the average efficiency of cells in the sample.
Figure \@ref(fig:vaginal-example)B shows the efficiencies of the three taxa and the mean efficiency in each sample.
The (multiplicative) difference between the taxon efficiencies and the mean efficiency determines the (multiplicative) error in each sample; for example, the mean efficiency in the second sample is 2.5, leading to a 29/2.5=11.6X error in the proportion of *Lactobacillus*.
The error in the second sample is larger—specifically, by 6X—than the first sample because the high proportion of *Gardnerella* has reduced the sample mean efficiency (and thus increased the efficiency of *Lactobacillus* relative to the mean) by 6X.

- see note on terminology for mult. error

The enrichment depends on the taxon’s efficiency relative to the sample average, rather than the efficiency itself, because the requirement that the proportions of all taxa sum to one locks the taxa in a zero-sum competition.
Notably, competition occurs (though in a less absolute form) even at the level of the raw sequencing reads: DNA extractions and PCR reactions have limited output capacity; high-concentration samples are typically subsampled to fixed, lower concentrations; and sequencing libraries are given limited sequencing effort; so that the experimental process itself typically imposes competition to be sequenced.
And it is competition that prevents the effects of taxonomic bias from canceling when we analyze how the proportions of individual taxa change across samples.

- When revising the above paragraph, pay attention to making clear that altough we are imposing competition via normalization to proportions, the experiment itself imposes competition on the absolute intermediates like DNA concentration and sequenced fragments.

Figure \@ref(fig:vaginal-example)C shows the consequences of measurement error on an analysis of the fold change in proportions between the two samples.
In reality, *Gardnerella* changes by 4X, *Lactobacillus* by 0.1X, and *Atopobium* by 0.5X.
Yet what is observed is that *Gardnerella* changes by 32X, *Lactobacillus* by 0.6X, and *Atopobium* by 3X.
In other words, the change in each case is multiplied by 6X, causing a much larger increase in *Gardnerella* and a smaller decrease *Lactobacillus* than actual, and the appearance that *Atopobium* increased when it in fact decreased.
Why do such errors occur?
The multiplicative error in Equation \@ref(eq:error) has two parts: a taxon-specific term in the numerator that is the same in each sample, and a taxon-independent term in the denominator that depends on the composition of the sample through the mean efficiency.
When we compute the fold change in a taxon i between samples s and t, 
\begin{align}
  \frac{\text{observed}_{i}(t)}{\text{observed}_{i}(s)}=\frac{\text{actual}_{i}(t)}{\text{actual}_{i}(s)}*\frac{\text{efficiency}_{i}}{\text{efficiency}_{i}}*\frac{\text{mean efficiency}(s)}{\text{mean efficiency}(t)}
\end{align}
the constant efficiency in the numerator of Equation \@ref(eq:error) cancels, but the varying mean efficiency does not, leaving an error that is equal to the inverse change in mean efficiency.
As a result, we are here left inferring that the fold-changes in proportions have all increased less (or decreased more) than they truly have.
In our example, the mean efficiency decreased by 6X (from 15 to 2.5) due to the shift in dominant taxon from the high-efficiency *Lactobacillus* to the low-efficiency *Gardnerella*, causing the observed fold changes to increase by 6X above the actual.
This increase distorted the magnitudes of the changes in *Lactobacillus* and *Gardnerella* (decreasing the first and increasing the second) and changed the sign of *Atopobium* from negative to positive.
Thus partial cancelling in the effect of a constant multiplicative bias may not be sufficient to protect against spurious results.

- TODO: Add strike-throughs to the above equation
- Revision should hone the selection of equations

Spurious results can occur for the same basic reason in proportion-based absolute-abundance inference.
Suppose that the true total abundance in each sample can be determined experimentally.
Because the absolute abundance estimate simply multiplies the erroneous observed proportions by this total, the estimated change in absolute abundance is wrong by the same factor as the fold change in proportions.
In our Figure \@ref(fig:vaginal-example) example, if the change in total abundance between samples were 2X, then we would observed both *Lactobacillus* and *Atopobium* to have increased in absolute abundance (by 1.2X and 6X), when in fact *Lactobacillus* decreased (by 5X) and *Atopobium* remained the same.

## Effect on regression of multiple samples

We showed how fold changes of individual samples may be unreliable.
But microbiome analyses often do not interpret such individual differences, and instead look for average patterns across many samples, such as how the abundance of a taxon varies with some covariate of interest.
The covariate may be discrete, such as whether the sample is from a healthy or sick person, or it may be continuous, such as a measure of space, time, or temperature.
Often these analyses can be framed as a regression problem.
For example, we might hypothesize that the log absolute abundance of taxon $i$ changes with a variable $x$ according to the simple linear regression, 
\begin{align}
  E[\text{log abundance}_{i} \mid x] = a_{0} + a_{1}x
\end{align}
where $x$ is either continuous (e.g., sediment depth) or binary (e.g., $x=1$ for treated patients and $x=0$ for controls).
Much of the effect of bias will be absorbed by the model intercept, which is typically not of interest to the researcher.
However, variation in the mean efficiency creates additional variation (reducing statistical power to detect variation in the slope) and can also create a statistical bias in our slope estimate equal to the inverse slope of $E[\text{log mean efficiency}\mid x]$ as a function of $x$.
Therefore if the mean efficiency varies across samples, but is not associated with the covariate $x$, its effect may simply be to create larger standard errors in our slope estimate.
But the larger worry may be the systematically distorted slope ($a_1$) estimates that arise in situations where the mean efficiency changes with $x$.
Notably, the error in an absolute sense is the same for all taxa—the slope is always reduced by the same amount—but the implications differ for various taxa depending on their slopes, causing magnitude or sign errors depending how the coefficient derived from the (unknown) true abundances compares to that of the mean efficiency (Figure \@ref(fig:regression-example)).

- on "(e.g., sediment depth)": Want to confirm with KL that I'm understanding this correctly. She notes: "Even though these values are sequential, we use them as discrete variables because the samples are only from one depth each. I'm not sure if this distinction matters here." To which I replied: "right....you first converted sediment depth to time, and then regressed log abundance on time to estimate growth rate? (new line) Here I was just using sediment depth as an example variable, but I could replace with time to more match your analysis"
- Consider different term than 'slope' since it's not a slope in the binary case
- Comment from AW indicated that it was hard to follow the shift from describing the hypothesized absolute abundance regression model to describing the effect of bias on the fitted model. It could help to spell out this transition and make clear we're back to discussing the situation where the absolute abundance measurements are observed (biased) relative abundance measurements multiplied by some observation of the total.
- Could be some potential for confusion in that the text notes we can get magnitude and sign errors for different taxa in the sample samples, but the figure shows a magnitude and sign error in two different sample sets on the same taxon. It could therefore be a better illustration to use two different taxa , both on the Eastern genotypes; We'll get a sign error for Penicillium but a magnitude error for any taxon that decreases more.

```{r regression-example, fig.cap = '(ref:regression-example)'}
svg_path <- here::here("figures/leopold2020host-example.svg")
if (knitr::is_html_output()) {
  knitr::include_graphics(svg_path)
} else if (knitr::is_latex_output()) {
  cowplot::ggdraw() + cowplot::draw_image(svg_path)
}
rm(svg_path)
```

(ref:regression-example) **Taxonomic bias distorts multi-sample differential abundance inference when the mean efficiency of samples is associated with the covariate of interest.** Details to synthesize into a caption: Regression of log2(Proportion) of the commensal fungus Penicillium versus timepoint; timepoints 1 and 2 are pre- and post-challenge with the pathogen Melampsora. Data is split by the region the host plants are derived from (Eastern and Western US). Calibrated proportions = Observed proportions in the real, experimental samples after adjustment for the bias measured in mock communities. Mean efficiency of each community is inferred by treating the calibrated proportions as the truth, and multiplying by efficiencies estimated from the mocks. The pathogen Melampsora has a high measurement efficiency; thus once it infects the plants, the mean efficiency of the sample increases (purple points). Efficiency is here taken as relative to the focal taxon Penicillium. West plants tend to be more resistant to the pathogen, which likely explains why the mean efficiency doesn't increase as much in the West plants. Penicillium is observed to decrease in log proportion, in both the East and West plants. But the calibrated measurements show that it actually slightly increases in the East plants, and has a lesser decrease in the West plants than what was observed before bias correction. The difference between the Calibrated and Observed data points and regression lines equals the regression line of the mean efficiency: Orange = Green + Purple; Green = Orange - Purple. The absolute error in regression coefficients is the same for all taxa. I picked Penicillium for illustration since it has the smallest observed decrease, which makes the error due to bias have a particularly significant impact.

- Note about this figure: It is currently framed as O = A - \bar B to show the effect of error on inferred DA results; it could be flipped to A = O + \bar B to show how a correction might be done.

# Implications for real-world inference

These observations show that the effects of taxonomic bias do in fact cause error in proportion-based differential abundance analyses.
Yet if the mean efficiency varies much less across samples than the abundances of individual taxa, this error will be negligible.
Alternatively, if variation in the mean efficiency is effectively random (not associated with regression covariates), the effect of bias may be to increase the residual variance and thus reduce our statistical power to assay differential abundance—unfortunate to be sure, but perhaps not as much as confidently inferring changes with the wrong sign or magnitude.
But there are salient biological scenarios where substantial associations between the mean efficiency and covariate can in fact lead to spurious differential abundance inference.

## Scenarios in which bias may cause spurious results

Figure \@ref(fig:vaginal-example) illustrates how a change in dominance by from a high- to low- efficiency taxon (or vice versa) can easily create spurious DA results.
Indeed, this scenario may be particularly common in vaginal microbiome studies.
Human microbiomes often dominated by *Lactobacillus*, but can become dominated by non-*Lactobacillus* species such as *Gardnerella* vaginalis and *Atopobium vaginae*, and these low-*Lactobacillus* states are associated with the disease bacterial vaginosis and, in pregnant women, an increased risk of preterm birth.
Using mock community measurements from @brooks2015thet, @mclaren2019cons showed that the efficiency of two common *Lactobacillus* species were 20-30X greater than *G. vaginalis* and *A. vaginae*.
Therefore we should expect samples dominated by *Lactobacillus* to have a much larger mean efficiency than those those dominated by *G. vaginalis* and *A. vaginae*, which would distort the observed associations of all taxa with any covariate that is also associated with *Lactobacillus* dominance.
The extraction and 16S sequencing protocol used by @brooks2015thet is similar to that used in the VaHMP MOMs-PI study also led by these researchers to investigate the associations of these taxa with preterm birth (@fettweis2019thev), demonstrating the potential for serious real-world consequences for these statistical/technical concerns.

- Discussion of the vaginal example in the above paragraph should be distilled in conjunction with the discussion of this example in the previous section
- Group-B strep / S. agalactiae is another "pathobiont" that has a low efficiency and I might as well included here for completeness. Can cite brooks2015thet and pace2021comp

Similar dynamics occur in a plant-fungal interactions experiment performed by @leopold2020host (Figure \@ref(fig:regression-example)).
The authors inoculated commensally-colonized trees with a fungal pathogen.
Using DNA mock communities, they showed that (excluding bias from DNA extraction) the pathogen was 10X more efficiently measured than the median commensal and 40X more efficiently measured than the lowest-efficiency commensal.
In most hosts, the pathogen increased rapidly between timepoints, driving an increase in the sample mean efficiency that in turn leads to larger decreases being observed in the proportions of commensal taxa than what is predicted using calibrated (bias-corrected) values (REF APPENDIX).
For example, the calibrated proportions indicate that the commensal *Penicillium* slightly increased in hosts from the Eastern US and slightly decreased in hosts from the Western US.
Yet the observed (uncalibrated) proportions show moderate and large decreases, respectively (Figure \@ref(fig:regression-example) above).

- In last sentence, insert the the values of the LFCs (ideally with evaluated R code)

These examples involved scenarios where individual taxa can constitute a large proportion of the community.
But in samples from highly-diverse ecosystems such as human gut, soil, and ocean sediment, one species rarely dominates.
In these cases it might seem intuitive that the mean efficiency should be relatively stable across samples: As diversity increases, the mean efficiency effectively averages over efficiencies from a greater number of taxa and so (by the central limit theorem) will converge to a constant value—if taxon efficiencies are statistically independent of taxon abundances (REF APPENDIX).
In practice, associations between efficiencies and abundances may be common and can allow large shifts in mean efficiency to occur even in highly diverse samples.

- on the example of 'ocean sediment' in the above P: I asked KL "is it reasonable to characterize the ocean sediment communities from your 2020 paper as "high diversity"? In particular, species- or OTU-level Inverse Simpson values are what's relevant here. I'm guessing it's intermediate diversity - much less than gut and soil, but much more than the vaginal microbiome", and she replied "The total microbial diversity in marine sediments worldwide is similar to that of soil. I'll email you Hoshino et al., 2020 which addresses this." However I'm not sure if the high diversity KL is noting applies to individual soil samples (i.e. what scale)
- Should make sure the above paragraph's claims are precise and accurate; see AW's feedback
- "As diversity increases" specifically refers to inverse simpson diversity / q=2. Should I explicitly state that?
- "statistically independent of taxon abundances" - I need to make sure I'm being accurate/precise here. What do we actually know? Maybe simpler to say, if taxon efficiencies are not correlated with taxon abundance

The human gut provides a clear example of how such associations might arise.
Gut microbiomes typically have high diversity at the species level but are dominated by just a small number of phyla, and two in particular: The Bacteroidetes and the Firmicutes.
The ratio of Bacteroidetes to Firmicutes can shift substantially between individuals and has been linked to a number of host traits and health conditions.
In addition, many DNA extraction protocols more efficiently lyse Gram-negative Bacteroidetes species than Gram-positive Firmicutes species (though there can also be significant variation in extraction efficiency among species within these phyla (@mclaren2019cons).
For such protocols, we should expect Bacteroidetes-dominated samples to have substantially larger mean efficiencies than Firmicutes-dominated samples, regardless of species-level diversity.
This association of efficiency with phylum abundance will distort DA inferences at any taxonomic level if the covariate is also associated with these phyla.

- mclaren2019cons measures overall extraction efficiency, not lysis efficiency; is that important for the above text?

This example illustrates a general mechanism by which association between the abundances and efficiencies of taxa are created by the common influence of evolutionary history.
Just as shared evolutionary history creates positive associations in ecological traits that drives positive correlations in how related taxa vary in abundance across samples, so does it create positive associations in bias-related traits that can lead to similar efficiencies among closely related taxa.
Such phylogenically-associated, bias-affecting traits include cell-wall toughness, ribosomal-operon copy number, binding affinity for a given set of primers, and representativeness in taxonomic databases.
Ecology and efficiency are both affected by the same evolutionary history, leading to positive associations between the two: A change in diet that increases the relative abundances of many Bacteroidetes species also increases the relative abundance of many easy-to-lyse species.

- first sentence in above P: "the abundances and efficiencies" - the idea is that both are determined by genetically encoded traits
- this is prob a tricky idea for readers and will have to take care in communicating it, and find nice refs

Associations between measurement efficiency and relative abundance can also arise because a single trait affects both.
For example, microbes at the ocean floor are slowly buried, sinking into a low nutrient, low oxygen environment.
@lloyd2020evid estimate the log fold change in the estimated absolute abundance of various taxa with sediment depth (as a proxy for time) to determine which taxa are able to persist and even grow in this difficult environment.
It is plausible that microbes with tougher cell walls would tend to persist longer (alive or dead) in the sediment, while at the same time being more difficult to extract DNA from than microbes with weaker cell walls.
As the relative abundance of tougher species increases with depth, the mean extraction efficiency decreases.
This decrease would increase the inferred log fold changes and could lead to inferred growth of taxa that are actually just persisting or even slowly dying off.
Another example of a trait that might simultaneously affect efficiency and relative abundance is ribosomal copy number, which increases the measurement efficiency in ribosomal amplicon experiments and is also linked to differences in ecology and population dynamics among species.

### Error in absolute abundance measurements

- Ben flags that this concept/section main idea should be highlighted more

So far we have ignored error associated with our absolute abundance measurements.
In fact, these measurements can also have a tendency to measure contributions from some taxa more efficiently than others.
**TODO: State how this could create problems.
Might also note the ambiguity in what the truth/target is (e.g.
live cells, or all cells; cells, or 16S rRNA gene copies; observations in Jian, Salonen, and Korpela 2021).**
Interestingly, taxonomic bias may actually make qPCR a better method for absolute quantification than cell counting for 16S rRNA gene sequencing experiments.
16S rRNA gene qPCR measures the concentration of 16S rRNA gene copies in the extracted DNA and is therefore necessarily affected by three large sources of bias in 16S rRNA gene experiments: extraction, amplification, and copy-number variation.
Yet these biases are shared by the sequencing measurement.
Though they make qPCR measurements a bad proxy for total cell density^[See note from KL below], when used for absolute differential abundance inference the shared bias in the qPCR and sequencing measurements can offset each other, leading to estimated fold changes that are relatively unaffected by taxonomic bias.
(Jian, Salonen, and Korpela 2021 suggests some of these ideas and APPENDIX gives a mathematical justification.)
If common primers are used by qPCR and 16S rRNA gene sequencing, the bias due to primer mismatches and other sources of variation in amplification could even be accounted for in this manner.^[See discussion below]
Yet ideally we would still be able to estimate and correct remaining error due to unshared bias or to mechanisms, such as saturation during DNA extraction, that could break the assumed proportionality between qPCR measurements and total abundance.

- from KL on qPCR being a bad proxy: We showed this empirically for all published measurements in marine sediments in Lloyd et al., 2013 doi:10.1128/AEM.02090-13. Apologies for the self-citation.
- In the above sentence "If common primers are used...", KL asked a question that shows I'm not being clear enough about what I'm talking about here. KL: "Wouldn't you want to use taxa specific primers for qPCR and general primers for the amplicon libraries to do this analysis? If they're the same primer sets, then it doesn't tell you anything about the relative abundance of individual taxa within the population grabbed by the primers." my reply: "The "solutions" section talks about using targeted qPCR; this section is talking about qPCR with "universal" primers for the purpose of estimating the total absolute abundance, for use in FRAxC instead of cell counting. There have been quite a few papers doing this recently and 16S sequencing companies are starting to offer it as a service so I've been thinking through the implications. This section needs to be expanded and the point I'm making about using qPCR for FRAxC is still somewhat preliminary and something we should chat about"
- should caveat something like 'Unfortunately, qPCR shares a limitation with DNA-based measurements: It assumes linearity of DNA extraction'
    - this should start own paragraph on this issue. For now, can be a 2 sentence paragraph and ref the appendix.

## We cannot simply ignore taxonomic bias

These examples show that there is potential for bias to distort DA results in both low and high diversity settings.
We therefore need methods to control or correct for bias to ensure the validity of future results, as well as methods to probe the robustness of past results.
It is possible that spurious results are unlikely in many experimental contexts.
By applying such methods across a range of biological and experimental contexts we will deepen our understanding of when bias is unlikely to significantly distort biological findings and when additional measures are needed to mitigate its effects.

# Potential solutions

## Using control samples to calibrate relative abundances

In principle, control ("mock") communities containing representative taxa from the environment of interest can be used to directly estimate the measurement efficiencies and correct bias in the MGS measurements ("calibration").
Such an approach may be sufficient for synthetic-community experiments, where all taxa are culturable, and for relatively simple natural communities like the vaginal microbiome that are dominated by a small number of culturable taxa.
But suitable mock controls may not be feasible for most complex natural ecosystems and require significant effort to develop.

An closely-related alternative to mocks are controls derivedd from natural samples, which provide a way to calibrate measurements from protocols to a reference protocol (@mclaren2019cons).
A natural fecal standard is currently being developed by [NIST](https://www.nist.gov/programs-projects/human-gut-microbiome-reference-material) and one has recently been made available commercially by Zymo Research ([ZymoBIOMICS Fecal Reference](https://www.zymoresearch.com/collections/zymobiomics-microbial-community-standards/products/zymobiomics-fecal-reference-with-trumatrix-technology)).
Ensuring the stability and homogeneity of such control samples can be challenging.
Careful testing is also needed to ensure that the preparation and storage of the controls has not significantly affected the taxonomic bias relative to the experimental samples in a given application.
As it is feasible to characterize a single standard much more extensively than a typical community sample, we may be able to obtain an estimated composition we feel comfortable treating as the ground truth. 
Yet even when this is not possible, such natural standards can allow us to reconcile results across studies despite not knowing the truth.

## Ratio-based relative and absolute abundance inference

A comprehensive approach of estimating the efficiency of all taxa is needed for getting calibrated relative abundances of all taxa within individual samples.
But what if we only want calibrated DA analysis?
That is, we are comfortable with not knowing whether *E. coli* is 10% of our sample if we can confidently determine that it doubled or halved between sample conditions?
This problem is easier to solve, and one solution is to use analysis methods based on the ratios among taxa rather than the proportions of individual taxa.

- from gdoc: this distinction of ratios vs proportions needs to be spelled out more explicitly; could even use my seminar figure (or equations) to help
- later thoughts: This distinction is so central for how I understand and explain the key ideas, should refer to it early and often

A subset of methods for analyzing differential relative abundance are derived from the field of Compositional Data Analysis (CoDA).
The defining feature of CoDA methods is that they are based on fold-changes in ratios among elements (here, the taxa) (REF AITCHISON).
The use of CoDA methods in microbiome analysis has largely been motivated by concerns over the negative correlations between taxa induced by the sum-to-one constraint in taxon proportions.
But @mclaren2019cons showed that, due to their property of perturbation invariance (REF AITCHISON), the results of CoDA differential relative abundance analysis are invariant to consistent taxonomic bias.
From Equation \@ref(eq:error), the observed ratio of a taxon $i$ to a taxon $j$ in a sample $s$ is
\begin{align}
  \frac{\text{observed}_{i}}{\text{observed}_{j}}=\frac{\text{actual}_{i}}{\text{actual}_{j}}*\frac{\text{efficiency}_{i}}{\text{efficiency}_{j}}
\end{align}
The error is independent of the sample composition and thus cancels in any function that is a fold-change of this (or any) ratio of taxa between samples.
(Perhaps insert example from Figure \@ref(fig:vaginal-example)ABC all giving same ratio FCs.)

Might ratio-based analysis also be used to overcome the effects of bias in differential absolute abundance?
In fact, some methods for determining absolute abundance are based on ratios among taxa in the MGS measurement instead of proportions.
Rather than using a measurement of total abundance, these methods require determining the abundance of one or more reference taxa.
To estimate the abundance of a focal taxon $i$, this approach multiplies the ratio of reads (or proportions) of taxon i to a reference taxon $r$ by a known or estimated abundance of the reference taxon, (while ignoring bias)
\begin{align}
  \text{estimate}_{i} = \frac{\text{observed}_{i}}{\text{observed}_{r}}*\text{abundance}_{r}
\end{align}
So far this approach has been mainly used with spike-in experiments, in which an extraneous taxon is added in a known (and typically constant) abundance to each sample so that it can serve as the reference taxon.
Yet the reference taxon could also be a naturally occurring taxon whose absolute abundance we have estimated using a method such as ddPCR directly on cells or (q/dd)PCR on the extracted DNA.
Because taxa do not compete (or competition is greatly reduced) in such targeted measurements, any taxonomic bias associated with them is expected to create constant multiplicative error across samples and so not affect fold change estimates [APPENDIX].
APPENDIX describe various theoretical and experimental considerations for both approaches.

- note from KL on choice of reference taxon: "It seems like you'd have to be pretty careful to pick a reference taxon that was not changing a lot between samples."
    - this comment indicates I need to be clearer about the three reference strategies - assume constant AA; (but may not be available so can) spike-in at constant AA; (but that might not be convenient, so can) measure AA to adjust for changes in AA.
    - my reply in the doc: "the idea is that either the reference is a taxon spiked at a fixed abundance to each sample, or it is a taxon you've measured independently with ddPCR or qPCR and so can account for its variation through the given equation." "There are also computational methods that attempt to identify taxa that don't vary (under the assumption that most taxa don't vary), which is an approach inspired from the RNAseq/gene expression literature (where it is probably a more valid assumption that most genes have consistent expression); Ben in his comments was suggesting we mention this approach too. "
- On taxa not competing in targeted measures: "I think DNA-based measurements would still be affected by competition if DNA extraction yield is a strongly saturating function of input. But if this saturation isn't very strong then the effect should remain much less than the MGS measurement, which also has competition imposed by various normalization steps "
- on 'considerations' noted in the last sentence: "perhaps most notable is how to deal with a given reference taxon not being present or detectable in every sample "
- clarify that there are 2 ways to use spike-ins: Proportion-based and ratio-based method; and only the ratio-based method is bias-invariant

## Estimating error with targeted measurements of a small number of taxa

Because targeted absolute-abundance measurements are expected to provide (relatively) accurate  estimates of fold changes, they can be used to validate and even correct the DA results derived from a proportion-based absolute DA analysis.
We illustrate the basic idea using the @lloyd2020evid experiment described above.
Absolute abundances in this study were estimated by multiplying MGS proportions by total abundance measured by cell counting; for comparison, qPCR was used to measure specific microbial taxa.^[see note]
Because the error in fold changes or a regression from the MGS measurements is taxon-independent, the difference between the two methods for these reference taxa informs us about the difference for all taxa.
@lloyd2020evid found close agreement for the specific taxa also measured by qPCR (Table 1), suggesting that variation in sample mean efficiency did not significantly distort their results.
In principle, a joint statistical analysis of all measurements would allow inferring variation in the sample mean efficiency across samples and obtaining calibrated fold change and regression estimates for all taxa [APPENDIX].

### Notes for revision

- We need an illustration of this idea. Might drive home the point that multiple measurements can be used both as qualitative checks and quantitative corrections. Could also mention the ability to use host and/or diet reads (or computational AA approaches) as a check against the experimental AA approach.
- 'specific microbial taxa' - I'm not sure if/how to address the taxonomic levels issue
- Details: Table 1 indicates that the doubling time estimated for Bathyarchaeota in Core 30 was 10.1 by FRAxC and 10.3 by qPCR, which corresponds to slopes (log abundance / time) of 1/10.1 and 1/10.3. The difference between these slopes, 1/10.3 - 1/10.1, is an estimate of the slope of the mean efficiency and hence is the correction that should be applied to all taxa’s FRAxC estimates.

> Might ratio-based analysis also be used to overcome the effects of bias in differential absolute abundance? In fact, some methods for determining absolute abundance are based on ratios among taxa in the MGS measurement instead of proportions. Rather than using a measurement of total abundance, these methods require determining the abundance of one or more reference taxa.

This needs to be revised to make clear the novelty of using non-spiked taxa as a reference; could help to replace "In fact, some methods" with clearer language.
Also make clear that some studies using reference taxa are using proportion-based methods.
Also, cite the appendix that gives an overview of AA methods, and explicitly state distinction between proportion-based and ratio-based methods.

Could help to change to:

- Ratio-based analysis can also be used to overcome the effects of bias in differential absolute abundance.
- /Make clear that the above ratio equation applies for relative and absolute observed or actual abudances (since totals cancel)/
- Suppose we knew the true absolute abundance of a reference taxon $r$; then we might estimate the absolute abundance of taxon $i$ from
  + Give the estimation equation
  + Explain the effect of taxonomic bias - get correct FCs
- It remains to determine the AA of one or more reference taxa.
  + Some existing AA methods involve using spike ins of extraneous taxa
    - discuss spike-in approach
    - often stated that bias a problem for spike-ins; however we see that it only affects the baseline and not the fold changes
      + this first part could be shunted to a discussion
    - most/all papers on spike-ins suppose the spike-in must be added in a known fixed amount.
    - however, the above argument indicates that it isn't required that the reference taxon is extraneous or has a fixed abundance; what is important is that we know the correct abundance up to a fixed multiplicative error
    - hence it can be a native taxon we've measured using a targeted method; this method can be biased as long as it has a fixed multiplicative error (unlike MGS bias.)

Need to consider if this is too detailed for the main text.
Might want to make this section of the appendix particularly clear if can't put this in the main text.

Key equations, to show side-by-side: Proportion-based and ratio-based AA inference.

## Computational approaches

What can we do for experiments that have already been conducted without mock controls, targeted control measurements, or spike-ins?
At least two purely computational approaches can still be used.

**Bias sensitivity analysis:**
A straightforward and universally available approach is to use computer simulation to determine how the results change under a range of possible sets of efficiencies, which can be randomly generated to reflect certain hypotheses about bias in the given system.
The utility of such simulations can increase the more we learn about the magnitude of bias in different systems and the taxonomic and protocol features that determine it.
Future work in developing tools and methods for simulating efficiency vectors and performing bias sensitivity analyses could be a valuable way to assay and improve the reliability of microbiome results—not just for differential absolute abundance, but for all microbiome analyses.

**Bias-aware meta-analysis:**
When the goal is to perform a meta-analysis that combines studies that have used different protocols, the unknown measurement efficiencies of each protocol can be explicitly included as parameters of the statistical model that is used.
Unknown efficiencies can be included in "compositional" linear modeling frameworks such as ALDEx2, DivNet, and fido simply by adding a protocol-specific term to the linear model of taxon log ratios.
[*Add sentence on how this can be used for non-compositional analyses.*]
Thus such bias-aware meta-analyses are already technically feasible and—if bias is truly consistent within a protocol or study—may provide a more powerful alternative to non-parametric or other meta-analysis methods that do not model bias explicitly.