### 2021-07-17 notebook/_posts/2021-07-09-illustrations/illustrations.Rmd
Develop new illustrations
This doc workshops new illustrations of spurious inference in
hypothetical examples involving vaginal taxa. This version was used to
create some figures that I shared with the Callahan lab on 2021-07-16.

```{r}
set.seed(42)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  echo = FALSE
  #> out.width = "70%",
  #> fig.align = 'center',
  #> fig.width = 6,
  #> fig.asp = 0.618,
  #> fig.show = "hold"
)

#> options(dplyr.print_min = 6, dplyr.print_max = 6)
#> options(crayon.enabled = FALSE)

library(tidyverse)
library(here)
library(cowplot)
library(patchwork)
theme_set(theme_cowplot(12))

close_elts <- function(x) x / sum(x)
```

## Example 1

Three species, with exponential growth rates of 0, <0, and >0.

```{r}
taxa <- tibble(
  taxon = c("Lactobacillus", "Gardnerella", "Atopobium", "Spike-in"),
  efficiency = c(4.68, 0.16, 0.285, 1),
  growth_rate = c(-1, 1, 0, 0) / 10
) %>%
# for now, ignore the spike-in
  filter(taxon != "Spike-in") %>%
  mutate(across(efficiency, ~. / min(.)))
```

Simulate expoential growth
```{r}
x <- taxa %>%
  expand(nesting(taxon, efficiency, growth_rate), time = seq(0, 20)) %>%
  with_groups(time, mutate,
    abundance = exp(growth_rate * time),
    total_abundance = sum(abundance),
    proportion = close_elts(abundance),
    observed_proportion = close_elts(efficiency * abundance),
    # Assuming perfect measurement of total abundance
    observed_abundance = observed_proportion * total_abundance
  )
```

```{r}
x %>%
  select(taxon:time, actual = abundance, observed = observed_abundance) %>%
  pivot_longer(c(actual, observed), names_to = "type", values_to = "abundance") %>%
  ggplot(aes(time, abundance, color = taxon)) +
  facet_wrap(~type) +
  scale_color_brewer(type = "qual", palette = 1) +
  geom_line()
```

Now let's try to show the values as stacked regions.
Strategy: Calculate ymin and ymax for each (taxon, time) value, then use `geom_ribbon()`.

We want

- ymin of first taxon = 0
- ymin of one taxon = ymax of the previous taxon
- ymax of each taxon = its ymin + its abundance
<!--  -->

We can calculate the values by just accumlating the abundances, then pull the mins and maxs.

```{r}
stack_abundance <- function(y, labels = NULL) {
  stopifnot(is.null(labels) | identical(length(y), length(labels)))
  vals <- accumulate(y, sum, .init = 0)
  ymin <- vals[seq_along(y)]
  ymax <- vals[1 + seq_along(y)]
  x <- tibble(ymin, ymax)
  if (!is.null(labels))
    x <- add_column(x, label = labels, .before = 1)
  x
}

stack_abundance(1:3)
stack_abundance(1:3, letters[1:3] %>% rev)
```

```{r}
x %>%
  select(taxon:time, actual = abundance, observed = observed_abundance) %>%
  pivot_longer(c(actual, observed), names_to = "type", values_to = "abundance") %>%
  with_groups(c(time, type), mutate,
    stack_abundance(abundance)
  ) %>%
  ggplot(aes(time, color = taxon, fill = taxon)) +
  facet_wrap(~type) +
  scale_color_brewer(type = "qual", palette = 1) +
  scale_fill_brewer(type = "qual", palette = 1) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax))
```

This actually seems to be working, though I'm not sure how reliable this method is.
E.g. this is assuming the taxa are in a consistent order.

```{r}
y <- x %>%
  select(taxon:time, actual = abundance, observed = observed_abundance) %>%
  pivot_longer(c(actual, observed), names_to = "type", values_to = "abundance") %>%
  arrange(time, type, taxon) %>%
  group_by(time, type) %>%
  nest() %>%
  mutate(stacked = map(data, ~stack_abundance(.$abundance))) %>%
  unnest(c(data, stacked))
```

```{r}
y %>%
  ggplot(aes(time, color = taxon, fill = taxon)) +
  facet_wrap(~type) +
  scale_color_brewer(type = "qual", palette = 1) +
  scale_fill_brewer(type = "qual", palette = 1) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax))
```

reverse the taxon order,

```{r}
z <- x %>%
  select(taxon:time, actual = abundance, observed = observed_abundance) %>%
  pivot_longer(c(actual, observed), names_to = "type", values_to = "abundance") %>%
  arrange(time, type, desc(taxon)) %>%
  group_by(time, type) %>%
  nest() %>%
  mutate(stacked = map(data, ~stack_abundance(.$abundance))) %>%
  unnest(c(data, stacked)) %>%
  ungroup
z %>%
  ggplot(aes(time, color = taxon, fill = taxon)) +
  facet_wrap(~type) +
  scale_color_brewer(type = "qual", palette = 1) +
  scale_fill_brewer(type = "qual", palette = 1) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax))
```

It seems like these plots don't actually do a good job of showing what is happening for the top taxon.
E.g. in this version, it looks like Atopobium is decreasing in actuality, even though it is actually fixed; and the observed increase is there but slight.

```{r}
z %>%
  filter(taxon == "Atopobium", type == "actual") %>%
  mutate(y = ymax - ymin) %>%
  count(y)
```

This suggests an intersting problem for the use of these types of plots generally.
I wonder if the problem goes away if we descretize / use stacked bars; this way we know the areas will be proportional to the abundance.

```{r}
x %>%
  select(taxon:time, actual = abundance, observed = observed_abundance) %>%
  pivot_longer(c(actual, observed), names_to = "type", values_to = "abundance") %>%
  ggplot(aes(time, abundance, color = taxon, fill = taxon)) +
  facet_wrap(~type) +
  scale_color_brewer(type = "qual", palette = 1) +
  scale_fill_brewer(type = "qual", palette = 1) +
  geom_col()
```

This visualization still suffers from the same problem - it is hard to tell that Atopobium is not really increasing.

What about the view of the proportions?

```{r}
x %>%
  select(taxon:time, actual = proportion, observed = observed_proportion) %>%
  pivot_longer(c(actual, observed), names_to = "type", values_to = "proportion") %>%
  ggplot(aes(time, proportion, color = taxon, fill = taxon)) +
  facet_wrap(~type) +
  scale_color_brewer(type = "qual", palette = 1) +
  scale_fill_brewer(type = "qual", palette = 1) +
  geom_col()
```

Perhaps it would be easier to see what's going on if we just compared two time-points, like in Figure 2 of the eLife paper.

```{r}
y <- x %>%
  filter(time %in% c(0, 10)) %>%
  mutate(timepoint = str_c("T", time)) %>%
  rename(
    "actual absolute" = abundance,
    "observed absolute" = observed_abundance,
    "actual proportion" = proportion,
    "observed proportion" = observed_proportion,
  ) %>%
  pivot_longer(c(starts_with("actual"), starts_with("observed"))) %>%
  separate(name, into = c("measurement_type", "abundance_type"), sep = " ")
```

```{r}
y %>%
  ggplot(aes(timepoint, value, color = taxon, fill = taxon)) +
  facet_grid(abundance_type~measurement_type, scales = "free_y") +
  scale_color_brewer(type = "qual", palette = 1) +
  scale_fill_brewer(type = "qual", palette = 1) +
  geom_col(width = 0.4)
```

This example is interesting; can see Lct increasing in absolute abundance despite decreasing, and Atp increasing despite saying fixed.
Probably instructive to think through why

## Example 2

Here the idea is to illustrate what happens when we increase Gard.
Let's keep Lacto and Ato fixed.

```{r}
taxa <- tibble(
  taxon = c("Lactobacillus", "Gardnerella", "Atopobium"),
  initial_density = c(2, 1, 1),
  efficiency = c(4.68, 0.16, 0.285),
  growth_rate = c(0, 1, 0) / 10
) %>%
  mutate(across(efficiency, ~. / min(.)))
x <- taxa %>%
  expand(
    nesting(taxon, initial_density, efficiency, growth_rate), 
    time = seq(0, 20)
  ) %>%
  with_groups(time, mutate,
    actual_density = initial_density * exp(growth_rate * time),
    total_density = sum(actual_density),
    actual_proportion = close_elts(actual_density),
    observed_proportion = close_elts(efficiency * actual_density),
    # Assuming perfect measurement of total abundance
    observed_density = observed_proportion * total_density
  )
xl <- x %>%
  pivot_longer(c(starts_with("actual"), starts_with("observed")), 
    values_to = "abundance") %>% 
  separate(name, into = c("type", "abundance_type"))
```

```{r}
xl %>%
  ggplot(aes(time, abundance, color = taxon)) +
  facet_grid(abundance_type~type, scales = "free_y") +
  scale_color_brewer(type = "qual", palette = 1) +
  geom_line()
```

## Example 3: Minimal AA example

Here the idea is to illustrate what happens when we increase Gard with just Lacto.

```{r}
taxa <- tibble(
  taxon = c("Lactobacillus", "Gardnerella", "Atopobium", "Spike-in"),
  initial_density = c(4, 1, 0, 0.5),
  efficiency = c(4.68, 0.16, 0.285, 1),
  growth_rate = c(0, 1, 0, 0) / 10
) %>%
  mutate(across(efficiency, ~. / min(.))) %>%
  filter(initial_density > 0)

x <- taxa %>%
  filter(taxon != "Spike-in") %>%
  expand(
    nesting(taxon, initial_density, efficiency, growth_rate), 
    time = seq(0, 20, by = 1)
  ) %>%
  with_groups(time, mutate,
    actual_density = initial_density * exp(growth_rate * time),
    total_density = sum(actual_density),
    actual_proportion = close_elts(actual_density),
    observed_proportion = close_elts(efficiency * actual_density),
    # Assuming perfect measurement of total abundance
    observed_density = observed_proportion * total_density,
    mean_efficiency = sum(actual_proportion * efficiency)
  )
xl <- x %>%
  pivot_longer(c(starts_with("actual"), starts_with("observed")), 
    values_to = "abundance") %>% 
  separate(name, into = c("type", "abundance_type"))
```

```{r}
xl %>%
  ggplot(aes(time, abundance, color = taxon)) +
  facet_grid(abundance_type~type, scales = "free_y") +
  scale_color_brewer(type = "qual", palette = 1) +
  geom_line()
```

We see Lactobacillus increasing in AA, because the growth of Gard raises the total AA while dropping the SME.

We can try plotting the total AA as well as the SME.

```{r}
# y <- x %>%
#   select(time, total_density, mean_efficiency) %>%
#   pivot_longer(
xl %>%
  filter(abundance_type == "density") %>%
  ggplot(aes(time, abundance, color = taxon)) +
  facet_wrap(~type) +
  scale_color_brewer(type = "qual", palette = 1) +
  geom_line() +
  geom_line(data = x, aes(y = total_density), color = "grey") +
  geom_line(data = x %>% mutate(type = "actual"), aes(y = mean_efficiency), color = "grey", linetype = "33")
```

Because here the mean efficiency drops faster than the total abundance grows, we see the observed abundance of Lactobacillus increase.

Does it also help to show what's happening with the relative abundances?

Here. Try showing this in the context of also showing how the QMP/AA method works, and the equations.

Logic of understanding the observed AA estimates:

Two ways:

- The actual AA estimates
- The observed proportions - Lacto dominates despite not actually dominating; hence the product with total AA shows Lacto increasing. In effect we're incorrectly assigning most of the increased cell density to Lacto
    - or: its observed prop has decreased much less than the total AA

- Compare the change in SME to the change in total abundance; use the equation for observed FC in terms of other stuff

<!--  -->

Might be better to put on log scale,
```{r}
xl %>%
  filter(abundance_type == "density") %>%
  ggplot(aes(time, abundance, color = taxon)) +
  facet_wrap(~type) +
  scale_y_log10() +
  scale_x_continuous(breaks = c(0, 10, 20)) +
  scale_color_brewer(type = "qual", palette = 1) +
  geom_line() +
  geom_line(data = x, aes(y = total_density), color = "grey") +
  geom_line(data = x %>% mutate(type = "actual"), aes(y = mean_efficiency), color = "grey", linetype = "33")
```

### Compare to using spike-in (todo)

### animated

```{r}
library(gganimate)
```

What if we show the growth of gard as an animation?
(Note, for this it might be better to have linear growth)

```{r}
xl %>%
  filter(abundance_type == "density") %>%
  ggplot(aes(type, abundance, fill = taxon)) +
  geom_col(width = 0.5) + 
  # scale_fill_brewer(type = "qual", palette = 1) +
  # Here comes the gganimate code
  transition_time(time)
# +
#   ease_aes('sine-in-out')
```

## Example 3 revisited

Taxa: Lactobacillus, Gardnerella, and a spike-in.
First show the error that arises from proportion-based inference. 
Here use just the two primary taxa to emphasize how the growth in the low-eff Gard messes up the the Lacto estimate.

```{r}
taxa <- tibble(
  taxon = c("Lactobacillus", "Gardnerella", "Atopobium", "Spike-in"),
  initial_density = c(4, 1, 0, 0.5),
  efficiency = c(4.68, 0.16, 0.285, 1),
  growth_rate = c(0, 1, 0, 0) / 10
) %>%
  mutate(across(efficiency, ~. / min(.))) %>%
  filter(initial_density > 0)
```

```{r}
x <- taxa %>%
  filter(taxon != "Spike-in") %>%
  expand(
    nesting(taxon, initial_density, efficiency, growth_rate), 
    time = seq(0, 20, by = 1)
  ) %>%
  with_groups(time, mutate,
    actual_density = initial_density * exp(growth_rate * time),
    total_density = sum(actual_density),
    actual_proportion = close_elts(actual_density),
    estimated_proportion = close_elts(efficiency * actual_density),
    # Assuming perfect measurement of total abundance
    estimated_density = estimated_proportion * total_density,
    mean_efficiency = sum(actual_proportion * efficiency)
  )
xl <- x %>%
  pivot_longer(c(starts_with("actual"), starts_with("estimated")), 
    values_to = "abundance") %>% 
  separate(name, into = c("type", "abundance_type"))
```

```{r}
clrs <- c(
  "Lactobacillus" = "#7FC97F",
  "Gardnerella" = "#BEAED4",
  "Spike-in" = "#FDC086",
  "Total" = "darkgrey",
  "Mean" = "black"
)
#
# RColorBrewer::brewer.pal(3, "Accent")
# #> [1] "#7FC97F" "#BEAED4" "#FDC086"
# RColorBrewer::display.brewer.all()
```

```{r}
# Restrict to density, and add the total as if it were another taxon.
x_dens <- xl  %>%
  filter(abundance_type == "density")
ttl <- x_dens %>%
  transmute(taxon = "Total", abundance = total_density, time, type) %>%
  distinct
x_dens <- bind_rows(x_dens, ttl) %>%
  mutate(
    label_y = case_when(
      type == "estimated" & taxon == "Total" ~ abundance * 1.2,
      type == "estimated" & taxon == "Lactobacillus" ~ abundance / 1.2,
      TRUE ~ abundance,
    )
  )

nudge_time <- 1

p_traj <- x_dens %>%
  ggplot(aes(time, abundance, color = taxon)) +
  facet_wrap(~type) +
  scale_y_log10(breaks = c(0.1, 1, 10), labels = c(0.1, 1, 10)) +
  scale_x_continuous(breaks = c(0, 10, 20)) +
  # scale_color_brewer(type = "qual", palette = 1) +
  scale_color_manual(values = clrs) +
  labs(y = "cell density") +
  theme(
    panel.spacing.x = unit(2.5, "cm"),
    legend.position = "none",
    plot.margin = unit(c(0.2, 2.3, 0.2, 0.2), units = "cm")
  ) +
  coord_cartesian(clip = "off") +
  geom_line() +
  geom_text(data = ~filter(., time == "20"), 
    aes(label = taxon, y = label_y), hjust = 0, nudge_x = nudge_time)
p_traj
```

To explain what is going on, let's consider the sample mean efficiency

```{r}
x_eff <- x %>%
  select(taxon, time, efficiency, Mean = mean_efficiency) %>%
  pivot_wider(names_from = taxon, values_from = efficiency) %>%
  pivot_longer(-time, names_to = "taxon", values_to = "efficiency")
```

```{r}
p_sme <- x_eff %>%
  ggplot(aes(time, efficiency, color = taxon)) +
  scale_y_log10() +
  # scale_y_log10(breaks = c(0.1, 1, 10), labels = c(0.1, 1, 10)) +
  scale_x_continuous(breaks = c(0, 10, 20)) +
  scale_color_manual(values = clrs) +
  theme(
    panel.spacing.x = unit(2.5, "cm"),
    legend.position = "none",
    plot.margin = unit(c(0.2, 2.3, 0.2, 0.2), units = "cm")
  ) +
  coord_cartesian(clip = "off") +
  geom_line() +
  geom_text(data = ~filter(., time == "20"), 
    aes(label = taxon), hjust = 0, nudge_x = nudge_time)
p_sme
```

Next, want to align the mean eff plot with the time axis on the actual plot.
To acheive, I might need to use distinct plots instead of facets in the first row.
Then, think of how to use the remaining quadrant - either insert the equation, or show the proportions.

Proportions at t=0 and t=20
```{r}
x_prop <- xl %>%
  filter(abundance_type == "proportion", time %in% c(0, 20)) %>%
  mutate(
    across(time, factor),
    across(taxon, as.factor),
    across(taxon, fct_rev),
  )
x_prop %>% pull(taxon) %>% levels
# x_lbl <- x_prop %>%
#   filter(type == "estimated") %>%
#   mutate(abundance = ifelse(taxon == "Lactobacillus", 0, 1))
p_prop <- x_prop %>%
  ggplot(aes(type, abundance, fill = taxon)) +
  facet_wrap(~time, labeller = label_both) +
  scale_color_manual(values = clrs[1:2]) +
  scale_fill_manual(values = clrs[1:2]) +
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  labs(y = "proportion") +
  theme(
    panel.spacing.x = unit(1, "cm"),
    legend.position = "bottom",
    plot.margin = unit(c(0.2, 0.2, 0.2, 0.2), units = "cm"),
    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
    axis.title.x = element_blank()
    # legend.title = element_text(6)
  ) +
  coord_cartesian(clip = "off") +
  geom_col(width = 0.6)
  # geom_text(data = x_lbl, 
  #   aes(color = taxon, label = taxon), hjust = 0, nudge_x = 0.4)
p_prop

```

```{r}
p_traj / (p_sme + p_prop) +
  plot_annotation(tag_levels = "A")
```

- save whole fig, and separate panels; add into lab meeting.

- practice how to explain it

<!--  -->

```{r}
pth <- "~/research/presentations/lab-meetings/2021-07-16-differential-abundance-manuscript/figures"
ggsave(fs::path(pth, "aa-example-traj.svg"), p_traj, units = "in", width = 6, height = 3.2)
ggsave(fs::path(pth, "aa-example-sme.svg"), p_sme, units = "in", width = 3.2, height = 3.0)
ggsave(fs::path(pth, "aa-example-prop.svg"), p_prop, units = "in", width = 3.6, height = 3.2)
```