### 2021-07-17 notebook/_posts/2021-05-10-vaginal-microbiome-example/vaginal-microbiome-example.Rmd
Update and render first vaginal example
Cleaned up the doc and updated to use the cleaner code and include the
caption from /article.Rmd. Rendered with distill_article class.
Multipanel figure appears in the Distill output in
vaginal-example-proportions-timepoints-1.svg so can be easily included
in the manuscript or other sources.

---
title: 'Hypothetical example of spurious fold changes in proportions in the vaginal microbiome'
author: Michael R. McLaren
date: '2021-05-10'
description: >
    This doc develops a hypothetical example in which bias leads to a spurious
    differential abundance result. It is based on a real case of bias measured
    among three species of vaginally-associated bacteria.
output:
  distill::distill_article:
    self_contained: false
    toc: true
    dev: svg
---

**Changelog**

- 2021-07-17: Cleaned up the doc and updated to use the cleaner code and
  include the caption from /article.Rmd

# Setup

```{r}
set.seed(42)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  echo = TRUE
)

library(tidyverse)
library(here)
library(cowplot)
library(patchwork)
theme_set(theme_cowplot(12))

close_elts <- function(x) x / sum(x)
```

# Simulate data

```{r}
bias <- tibble(taxon = c("Lactobacillus", "Gardnerella", "Atopobium", "Spike-in"),
  efficiency = c(4.68, 0.16, 0.285, 1)) %>%
  mutate(across(efficiency, ~. / min(.)))
# efficiency associated with the targeted measurement of Lactobacillus
targeted_efficiency <- 2
a0 <- tribble(
  ~taxon, ~timepoint, ~abundance,
  "Lactobacillus", "T1", 5,
  "Lactobacillus", "T2", 0.5,
  "Gardnerella", "T1", 2,
  "Gardnerella", "T2", 8,
  "Atopobium", "T1", 3,
  "Atopobium", "T2", 1.5,
) %>%
  # Shrink abundance to have similar scale as proportions
  mutate(across(abundance, ~ . / max(.))) %>%
  left_join(bias, by = "taxon") %>%
  with_groups(timepoint, mutate,
    total_abundance = sum(abundance),
    proportion = close_elts(abundance),
    biased_abundance = efficiency * abundance,
    biased_proportion = close_elts(biased_abundance),
    biased_count = biased_proportion * 1e3,
    abundance_estimate_bulk = biased_proportion * total_abundance,
    #> abundance_estimate_spikein = biased_count / biased_count[taxon == "Spike-in"],
    abundance_estimate_targeted = biased_count * targeted_efficiency *
      abundance[taxon == "Lactobacillus"] / biased_count[taxon == "Lactobacillus"],
  )
a1 <- a0 %>%
  select(-total_abundance) %>%
  pivot_longer(-c(taxon, efficiency, timepoint), names_to = "type")
# data frame for plots
ptb <- a1 %>%
  filter(
    type %in% c("proportion", "biased_proportion"),
  ) %>%
  mutate(
    across(type, fct_relevel, "proportion", "biased_proportion"),
    across(type, fct_recode,
      "Actual" = "proportion",
      "Observed" = "biased_proportion"),
  )
```

# Plots

Panels showing the error in measurement and in differential abundance

```{r}
shared_layers <- list(
  geom_path(aes(group = taxon),
    arrow = grid::arrow(length = unit(0.15, "inches"))),
  geom_point(size = 2),
  scale_color_brewer(type = "qual", palette = 1, guide = "none"),
  labs(y = "Proportion", color = "Taxon"),
  scale_y_log10(),
  coord_cartesian(clip = "off"),
  # scale_y_log10(breaks = c(1e-2, 3e-2, 1e-1, 3e-1, 1)) +
  expand_limits(y = 1e-2),
  theme(plot.margin = unit(c(0, 1.3, 0, 0), units = "in"))
)
# How much to nudge the taxon labels and proportions
nudge.taxon <- 0.48
nudge.prop <- 0.24
# In future iterations, consider labelling the taxa in both facets
p.meas <- ptb %>%
  ggplot(aes(type, value, color = taxon)) +
  facet_wrap(~timepoint, nrow = 1, scales = "fixed",
    labeller = as_labeller(function(x) str_c("Time point ", x))
  ) +
  shared_layers +
  geom_text(data = ~filter(., type == "Actual"),
    aes(label = round(value, 2)), nudge_x = -nudge.prop) +
  geom_text(data = ~filter(., type == "Observed"),
    aes(label = round(value, 2)), nudge_x = nudge.prop) +
  geom_text(data = ~filter(., timepoint == "T2", type == "Observed"),
    aes(label = taxon), nudge_x = nudge.taxon, hjust = 0) +
  labs(
    x = "Type",
    title = "Measurement error at each time point"
  )
p.fc <- ptb %>%
  ggplot(aes(timepoint, value, color = taxon)) +
  facet_wrap(~type, nrow = 1, scales = "fixed") +
  shared_layers +
  geom_text(data = ~filter(., timepoint == "T1"),
    aes(label = round(value, 2)), nudge_x = -nudge.prop) +
  geom_text(data = ~filter(., timepoint == "T2"),
    aes(label = round(value, 2)), nudge_x = nudge.prop) +
  geom_text(data = ~filter(., timepoint == "T2", type == "Observed"),
    aes(label = taxon), nudge_x = nudge.taxon, hjust = 0) +
  labs(
    x = "Time point",
    title = "Actual and observed fold changes"
  )
```

Panel showing the efficiencies of individual taxa and the sample means

```{r}
# First, compute the mean efficiency, then join with the taxon efficiencies in
# a table for plotting.
sme <- a0 %>%
  with_groups(timepoint, summarize, mean_efficiency = sum(proportion * efficiency))
sme0 <- sme %>%
  mutate(label = str_glue("mean ({timepoint})")) %>%
  select(label, efficiency = mean_efficiency)
sme1 <- sme %>%
  mutate(taxon = "Mean", type = "Mean") %>%
  select(taxon, efficiency = mean_efficiency, timepoint, type)
bias1 <- bias %>%
  filter(taxon != "Spike-in") %>%
  expand(nesting(taxon, efficiency), timepoint = c("T1", "T2")) %>%
  mutate(type = "Taxon")
lvls = c("Atopobium", "Gardnerella", "Lactobacillus", "Mean")
etb1 <- bind_rows(bias1, sme1) %>%
  mutate(across(taxon, factor, levels = lvls))

lvls = c("Gardnerella", "Atopobium", "Lactobacillus")
bias2 <- bias %>%
  filter(taxon != "Spike-in") %>%
  expand(nesting(taxon, efficiency), timepoint = c("T1", "T2")) %>%
  mutate(
    type = "Taxon",
    across(taxon, factor, levels = lvls),
    x = as.integer(taxon)
  ) %>%
  left_join(sme, by = "timepoint")

clrs <- c(RColorBrewer::brewer.pal(n = 3, "Accent"), rep("#585858", 2))
p.eff <- bias2 %>%
  ggplot(aes(x = x, y = efficiency, color = taxon)) +
  geom_point(size = 2) +
  geom_text(data = ~filter(., timepoint == "T2"),
    aes(label = taxon), x = 3.3, hjust = 0) +
  geom_segment(aes(xend = x, yend  = efficiency, y = mean_efficiency),
    arrow = grid::arrow(length = unit(0.15, "inches"))) +
  geom_segment(data = sme,
    aes(x = 1, xend = 3, y = mean_efficiency, yend = mean_efficiency),
    color = "#585858",
    inherit.aes = FALSE) +
  geom_text(data = sme0 %>% mutate(timepoint = "T2"),
    aes(label = label, y = efficiency), x = 3.3, hjust = 0,
    color = "#585858",
    inherit.aes = FALSE) +
  scale_color_manual(values = clrs) +
  labs(y = "Relative efficiency", x = NULL, color = "Taxon",
    title = "Taxonomic bias of protocol"
  ) +
  # Set the vertical span to match the other plots
  scale_y_log10(limits = c(1, 100) / 2,
    breaks = etb1$efficiency,
    labels = signif(etb1$efficiency, 2)
  ) +
  xlim(c(0.0, 4)) +
  facet_wrap(~timepoint, nrow = 1, scales = "fixed",
    labeller = as_labeller(function(x) str_c("Time point ", x))
  ) +
  coord_cartesian(clip = "off") +
  theme(
    axis.line.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank(),
    legend.position = "none",
    plot.margin = unit(c(0, 1.0, 0, 0), units = "in")
  )
```

Plot as a multi-panel figure

```{r vaginal-example-proportions-timepoints, fig.dim = c(8, 6) * 1.5}
p.meas + p.eff + p.fc + plot_spacer() + 
  plot_layout(byrow = TRUE, ncol = 2, widths = c(1, 0.6)) +
  plot_annotation(tag_levels = "A")
```

```{r save, include = FALSE, eval = FALSE}
fmts <- c("svg", "png", "pdf")
str_c("figures/example-vaginal-microbiome.", fmts) %>%
  here %>%
  walk(ggsave, units = "in", width = 8, height = 6, scale = 1.5)
```

Caption (copied from main article version 47ce39c): 
**Taxonomic bias can distort differential abundance results even when it is consistent for each taxon across samples.** 
Panel A shows the actual and observed proportions for hypothetical community samples from two time points, which differ in their relative abundance of three taxa.
Panel B shows taxonomic bias in terms of the relative efficiencies of the three taxa against the mean efficiency of each sample; the difference between the taxon's efficiency and the sample's mean (vertical arrows) determines the fold error seen in Panel A.
Panel C rearranges the plot from Panel A to show the actual and observed fold changes between time points.
The efficiencies of individual taxa were estimated by @mclaren2019cons from mock community data from @brooks2015thet.
The abundances are hypothetical but inspired by observations from the human vaginal microbiome; see main text.

Some ideas for improvement:

- Experiment with ways to make the connections even clearer between the efficiency differences and the error. 
- Try making the arrows vertical, so that they are directly connected to the fold changes in proportion and in efficiency.
- Can also split the third plot into T1 and T2.
- Might consider changing "Time point T1" to "Sample S1", to fit the primary language of the text.
- Can mention in the text that these could be samples from different women, or the same woman at different times.
- also use a shape scale for the taxa, to give a second way besides color to distinguish them

## Session info {.appendix}

<details><summary>Click for session info</summary>

```{r, R.options = list(width = 83)}
sessioninfo::session_info()
```

</details>