### 2021-08-23 notebook/_posts/2021-08-03-simulate-regression-example/simulate-regression-example.Rmd
Simulate regression example

---
title: "Simulate regression example"
description: |
  A short description of the post.
author:
  - name: Michael R. McLaren
date: 08-03-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r}
# library(here)
library(tidyverse)

library(cowplot)
library(patchwork)
theme_set(theme_cowplot(12))

library(ggdist)
library(ggbeeswarm)

library(speedyseq)
library(metacal)

set.seed(42)
```

# Simulate actual abundances

```{r}
n_species <- 10
n_samples <- 50

species <- str_c("sp", seq(n_species))
```

## Actual abundances

Log abundances are independent and follow a simple linear model,

```
log density = a_0 + a_1 * X + epsilon
```

```{r}
n_samples <- 100
x <- c(
  rep(0, n_samples / 2),
  rep(1, n_samples / 2)
)
X <- rbind(1, x)

Sigma <- diag(1, nrow = n_species)
rownames(Sigma) <- colnames(Sigma) <- species
a_0 <- MASS::mvrnorm(
  mu = rep(0, n_species), 
  Sigma = diag(1, nrow = n_species)
)
a_1 <- MASS::mvrnorm(
  mu = rep(0, n_species), 
  Sigma = diag(1, nrow = n_species)
)
a <- cbind(a_0, a_1)
names(a_0) <- names(a_1) <- rownames(a) <- species

epsilon <- MASS::mvrnorm(
  n = n_samples,
  mu = rep(0, n_species), 
  Sigma = diag(0.5, nrow = n_species)
) %>% 
  t

y <- a %*% X + epsilon
colnames(y) <- str_glue("sa{1:n_samples}")
```

```{r}
sam <- tibble(.sample = colnames(y), x) %>%
  mutate(across(x, factor))
actual <- phyloseq(
  otu_table(y, taxa_are_rows = TRUE) %>% transform_sample_counts(exp),
  sample_data(sam)
)
```

```{r}
actual %>%
  as_tibble %>%
  ggplot(aes(y = .otu, x = .abundance, fill = x)) +
  scale_x_log10() +
  stat_dots()
```

## Biased abundances

```{r}
log_efficiency <- MASS::mvrnorm(
  mu = rep(0, n_species),
  Sigma = diag(3, nrow = n_species)
) %>%
  {. - mean(.)}
efficiency <- exp(log_efficiency)
# names(efficiency) <- species
```

```{r}
log_efficiency %>% summary
max(efficiency) / min(efficiency)
efficiency %>% qplot + scale_x_log10()
```

To create an association of log mean efficiency and the covariate, I will set set it so that the species with the largest slope coefficient also has the largest efficiency.
This scenario is inspired by the leopold2020host and brooks2015thet experiments.

```{r}
idx_slope <- which.max(a_1)
idx_eff <- which.max(efficiency)
names(efficiency) <- species
names(efficiency)[idx_eff] <- species[idx_slope]
names(efficiency)[idx_slope] <- species[idx_eff]
```

The observed (i.e. estimated) proportions and densities are given by perturbing the actual densties by the efficiencies, and normalizing to proportions or to the original (correct) total.

```{r}
observed_prop <- actual %>% perturb(efficiency)
observed_dens <- actual %>% perturb(efficiency, norm = "keep")
```

## Mean efficiencies

```{r mean-efficiency}
mean_efficiency <- actual %>%
  transform_sample_counts(close_elts) %>%
  perturb(efficiency, norm = "none") %>%
  sample_sums %>%
  enframe(".sample", "mean_efficiency") %>%
  left_join(sam, by = ".sample")
```

```{r}
mean_efficiency %>%
  ggplot(aes(y = as.factor(x), x = mean_efficiency)) +
  scale_x_log10() +
  stat_dotsinterval()
```

## Regressions and summary plots

what is a nice pipeline for doing all the regressions?

```{r}
analysis_ps <- function(ps, ...) {
  mat <- otu_table(ps) %>% orient_taxa(as = "columns") %>% log2
  lm(mat ~ 1 + x, data = sample_data(ps) %>% as_tibble) %>%
    broom::tidy()
}
analysis_sme <- function(x, ...) {
  lm(log2(mean_efficiency) ~ 1 + x, data = mean_efficiency) %>%
    broom::tidy()
}

true_coeffs <- list('(Intercept)' = a_0, x1 = a_1) %>%
  map_dfr(enframe, "response", "truth", .id = "term")

df <- bind_rows(
  analysis_ps(actual) %>% add_column(type = "Actual"),
  analysis_ps(observed_dens) %>% add_column(type = "Observed"),
  analysis_sme(mean_efficiency) %>% 
    add_column(type = "Mean efficiency", response = "Mean efficiency"),
) %>%
  mutate(
    across(type, factor, levels = c("Actual", "Observed", "Mean efficiency")),
    lower = estimate - 2 * std.error,
    upper = estimate + 2 * std.error,
  ) %>%
  # Add the true coefficients
  left_join(true_coeffs, by = c("response", "term"))
```

```{r}
p1 <- df %>% 
  filter(term == "x1") %>%
  mutate(
    across(response, fct_reorder, estimate)
  ) %>%
  ggplot(aes(y = response, x = estimate, color = type)) +
  geom_vline(xintercept = 0, color = "grey") +
  geom_pointinterval(aes(xmin = lower, xmax = upper)) +
  geom_point(data = ~filter(., type == "Actual"),
    aes(x = truth),
    color = 'black', shape = '+', size = 4)
p2 <- df %>% 
  filter(term == "x1") %>%
  mutate(
    across(response, fct_reorder, estimate)
  ) %>%
  ggplot(aes(y = type, x = estimate, fill = type)) +
  geom_vline(xintercept = 0, color = "grey") +
  stat_dots()
p1 / p2 + 
  plot_layout(heights = c(1, 0.3)) &
  colorblindr::scale_color_OkabeIto() &
  colorblindr::scale_fill_OkabeIto()
  # scale_color_brewer(type = 'qual', palette = 7) &
  # scale_fill_brewer(type = 'qual', palette = 7)
```

```{r}
ggsave('/tmp/regression-example.pdf', width = 6, height = 5, units = 'in')
ggsave('/tmp/regression-example.png', width = 6, height = 5, units = 'in')
```

here

- understand why the 'actual' estimates are generally more extreme than the true lfc
<!--  -->

## Data plots of specific examples

Plots like the first ones I made for the Leopold data.

```{r}
species_to_plot <- c('sp9', 'sp1', 'sp5', 'sp7', 'sp8')
```

```{r}
```

### attempt 1

```{r}
mean_efficiency1 <- mean_efficiency %>%
  rename(.abundance = mean_efficiency) %>%
  add_column(.otu = "Mean efficiency")
data_long <- 
  bind_rows(
    Actual = actual %>% as_tibble,
    Observed = observed_dens %>% as_tibble,
    'Mean efficiency' = mean_efficiency1,
    .id = 'type'
  )
```

```{r}
species_to_plot <- c('Mean efficiency', 'sp9', 'sp1', 'sp5', 'sp7', 'sp8')

data_long %>%
  filter(.otu %in% species_to_plot) %>%
  mutate(
    across(.otu, factor, levels = species_to_plot)
  ) %>% 
  ggplot(aes(x, log2(.abundance), color = type)) +
  facet_wrap(~.otu, scales = "free_y", ncol = 1) +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_smooth(aes(x = as.integer(x)),
    method = "lm", geom = "line", size = 0.9
  )
```

### attempt 2

```{r}
data_long <- 
  bind_rows(
    Actual = actual %>% as_tibble,
    Observed = observed_dens %>% as_tibble,
    .id = 'type'
  )
```

```{r}
species_to_plot <- c('sp9', 'sp1', 'sp5', 'sp7', 'sp8')

p3 <- data_long %>%
  filter(.otu %in% species_to_plot) %>%
  mutate(
    across(.otu, factor, levels = species_to_plot)
  ) %>% 
  ggplot(aes(x, log2(.abundance), color = type)) +
  facet_grid(.otu~type, scales = "free_y") +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_smooth(aes(x = as.integer(x)),
    method = "lm", geom = "line", size = 0.9
  ) +
  theme(legend.position = "none")

p4 <- mean_efficiency %>%
  ggplot(aes(x, log2(mean_efficiency))) +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
  stat_smooth(aes(x = as.integer(x)),
    method = "lm", geom = "line", size = 0.9
  )
```

```{r}
p1 + p3 + p2 + p4 +
  plot_layout(ncol = 2, heights = c(1, 0.3)) &
  colorblindr::scale_color_OkabeIto() &
  colorblindr::scale_fill_OkabeIto()

```

### Testing plot types

The difference plots in 'dabestr' seem like they might be ideal; specifically the Gardner-Altman Two Group Estimation Plot

```{r}
# library(dabestr)
```

But perhaps it will be better to manually create this plot using ggdist.
Could compute the bootstrapped difference estimates, then plot them with stat_halfeye in vertical orientation at x=3.

HERE.

https://mjskay.github.io/ggdist/

```{r}
```

```{r}
data_long %>%
  filter(.otu == 'sp5', type == 'Actual') %>%
  ggplot(aes(x, log2(.abundance))) +
  geom_quasirandom(alpha = 0.3, groupOnX = TRUE)

```