### 2021-10-19 index.Rmd
Add authors
Add Jacob Nearing and Amy Willis as authors

---
title: "Accurate inference of microbial differential abundance from taxonomically-biased microbiome measurements"
author:
  - Michael R. McLaren^[North Carolina State University; send correspondence to m.mclaren42@gmail.com]
  - Jacob T. Nearing^[Dalhousie University]
  - Amy D. Willis^[University of Washington]
  - Karen G. Lloyd^[University of Tennessee]
  - Benjamin J. Callahan^[North Carolina State University]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
bibliography: [main.bib, packages.bib]
biblio-style: apalike
link-citations: yes
# description:
# For pandoc latex processing
documentclass: article
geometry:
  - top=1in
  - bottom=1in
  - left=1.5in
  - right=1.5in
output:
  bookdown::gitbook:
    css: style.css
    number_sections: true
    includes:
      in_header: header.html
    config:
      sharing:
        facebook: no
        github: yes
        twitter: no
        linkedin: no
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'twitter', 'linkedin', 'weibo', 'instapaper']
  bookdown::pdf_book:
    includes:
      in_header: latex/preamble.tex
    keep_tex: true
    dev: "pdf"
    papersize: letter
    toc_depth: 2
    toc_unnumbered: yes
    toc_appendix: yes
---

# Preface {-}

```{r}
repo <- git2r::repository(".")
sha <- repo %>% git2r::commits(n = 1) %>% .[[1]] %>% git2r::sha()
status <- repo %>% git2r::status()
```

```{r}
if (length(status$unstaged) > 0)
  print(status$unstaged)
```

_This manuscript was rendered from commit `r sha`._

::: {#preface-warning .rmdnote}

**This in-progress manuscript is not intended for general scientific use.**
It is incomplete, has not been carefully reviewed, and may contain mistakes or other inaccuracies.
Please post comments or questions on the [GitHub Issues page](https://github.com/mikemc/differential-abundance-theory/issues) or [email Mike](m.mclaren42@gmail.com).

:::

This manuscript addresses the effect that the taxonomic bias inherent in microbiome measurement has on microbial differential-abundance analysis.
We describe the basic problem posed by taxonomic bias for measuring changes in the abundance of particular taxa across conditions and describe new strategies for mitigating the errors it induces.
Analyses of both relative and absolute abundances are considered.
In its current form, the manuscript sits somewhere between a standard scientific article and a monograph; 
it consists of an article followed by a series of appendices which together give a comprehensive treatment of the implications of the @mclaren2019cons model of taxonomic bias for differential-abundance analysis and experimental design.
It is licensed under a [CC BY 4.0 License](https://creativecommons.org/licenses/by/4.0/).
See [the Zenodo record](https://doi.org/10.5281/zenodo.4552717) for how to cite the latest version.