### 2021-10-27 theory/index.Rmd

---
title: "Theory for taxonomic bias and differential abundance"
author:
  - Michael R. McLaren^[North Carolina State University; send correspondence to m.mclaren42@gmail.com]
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
bibliography: [../main.bib, ../packages.bib]
biblio-style: apalike
link-citations: yes
# description:
# For pandoc latex processing
documentclass: article
geometry:
  - top=1in
  - bottom=1in
  - left=1.5in
  - right=1.5in
output:
  bookdown::gitbook:
    css: ../style.css
    number_sections: true
    includes:
      in_header: ../header.html
    config:
      sharing:
        facebook: no
        github: yes
        twitter: no
        linkedin: no
        weibo: no
        instapaper: no
        vk: no
        all: ['facebook', 'twitter', 'linkedin', 'weibo', 'instapaper']
  bookdown::pdf_book:
    includes:
      in_header: ../latex/preamble.tex
    keep_tex: true
    dev: "pdf"
    papersize: letter
    toc_depth: 2
    toc_unnumbered: yes
    toc_appendix: yes
---

# Preface {-}

```{r}
repo <- git2r::repository(".")
sha <- repo %>% git2r::commits(n = 1) %>% .[[1]] %>% git2r::sha()
status <- repo %>% git2r::status()
```

```{r}
if (length(status$unstaged) > 0)
  print(status$unstaged)
```

_This manuscript was rendered from commit `r sha`._

::: {#preface-warning .rmdnote}

**This in-progress manuscript is not intended for general scientific use.**
It is incomplete, has not been carefully reviewed, and may contain mistakes or other inaccuracies.
Please post comments or questions on the [GitHub Issues page](https://github.com/mikemc/differential-abundance-theory/issues) or [email Mike](m.mclaren42@gmail.com).

:::

This Bookdown document contains theoretical calculations which supported the development of the manuscript.