### 2021-10-27 notebook/_posts/2021-10-27-illustrate-the-problem/illustrate-the-problem.Rmd
Illustrate the basic problem
Rmd notebook post illustrating the fact that the error in proportions is
not necessarily consistent, and hence the fold changes are not
necessarily accurate. This post looks at an example in the Brooks
dataset, but also briefly explores possible illustrations using the
Leopold dataset, including a look at ratios.

---
title: "Illustrate the basic problem"
description: |
author:
  - name: Michael R. McLaren
    url: {}
categories:
  - ref:brooks2015thet
  - ref:leopold2020host
date: 2021-10-27
draft: false
output:
  distill::distill_article:
    self_contained: false
    dev: svg
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  autodep = TRUE,
  cache.comments = FALSE,
  dpi = 300
)
```

# R setup

```{r libraries}
# Tools for microbiome data
library(speedyseq)
# Tools for general purpose data manipulation and plotting
library(tidyverse)
library(fs)
# ggplot helpers
library(ggbeeswarm)
library(cowplot)
library(patchwork)
library(scales)
theme_set(theme_cowplot())
# stats helpers
# library(broom)

library(metacal); packageVersion("metacal")
```

# Brooks et al

## Setup

```{r}
colors_brooks <- c(
  "Atopobium_vaginae" = "#009E73",
  "Gardnerella_vaginalis" = "#56B4E9",
  "Lactobacillus_crispatus" = "#D55E00",
  "Lactobacillus_iners" = "#505050",
  "Prevotella_bivia" = "#0072B2",
  "Sneathia_amnii" = "#CC79A7",
  "Streptococcus_agalactiae" = "#E69F00")

scale_y_custom <- scale_y_continuous(
    trans = 'log10',
    breaks = trans_breaks('log10', function(x) 10^x),
    labels = trans_format('log10', math_format(10^.x))
  )
```

Load data from the cellular mock communities of Brooks et al 2015 from metacal,

```{r}
dr <- system.file("extdata", package = "metacal")
list.files(dr)
actual <- file.path(dr, "brooks2015-actual.csv") |>
  read.csv(row.names = "Sample") |>
  otu_table(taxa_are_rows = FALSE)
observed <- file.path(dr, "brooks2015-observed.csv") |>
  read.csv(row.names = "Sample") |>
  subset(select = - Other) |>
  otu_table(taxa_are_rows = FALSE)
```

Note that the single-species samples present.

```{r}
actual %>% 
  as_tibble %>%
  filter(.abundance == 1) %>%
  count(.otu)
```

## Plots

Idea is to focus on a single species and show that the error in log proportions is inconsistent, and therefore there is error in the fold changes.
_L. crispatus_ is a species MM and BC commonly use to illustrate the inconsistent error, as since it has an intermediate efficiency, we see its error varying in sign/direction.
However, the errors in fold changes don't depend on this aspect, but only the variation in mean efficiency, and is therefore the same for all species.

Data frame for comparing measured and actual proportions:

```{r}
brooks_prop <- list(
  Actual = actual, 
  Measured = observed
) %>%
  map(transform_sample_counts, close_elts) %>%
  map_dfr(as_tibble, .id = 'type') %>%
  pivot_wider(names_from = type, values_from = .abundance)

brooks_fc <- list(
  Actual = actual, 
  Measured = observed
) %>%
  map(transform_sample_counts, close_elts) %>%
  map(pairwise_ratios, margin = 'samples', filter = FALSE) %>%
  map_dfr(as_tibble, .id = 'type') %>%
  pivot_wider(names_from = type, values_from = .abundance) %>%
  separate(.sample, c('sample1', 'sample2'), sep = ':') %>%
  # Filter to remove cases where Actual is 0 in one or both samples, or where
  # the numerator and denominator sample are equal
  filter(is.finite(Actual), Actual != 0, sample1 != sample2)
```

```{r}
set.seed(42)
species <- 'Lactobacillus_crispatus'

samples <- actual %>%
  as_tibble %>%
  filter(.otu == species, .abundance > 0) %>%
  pull(.sample)

p1 <- brooks_prop %>%
  filter(.otu == species, Actual > 0) %>%
  ggplot(aes(Actual, Measured, color = .otu)) +
  scale_color_manual(values = colors_brooks) +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(color = 'grey', size = 1) +
  geom_point(position = position_jitter(width = 0.02, height = 0.02), alpha = 0.5) +
  # geom_quasirandom() +
  coord_fixed() +
  expand_limits(x = 0.1, y = 0.1) +
  labs(x = 'Actual proportion', y = 'Measured proportion', color = 'Species') +
  theme(legend.position = 'none')
p2 <- brooks_fc %>%
  filter(.otu == species, sample1 %in% samples, sample2 %in% samples) %>%
  ggplot(aes(Actual, Measured, color = .otu)) +
  scale_color_manual(values = colors_brooks) +
  scale_x_log10() +
  # scale_y_custom +
  scale_y_log10() +
  geom_abline(color = 'grey', size = 1) +
  geom_point(position = position_jitter(width = 0.02, height = 0.02), alpha = 0.5) +
  # geom_quasirandom() +
  coord_fixed() +
  labs(x = 'Actual fold change', y = 'Measured fold change', color = 'Species') +
  theme(legend.position = 'none')
# p2
```

```{r brooks_lcrisp_prop_and_fc, fig.dim = c(7, 3.7), dev = 'svg'}
(p1 + labs(x = 'Actual', y = 'Measured', title = 'Proportion')) + 
  (p2 + labs(x = 'Actual', y = 'Measured', title = 'Fold change')) + 
  plot_annotation(tag_levels = 'A')
```

# Leopold and Busby 2020

This dataset is for DNA mocks rather than cellular mocks; however the magnitude of bias is similar.
This dataset has a couple advantages for illustration purposes:

- Each species has a wider range of actual proportions than in the Brooks dataset.
- The ratios between species varies, whereas in the Brooks dataset it is always 1:1.
<!--  -->

## Setup

```{r leopold2020host-data, include = TRUE}
## From the metacal 2.0 tutorial
# Download data from https://zenodo.org/record/3872145
data_path <- here::here("notebook/_data", "leopold2020host")
# To use a temporary directory:
# data_path <- file.path(tempdir(), "leopold2020")
if (!dir.exists(data_path)) {
  dir.create(data_path, recursive = TRUE)
  download.file(
    "https://zenodo.org/record/3872145/files/dleopold/Populus_priorityEffects-v1.2.zip",
    file.path(data_path, "Populus_priorityEffects-v1.2.zip")
  )
  unzip(
    file.path(data_path, "Populus_priorityEffects-v1.2.zip"), 
    exdir = data_path
  )
}
#> The microbiome data is stored in a phyloseq object,
ps <- file.path(data_path, 
  "dleopold-Populus_priorityEffects-8594f7c/output/compiled/phy.rds") %>%
  readRDS

sample_data(ps) <- sample_data(ps) %>%
  transform(
    Timepoint = factor(Timepoint)
  )

mock_actual <- file.path(data_path, 
  "dleopold-Populus_priorityEffects-8594f7c/data/MockCommunities.csv") %>%
  read.csv(row.names = 1) %>%
  select(-Sym4) %>%
  as("matrix") %>%
  otu_table(taxa_are_rows = FALSE) %>%
  transform_sample_counts(function(x) close_elts(1 / x))
mock_taxa <- taxa_names(mock_actual)
sam <- sample_data(ps) %>% as("data.frame") %>% as_tibble(rownames = "Sample")
tax <- tax_table(ps) %>% as("matrix") %>% as_tibble(rownames = "Taxon")
ps.mock <- ps %>% 
  subset_samples(Samp_type == "Mock") %>%
  prune_taxa(mock_taxa, .)
leopold_actual <- mock_actual
leopold_observed <- ps.mock %>% otu_table
rm(mock_actual, mock_taxa, sam, tax, ps.mock)
```

## Plots

```{r}
leopold_prop <- list(
  Actual = leopold_actual, 
  Measured = leopold_observed
) %>%
  map(transform_sample_counts, close_elts) %>%
  map_dfr(as_tibble, .id = 'type') %>%
  pivot_wider(names_from = type, values_from = .abundance)

leopold_fc <- list(
  Actual = leopold_actual, 
  Measured = leopold_observed
) %>%
  map(transform_sample_counts, close_elts) %>%
  map(pairwise_ratios, margin = 'samples', filter = FALSE) %>%
  map_dfr(as_tibble, .id = 'type') %>%
  pivot_wider(names_from = type, values_from = .abundance) %>%
  separate(.sample, c('sample1', 'sample2'), sep = ':') %>%
  # Filter to remove cases where Actual is 0 in one or both samples, or where
  # the numerator and denominator sample are equal
  filter(is.finite(Actual), Actual != 0, sample1 != sample2)
```

Note, in this case all species are (nominally) in all samples, but that Epicoccum is observed in 0 reads in one sample.

```{r}
species <- 'Cladosporium'
# species <- 'Alternaria'
# species <- 'Trichoderma'

p1 <- leopold_prop %>%
  filter(.otu == species, Actual > 0) %>%
  ggplot(aes(Actual, Measured, color = .otu)) +
  # scale_color_manual(values = colors_brooks) +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(color = 'grey', size = 1) +
  geom_point(position = position_jitter(width = 0.02, height = 0.02), alpha = 0.5) +
  # geom_quasirandom() +
  coord_fixed() +
  expand_limits(x = 0.1, y = 0.1) +
  labs(x = 'Actual proportion', y = 'Measured proportion', color = 'Species') +
  theme(legend.position = 'none')
p2 <- leopold_fc %>%
  filter(.otu == species) %>%
  ggplot(aes(Actual, Measured, color = .otu)) +
  # scale_color_manual(values = colors_brooks) +
  scale_x_log10() +
  # scale_y_custom +
  scale_y_log10() +
  geom_abline(color = 'grey', size = 1) +
  geom_point(position = position_jitter(width = 0.02, height = 0.02), alpha = 0.5) +
  # geom_quasirandom() +
  coord_fixed() +
  labs(x = 'Actual fold change', y = 'Measured fold change', color = 'Species') +
  theme(legend.position = 'none')
```

```{r, fig.dim = c(7, 4), dev = 'svg'}
p1 + p2
```

Same but for the ratios between two species:

```{r}
x <- list(
  Actual = leopold_actual, 
  Measured = leopold_observed
) %>%
  map(transform_sample_counts, close_elts) %>%
  map(pairwise_ratios, margin = 'taxa', filter = FALSE)

leopold_ratio <- x %>%
  map_dfr(as_tibble, .id = 'type') %>%
  pivot_wider(names_from = type, values_from = .abundance) %>%
  separate(.otu, c('otu1', 'otu2'), sep = ':') %>%
  filter(otu1 != otu2)

leopold_ratio_fc <- x %>%
  map(pairwise_ratios, margin = 'samples', filter = FALSE) %>%
  map_dfr(as_tibble, .id = 'type') %>%
  pivot_wider(names_from = type, values_from = .abundance) %>%
  separate(.otu, c('otu1', 'otu2'), sep = ':') %>%
  separate(.sample, c('sample1', 'sample2'), sep = ':') %>%
  # Filter to remove cases where Actual is 0 in one or both samples, or where
  # the numerator and denominator sample are equal
  filter(otu1 != otu2, sample1 != sample2)
```

```{r}
species1 <- 'Cladosporium'
species2 <- 'Fusarium'
# species2 <- 'Melampsora'
# species <- 'Trichoderma'

p3 <- leopold_ratio %>%
  filter(otu1 == species1, otu2 == species2) %>%
  ggplot(aes(Actual, Measured)) +
  # scale_color_manual(values = colors_brooks) +
  scale_x_log10() +
  scale_y_log10() +
  geom_abline(color = 'grey', size = 1) +
  geom_point(position = position_jitter(width = 0.02, height = 0.02), alpha = 0.5) +
  # geom_quasirandom() +
  coord_fixed() +
  expand_limits(x = 0.1, y = 0.1) +
  labs(x = 'Actual ratio', y = 'Measured ratio', color = 'Species') +
  theme(legend.position = 'none')
p4 <- leopold_ratio_fc %>%
  filter(otu1 == species1, otu2 == species2) %>%
  ggplot(aes(Actual, Measured)) +
  # scale_color_manual(values = colors_brooks) +
  scale_x_log10() +
  # scale_y_custom +
  scale_y_log10() +
  geom_abline(color = 'grey', size = 1) +
  geom_point(position = position_jitter(width = 0.02, height = 0.02), alpha = 0.5) +
  # geom_quasirandom() +
  coord_fixed() +
  labs(x = 'Actual fold change', y = 'Measured fold change', color = 'Species') +
  theme(legend.position = 'none')
```

```{r, fig.dim = c(7, 7)}
(p1 + p2) / (p3 + p4)
```

Note, the ratio-view isn't always _obviously_ better; e.g. when Mel is involved; perhaps because of noise?

To make things more directly comparable, it might be best to use faceting and to fix the axes to have the same span.

# Session info {.appendix}

<details><summary>Click for session info</summary>
```{r, R.options = list(width = 83)}
sessioninfo::session_info()
```
</details>