### 2022-01-06 notebook/_posts/2022-01-06-leopold2020host-original-regression-analysis/leopold2020host-original-regression-analysis.Rmd
Run leopold2020host regression analysis
Run regression analysis from leopold2020host with and without bias
correction, to show that bias correction has little impact.

---
title: "Perform Leopold and Busby (2020) regression analysis with and without bias correction"
description: |
  We perform the regression analysis of Leopold and Busby (2020) with and without bias correction, finding that there is negligible impact of bias correction on the results.
author:
  - name: Michael R. McLaren
    url: {}
date: 2022-01-06
bibliography: ../../../main.bib
categories:
  - ref:leopold2020host
  - differential abundance
  - bias sensitivity
output:
  distill::distill_article:
    self_contained: false
    toc: true
    dev: svg
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE
  # autodep = TRUE,
  # cache.comments = FALSE
)
```

This document asks whether bias correction has a significant impact on the regression analysis of @leopold2020host.

# Run analyses with and without bias correction

```{r libraries}
library(here)

set.seed(42)
```

This script follows the analysis of `code/jsdModels.R` in https://github.com/dleopold/Populus_priorityEffects.

We start by running the setup of libraries and loading the phyloseq object and bias estimate.
The code is modified from `jsdModels.R` to adjust working directories for reading files.

```{r lp-setup}
library(tidyverse)
library(magrittr)
library(phyloseq)
library(mvabund)
library(gt)

this_dir <- getwd()
setwd(here('notebook/_data/leopold2020host/dleopold-Populus_priorityEffects-8594f7c/'))

source("code/Rfunctions.R")

# load phyloseq data
(phy <- loadPhyloseq())

# load bias correction factors - estimated from mock communities
bias <- read.csv("output/tabs/bias.csv")

setwd(this_dir)
```

## Compute offsets with and without bias correction

Bias is accounted for in the original regression analysis through the use of taxon and sample-specific offsets, stored in a matrix `effort`.
These offsets also account for the variation in read depth across samples.
To perform the analysis with and without bias correction, we therefore compute two versions of the offsets, using the estimated bias vector and a dummy bias vector where the efficiencies are all set to 1.

```{r}
bias_list <- list(
  'Corrected' = bias, 
  'Uncorrected' = bias %>% mutate(Bhat = 1)
)

# 'effort' (offset) is calculated following jsdModels.R
effort_fun <- function(bias, phy) {
  (sample_sums(unbias(phy, bias)) %*% 
    t(bias$Bhat[match(taxa_names(phy), bias$Taxon)])
  ) %>% log 
}

effort_list <- bias_list %>%
  map(effort_fun, phy = phy)
```

The offset calculation is done as in `jsdModels.R`.
In section 'Quantification and statistical analysis'), @leopold2020host explain

> We accounted for 2 sources of unequal sampling effort, variable sampling depth and the species-specific sequencing biases, by including an offset term ($effort$) for each species $i$ in each sample $j$, in the form: $effort_{ij} = \log(bias_{i} \times depth_{j})$, where $bias_i$ is the sequencing bias correction factor for species $i$, estimated from our mock communities (see above, Estimating Sequencing Bias), and $depth_j$ is the total sum of all species in sample $j$, after dividing each by their species-specific sequencing bias correction factor.

The call to `sample_sums(unbias(phy, bias))` returns the $depth_{j}$ terms as described in this paragraph.

To understand why this calculation yields the correct offset, consider that under our model (and in our notation) we have
\begin{align}
  \log \text{reads}_i(a)
    &= \log \text{prop}_i (a) +
    \underbrace{\log \left[\frac{\text{efficiency}_i}{\text{efficiency}_S(a)} \cdot \text{reads}_S(a) \right]}_{\text{offset}}.
\end{align}
It can be shown with some algebra that
\begin{align}
  \frac{\text{reads}_S(a)}{\text{efficiency}_S(a)} 
  = \sum_i \frac{\text{reads}_i(a)}{\text{efficiency}_i}.
\end{align}
The left-hand side is the ratio of total reads to mean efficiency of the sample $a$, and the right-hand side is the $depth_{a}$ of sample $a$ defined by @leopold2020host.
It follows that the 'effort' terms computed by @leopold2020host are indeed the offsets implied by our model.

## Run analysis with each set of offsets

We'll wrap the analysis from `jsdModels.R` in a function to call with the corrected and uncorrected 'effort' offsets.
Toggle to see the code.

```{r, code_folding = TRUE}
analyze_jsd_models <- function(effort, phy, nBoot=4999) { 
  ##########################
  ### Fit genotype model ###
  ##########################

  # Make mvabund object 
  mvDat <- otu_table(phy) %>% data.frame  %>% mvabund

  # Fit joint-species model for genotype effect
  mv.full <- manyglm(mvDat ~ Genotype*Treatment, 
                     offset=effort, 
                     family="negative.binomial",
                     data=data.frame(sample_data(phy)))

  # Check model assumptions
  #plot(mv.full)
  #meanvar.plot(mvDat~sample_data(phy)$Treatment)
  #meanvar.plot(mvDat~sample_data(phy)$Genotype)

  # Test with anova.manyglm 
  # Using unstructured correlation matrix and wald tests.  
  # Including univariate test with adjustment for multiple testing.  
  mv.anova <- anova(mv.full, nBoot=nBoot, p.uni="adjusted", cor.type="shrink", test="wald")
  #> saveRDS(mv.anova, "output/rds/mv.genotype.rds")

  ########################
  ### Fit region model ###
  ########################

  #' # Fit joint-species model for genotype effect
  mv.region <- manyglm(mvDat ~ Region*Treatment, 
                       offset=effort, 
                       family="negative.binomial",
                       data=data.frame(sample_data(phy)))

  #' ## Check model assumptions
  #plot(mv.region)
  #meanvar.plot(mvDat~sample_data(phy)$Region)

  #' ## Test with anova.manyglm 
  #+ cache=T, results='asis'
  mv.region.anova <- anova(mv.region, nBoot=nBoot, p.uni="adjusted", cor.type="shrink", test="wald")
  #> saveRDS(mv.region.anova, "output/rds/mv.region.rds")

  mv.results <- bind_cols(
    mv.anova$table %>% 
      rownames_to_column() %>%
      filter(rowname!='(Intercept)') %>%
      mutate(rowname = gsub("Genotype","Host",rowname)),
    mv.region.anova$table %>% 
      rownames_to_column() %>%
      filter(rowname!='(Intercept)') %>%
      select(-rowname),
    # NOTE: [MRM] I set the name repair function to match the original behavior
    .name_repair = function(x) make.unique(x, sep = '')
  ) %>%
    gt(rowname_col = "rowname") %>%
      tab_spanner(
        label = "Genotype",
        columns = vars(Res.Df, Df.diff, wald, 'Pr(>wald)')
      ) %>%
      tab_spanner(
        label = "Ecotype",
        columns = vars(Res.Df1, Df.diff1, wald1, 'Pr(>wald)1')
      ) %>%
      fmt_number(c(4,8),
                 decimals = 1) %>%
      fmt(c(5,9),
          fns = function(x) {
            ifelse(x>=0.001,round(x,3),"< 0.001")
          }) %>%
      cols_label('Pr(>wald)'=md("*P*-value"),
                 wald=md("Wald-χ<sup>2<sup>"),
                 Res.Df=md("Df.resid"),
                 Df.diff="Df",
                 'Pr(>wald)1'=md("*P*-value"),
                 wald1=md("Wald-χ<sup>2<sup>"),
                 Res.Df1=md("Df.resid"),
                 Df.diff1="Df") %>%
      cols_move_to_start(3) %>%
      cols_move(7,5) %>%
      cols_align("center")
  #> gtsave(mv.results,"output/figs/jsdModels.png")

  list(
    mv.full = mv.full,
    mv.anova = mv.anova,
    mv.region = mv.region,
    mv.region.anova = mv.region.anova,
    mv.results = mv.results
  )
}
```

Finally, we run the analysis.
The call is wrapped in `xfun::cache_rds()` to cache the results; the hash is based on the inputs and the analysis function's source code.

```{r analyses, message = FALSE, warning = FALSE, results = 'hide'}
fun_src <- attr(analyze_jsd_models, 'srcref') %>% as('character')
res <- xfun::cache_rds({
  effort_list %>% map(analyze_jsd_models, phy = phy)
}, hash = list(fun_src, effort_list, phy),
  clean = FALSE
)
```

# Compare results

Next we compare the results of the analysis with and without bias correction.

```{r}
library(cowplot)
library(patchwork)
theme_set(theme_cowplot())
```

## ANOVA summary tables

First, let's compare the results summary tables.

**With bias correction**

```{r, echo = FALSE}
res$Corrected$mv.results
```

**Without bias correction**

```{r, echo = FALSE}
res$Uncorrected$mv.results
```

## Full model

**With bias correction**

```{r}
res[['Corrected']][['mv.full']]
```

**Without bias correction**

```{r}
res[['Uncorrected']][['mv.full']]
```

Let's compare the estimated coefficients for the various response variables.

```{r}
tidy_manyglm <- function(x) {
  x %>% coef %>%
    as_tibble(rownames = 'term') %>%
    pivot_longer(-term, names_to = 'response', values_to = 'estimate')
}
```

```{r, fig.dim = c(6, 6.5)}
coef_ests <- res %>%
  map('mv.full') %>%
  map_dfr(tidy_manyglm, .id = 'type')
x <- coef_ests %>%
  filter(term != '(Intercept)') %>%
  pivot_wider(names_from = type, values_from = estimate)
mse <- x %>%
  summarize(
    mse = mean((Corrected - Uncorrected)^2)
  ) %>% .[[1]] %>% signif(2)
x %>%
  ggplot(aes(Uncorrected, Corrected)) +
  coord_fixed() +
  geom_abline(color = 'darkred') +
  geom_point() +
  labs(title = 'Non-intercept coefficients, full model', 
    subtitle = str_glue('Mean squared difference: {mse}')
  ) +
  theme(plot.title.position = 'plot') +
  scale_color_brewer(type = 'qual', palette = 2)
```

## Region model

**With bias correction**

```{r}
res[['Corrected']][['mv.region']]
```

**Without bias correction**

```{r}
res[['Uncorrected']][['mv.region']]
```

Let's compare the estimated coefficients for the various response variables.

```{r, fig.dim = c(6, 6.5)}
coef_ests <- res %>%
  map('mv.region') %>%
  map_dfr(tidy_manyglm, .id = 'type')
x <- coef_ests %>%
  filter(term != '(Intercept)') %>%
  pivot_wider(names_from = type, values_from = estimate)
mse <- x %>%
  summarize(
    mse = mean((Corrected - Uncorrected)^2)
  ) %>% .[[1]] %>% signif(2)
x %>%
  ggplot(aes(Uncorrected, Corrected)) +
  coord_fixed() +
  geom_abline(color = 'darkred') +
  geom_point() +
  labs(title = 'Non-intercept coefficients, region model', 
    subtitle = str_glue('Mean squared difference: {mse}')
  ) +
  theme(plot.title.position = 'plot') +
  scale_color_brewer(type = 'qual', palette = 2)
```

Bias correction has little impact on the estimated coefficients for the non-intercept terms for either model.

# Session info {.appendix}

<details><summary>Click for session info</summary>
```{r, R.options = list(width = 83)}
sessioninfo::session_info()
```
</details>