### 2022-01-08 supplemental-figures.Rmd
Add supplemental figures for fungal case study
Add two additional supplemental figures for the fungal case study, for
the analysis of how the commensals respond to infection.

# Supplemental figures

<!-- Case studies: leopold2020host -->

```{r leopold2020host-variation, fig.cap = '(ref:cap-leopold2020host-variation)'}
fs::path(
  'notebook/_posts/2022-01-08-leopold2020host-case-study',
  'leopold2020host-case-study_files/figure-html5',
  'variation-in-proportions-and-mean-efficiency-1.svg'
  ) %>%
  include_svg
```

(ref:cap-leopold2020host-variation) **In the pre-infection samples from @leopold2020host, multiplicative variation in taxa proportions is much larger than that in the mean efficiency.** Panel A shows the distribution of the proportions of each commensal isolate (denoted by its genus) across all samples collected prior to pathogen inoculation; Panel C shows the distribution of the (estimated) sample mean efficiency across these same samples on the same scale; and Panel B shows the efficiency of each taxon estimated from DNA mock communities as point estimates and 90% bootstrap percentile confidence intervals. Efficiencies are shown relative to the most efficiently measured taxon (_Fusarium_).

```{r leopold2020host-infection-mean-efficiency-dist, fig.cap = '(ref:cap-leopold2020host-infection-mean-efficiency-dist)'}
fs::path(
  'notebook/_posts/2022-01-08-leopold2020host-case-study',
  'leopold2020host-case-study_files/figure-html5',
  'infection-mean-efficiency-dist-1.svg'
  ) %>%
  include_svg
```

(ref:cap-leopold2020host-infection-mean-efficiency-dist) **The mean efficiency tends to increase after infection due to the high proportion of the pathogen.**

```{r leopold2020host-infection-lfc, fig.cap = '(ref:cap-leopold2020host-infection-lfc)'}
fs::path(
  'notebook/_posts/2022-01-08-leopold2020host-case-study',
  'leopold2020host-case-study_files/figure-html5',
  'infection-lfc-estimates-1.svg'
  ) %>%
  include_svg
```

(ref:cap-leopold2020host-infection-lfc) **Bias correction increases the estimated increase in log proportion in response to infection for commensal taxa across all host genotypes.** Shown are the estimated log fold change (LFC) and 95% confidence intervals from simple linear regression of log (base e) proportion against experimental timepoint for commensal taxa. Negative values indicate that the proportion of the taxon decreased on average in response to infection, which we expect due to the increase in pathogen abundance and the sum-to-one constraint. Bias leads to artificially low estimates, as the pathogen drives an increase in mean efficiency. In Western genotypes, several taxa whose proportion is observed to decrease without correction are instead found to remain stable or even increase when bias is corrected.

<!-- Implications: MOMSPI -->

```{r momspi-mean-efficiency-fcs, fig.cap = '(ref:cap-momspi-mean-efficiency-fcs)'}
fs::path(
  "notebook/_posts/2021-11-01-momspi-summary/momspi-summary_files", 
  "figure-html5/momspi-mean-efficiency-fcs-1.svg"
) %>%
  include_svg
```

(ref:cap-momspi-mean-efficiency-fcs) **Fold changes in the mean efficiency within and between women in the MOMS-PI study.**

<br><br>

<!-- Solutions -->

```{r sensitivity-example, fig.cap = '(ref:cap-sensitivity-example)'}
fs::path(
  "notebook/_posts/2021-10-18-evaluate-robustness-example/",
  "evaluate-robustness-example_files/figure-html5/summary_plot-1.svg"
) %>%
  include_svg
```

(ref:cap-sensitivity-example) **A bias-sensivity analysis can be performed to examine how sensitive the results of a DA analysis are to assumptions about taxonomic bias in community measurements.** The figure shows the results of a bias-sensitivity analysis used to study the effect of bias on the association of _Gardnerella vaginalis_ and preterm birth that was investigated by @callahan2017repl. 100 random efficiency vectors were drawn at 6 different bias strengths (quantified by the variance in log efficiency, $\sigma_{e}^{2}$). Each efficiency vector was used to calibrate the MGS profiles and perform a DA association test of _G. vaginalis_ versus the host's preterm birth outcome; regression coefficients $\hat \beta$ indicate the increase of average logit proportion of  _G. vaginalis_ in women who experienced preterm birth.