### 2022-01-10 notebook/_posts/2021-12-15-momspi-regression-diversity/momspi-regression-diversity.Rmd

---
title: "MOMS-PI: Regression on diversity"
description: Analyze the impact of bias on a DA analysis of relative abundances versus diversity.
author:
  - name: Michael R. McLaren
    url: {}
categories:
  - ref:brooks2015thet
  - ref:fettweis2019thev
date: 2021-12-15
draft: false
bibliography: ../../../main.bib
output:
  distill::distill_article:
    self_contained: false
    dev: svg
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  autodep = TRUE,
  cache.comments = FALSE,
  dpi = 300
)
```

# Setup

## R setup

```{r libraries}
# Tools for microbiome data
library(speedyseq)
# Tools for general purpose data manipulation and plotting
library(tidyverse)
library(fs)
library(here)
# ggplot helpers
library(ggbeeswarm)
library(cowplot)
library(patchwork)
library(scales)
# stats helpers
# library(broom)

library(metacal); packageVersion("metacal")
```

```{r ggplot}
theme_set(theme_cowplot())

scale_y_custom <- scale_y_continuous(
    trans = 'log10',
    breaks = trans_breaks('log10', function(x) 10^x),
    labels = trans_format('log10', math_format(10^.x))
  )
```

## MOMS-PI setup

Load the MOMS-PI profiles and estimate bias,

```{r}
source(here('notebook/_code/momspi-setup.R'))
```

To simplify interpretation of some of the regression models, let's rarefy samples to 10K reads (we'll be dropping some samples that are below this limit).

```{r}
momspi_rare <- xfun::cache_rds({
  momspi %>% 
    rarefy_even_depth(
      sample.size = 1e4, 
      rngseed = 42,
      replace = FALSE,
      trimOTUs = FALSE
    )
  }, hash = momspi
)
# Make sure no taxa have 0 counts, in case this causes issues later on
stopifnot(all(momspi_rare %>% taxa_sums > 0))
```

Let's also add the new sample sums and estimated mean efficiency and

```{r}
mean_eff <- momspi_rare %>%
 calibrate(bias_all_vec, norm = 'close') %>%
 perturb(bias_all_vec, norm = 'none') %>% 
 sample_sums %>%
 enframe('.sample', 'mean_efficiency')

momspi_rare <- momspi_rare %>% 
  mutate_sample_data(., 
    sample_sum_original = sample_sum,
    sample_sum = sample_sums(.),
  ) %>%
  left_join_sample_data(mean_eff, by = '.sample')
```

Let's also add the top species in each sample,

```{r}
get_top_species <- function(ps) {
  x <- ps %>%
    transform_sample_counts(close_elts) %>%
    otu_table %>%
    as_tibble(pivot = TRUE) %>%
    with_groups(.sample, slice_max, .abundance, n = 1)
  if (anyDuplicated(x$.sample))
    stop('Ties present')
  x %>% select(.sample, top_species = .otu)
}
top_species <- list(
  'obs' = momspi_rare,
  'cal' = momspi_rare %>% calibrate(bias_all_vec)
  ) %>%
  map_dfr(get_top_species, .id = 'type')
top_species_wide <- top_species %>%
  pivot_wider(values_from = top_species, names_from = type, 
    names_prefix = 'top_species_') 
momspi_rare <- momspi_rare %>%
  left_join_sample_data(top_species_wide, by = '.sample')
```

```{r, include = FALSE, eval = FALSE}
# check that see expected difference between obs and cal
# top_species_wide  %>% count(top_species_obs, top_species_cal) %>% arrange(-n) %>% print(n=30)
top_species_wide  %>% count(top_species_obs) %>% arrange(-n) %>% print(n=10)
top_species_wide  %>% count(top_species_cal) %>% arrange(-n) %>% print(n=10)
```

# Regression analysis

NOTE: Consider saving output so can be used in other notebook files

## Classify samples into high and low diversity groups

Compute Shannon and Inverse-Simpson diversity indices and add to the sample metadata.
`Phyloseq::estimate_richness()` will complain about the lack of singletons, but we can ignore this since we aren't using species richness estimators.
I will exponentiate the Shannon index to convert it to the q1 Hill number so that (like Inverse Simpson) it has the scale of number of species.

```{r}
div <- list(
  'obs' = momspi_rare,
  'cal' = momspi_rare %>% calibrate(bias_all_vec)
) %>%
  map(estimate_richness, measures = c('Shannon', 'InvSimpson')) %>%
  map_dfr(as_tibble, rownames = '.sample', .id = 'type') %>%
  transmute(
    .sample, type,
    div_q1 = exp(Shannon),
    div_q2 = InvSimpson,
  ) %>%
  pivot_longer(starts_with('div_'), names_to = 'index', values_to = 'diversity')
div_wide <- div %>%
  pivot_wider(values_from = diversity, names_from = c('index', 'type')) 
```

```{r, fig.dim = c(7, 7)}
p <- div_wide %>%
  mutate(across(starts_with('div'), log2)) %>%
  ggplot() +
  geom_point(aes(x = .panel_x, y = .panel_y), alpha = 0.5) +
  ggforce::geom_autodensity() +
  geom_bin2d(aes(x = .panel_x, y = .panel_y)) +
  ggforce::facet_matrix(vars(starts_with('div')), layer.diag = 2, layer.lower = 3)
ggrastr::rasterize(p, dpi = 300)
```

The observed (Log) order-1 and order-2 diversity show a tight correlation with each other, as do the calibrated diversities, while the correlation is much lower when comparing observed and calibrated numbers.
The reason is likely that calibration will tend to increase the diversity of samples that are observed to be dominated by a high-efficiency taxon (like L. iners) and decrease the diversity of samples dominated by a low efficiency tason (like G. vaginalis).

Shannon diversity of the observed profiles is what is most often considered in vaginal microbiome studies, so let's use this metric to partition the samples.

```{r}
div_wide %>%
  ggplot(aes(x = div_q1_obs)) +
  scale_x_log10() +
  annotate('rect', xmin = 0, xmax = 1.5, ymin = -Inf, ymax = Inf, 
    fill = 'blue', alpha = 0.3) +
  annotate('rect', xmin = 1.5, xmax = 3, ymin = -Inf, ymax = Inf, 
    fill = 'grey', alpha = 0.3) +
  annotate('rect', xmin = 3, xmax = Inf, ymin = -Inf, ymax = Inf, 
    fill = 'green', alpha = 0.3) +
  geom_histogram() +
  geom_vline(xintercept = c(1.5, 3), color = 'black')
```

I'll pick the first mode as the 'Low diversity' group, treat the second mode as 'Medium', and the tail as 'High'.

```{r}
div_wide <- div_wide %>%
  mutate(
    div_group = case_when(
      div_q1_obs < 1.5 ~ 'Low',
      div_q1_obs > 3 ~ 'High',
      TRUE ~ 'Medium'
    ) %>% factor(levels = c('Low', 'Medium', 'High'))
  )
```

## Regression setup

Add the diversity grouping information, subset to the Low and High samples, and compute the offsets for the Gamma-Poisson model

```{r}
momspi_div <- momspi_rare %>%
  left_join_sample_data(div_wide, by = '.sample') %>%
  filter_sample_data(div_group %in% c('Low', 'High')) %>%
  mutate_sample_data(
    gp_offset_obs = log(sample_sum),
    gp_offset_cal = log(sample_sum) - log(mean_efficiency),
  )
```

We'll use the criterion from Fettweis et al to pick the taxa to test.
Note, we may not get the same exact taxa as them since we're using a larger set of women and ignoring the medium-diversity samples.
Note, the log proportions used in the regression analysis below are with respect to the full species set.

```{r}
taxa_to_test <- momspi_div %>%
  transform_sample_counts(close_elts) %>%
  filter_taxa2(~ mean(. >= 1e-2) >= 0.05 | mean(. >= 1e-3) >= 0.15) %>%
  taxa_names %>%
  print
```

## Linear regression on log proportion

The simple linear regression is the fastest and simplest approach, and the easiest to understand using results from the manuscript.

First, we need to ensure there are no zeros.
I'll try two approaches; the first is to use a pseudocount of 0.5.

```{r}
stat_test <- function(ps, taxa) {
  ps %>%
    filter_sample_data(div_group %in% c('Low', 'High')) %>%
    transform_sample_counts(close_elts) %>%
    prune_taxa(taxa, .) %>%
    as_tibble %>%
    with_groups(.otu, nest) %>%
    mutate(
      fit = map(data,
        ~lm(log(.abundance) ~ div_group, data = .x)
      ),
    ) %>%
    select(-data)
}
res_div_lm <- list(
  'Observed' = momspi_div %>% transform_sample_counts(~. + 0.5),
  'Calibrated' = momspi_div %>% transform_sample_counts(~. + 0.5) %>%
    calibrate(bias_all_vec)
) %>%
  map_dfr(stat_test, taxa = taxa_to_test, .id = 'type')
res_div_lm_tidy <- res_div_lm %>%
  mutate(across(fit, map, broom.mixed::tidy)) %>%
  unnest(fit) %>% 
  filter(term != '(Intercept)')
```

```{r}
# x should be the outout of a tidier
plot_estimates <- function(x) {
  err <- x %>%
    select(name = type, .otu, value = estimate) %>%
    pivot_wider() %>%
    mutate(
      error = Observed - Calibrated,
      error_sign = sign(Observed) == sign(Calibrated),
      sign = case_when(
        sign(Observed) > 0 & sign(Calibrated) > 0 ~ 'Positive',
        sign(Observed) < 0 & sign(Calibrated) < 0 ~ 'Negative',
        TRUE ~ 'Depends',
      )
    ) %>%
    select(.otu, error_sign, sign)

  p1 <- x %>%
    left_join(err, by = '.otu') %>%
    ggplot(aes(y = type, x = estimate, fill = sign)) +
    scale_fill_brewer(type = 'qual', palette = 2,
      guide = guide_legend(reverse = TRUE)) +
    geom_vline(xintercept = 0, color = 'grey') +
    ggdist::geom_dots() +
    labs(fill = 'Sign of estimate', x = 'Estimated LFC', y = 'Meas. type')
  p2 <- x %>%
    mutate(
      .otu_fct = fct_reorder(.otu, estimate, .fun = min) %>%
        fct_relabel(function(x) str_replace_all(x, '_', ' '))
    ) %>%
    ggplot(aes(y = .otu_fct, x = estimate, color = type)) +
    # scale_color_brewer(type = 'qual', palette = 7) +
    colorblindr::scale_color_OkabeIto() +
    geom_vline(xintercept = 0, color = 'grey') +
    geom_linerange(size = 1,
      aes(
        xmin = estimate - 2*std.error,
        xmax = estimate + 2*std.error
      )) +
    geom_point(size = 1.5) +
    labs(y = 'Species-level OTU', x = 'Estimated LFC', color = 'Meas. type')
  p1 / p2 +
    plot_layout(ncol = 1, heights = c(0.3, 1)) +
    plot_annotation(tag_levels = 'A') &
    theme_minimal_hgrid()
}
```

```{r regression-lm, fig.dim = c(7, 7) * 1.1}
plot_estimates(res_div_lm_tidy)
```

The data is unlikely to meet the assumptions generally considered necessary for simple linear regression.
In particular, the standard errors (and confidence intervals) are likely to be much too small.
Bootstrapped intervals could be used to address this issue, if we are interested in understanding the impact of bias on confidence intervals in a more realistic scenario.

```{r, include = FALSE, eval = FALSE}
# A simple approach to improving the validity of the results could be to
# generate bootstrap CIs.
lm_boot <- function(.data, ...) {
  fit <- lm(log(.abundance) ~ div_group, data = .data)
  car::Boot(fit, ...)
}
x <- momspi_div %>% transform_sample_counts(~. + 0.3) %>% 
  filter_sample_data(div_group %in% c('Low', 'High')) %>%
  filter_tax_table(.otu == 'Gardnerella_vaginalis') %>% 
  as_tibble
fit <- lm(log(.abundance) ~ div_group, data = x)
bt <- car::Boot(fit)
bt %>% broom::tidy()
broom::tidy(bt, conf.int = TRUE)
confint(bt, type = 'perc') %>% broom::tidy()
```

### Alternate zero-replacement strategy

Let's also try adjust the proportions to the Dirichlet posterior mean, with the prior set to the average proportions across all samples.
This amounts to a pseudocount of `ntaxa` times the mean (observed) proportion of the species, and so adjusts for differences in mean proportion among species.

```{r}
momspi_div_alt <- momspi_div %>% 
  adjust_dirchlet
```

```{r regression-lm-alt, fig.dim = c(7, 7) * 1.1}
res_div_alt_lm <- list(
  'Observed' = momspi_div_alt,
  'Calibrated' = momspi_div_alt %>% calibrate(bias_all_vec)
) %>%
  map_dfr(stat_test, taxa = taxa_to_test, .id = 'type')
res_div_alt_lm_tidy <- res_div_alt_lm %>%
  mutate(across(fit, map, broom.mixed::tidy)) %>%
  unnest(fit) %>% 
  filter(term != '(Intercept)')

plot_estimates(res_div_alt_lm_tidy)
```

## Interpretion

Check the diversity-efficiency relationship

```{r}
sam <- momspi_div %>% sample_data %>% as_tibble
sam %>%
  ggplot(aes(y = div_group, x = mean_efficiency)) +
  scale_x_log10() +
  ggdist::stat_slab()
```

From this plot, we can see that the mean efficiency is 3-10X lower in the High diversity relative to the low diversity group.

The results of a linear regression analysis give us the systematic error in LFCs for the simple linear regression,

```{r}
mean_eff_lm <- lm(log(mean_efficiency) ~ div_group,
  data = sam
)
mean_eff_lm %>% summary
mean_eff_lfc <- mean_eff_lm %>% broom::tidy() %>% 
  filter(term == 'div_groupHigh') %>% pull(estimate)
```

```{r}
res_div_lm_tidy %>%
  select(name = type, .otu, value = estimate) %>%
  pivot_wider() %>%
  mutate(
    error = Observed - Calibrated,
  ) %>%
  pull(error) %>%
  summary
mean_eff_lfc 
```

Can see that the errors are the same for all species, and nearly equal to the negative LFC in mean efficiency.
A small difference is expected from the addition of pseudocounts for the species regression.

To understand why the mean efficiency is lower, can try a scatterplot of mean efficiency against diversity with samples colored by most abundant species.

```{r}
div1 <- div %>%
  left_join(
    momspi_rare %>% sample_data %>% as_tibble,
    by = '.sample'
  ) %>%
  mutate(
    top_species_plot = fct_lump_n(top_species_cal, 7)
  )
```

```{r, fig.dim = c(8, 6)}
colors_top <- RColorBrewer::brewer.pal(8, 'Accent')[c(1,2,3,5,4,6,7,8)] 
div1 %>%
  ggplot(aes(diversity, mean_efficiency, color = top_species_plot)) +
  facet_grid(type~index) +
  scale_x_log10() +
  scale_y_log10() +
  ggrastr::rasterize(geom_point(alpha = 0.6), dpi = 300) +
  stat_smooth(aes(group = 'none'), color = 'black') +
  labs(color = 'Top species', x = 'Diversity', y = 'Mean efficiency') +
  scale_color_manual(
    values = colors_top,
    labels = function(x) str_replace_all(x, '_', ' ')
  )
```