### 2022-01-16 conclusion.Rmd
Add placeholder for conclusion section

# Conclusion

_This section is a placeholder for a short (2-3 paragraph) conclusion. We are planning to keep 'discussion' to the earlier sections and appendix, at least for initial submission._