### 2022-02-06 notebook/_posts/2022-01-25-explore-gut-data-2/_cut.Rmd

### CHECKs

```{r}
dataset <- 'KieserS_2018'
tse <- curatedMetagenomicData(
  str_c(dataset, ".relative_abundance"),
  dryrun = FALSE
)
stopifnot(identical(length(tse), 1L))
nm <- names(tse)
tse <- tse[[1]]
ps <- mia::makePhyloseqFromTreeSummarizedExperiment(
  tse,
  abund_values = "relative_abundance"
) %>%
  rename_with_sample_data(janitor::make_clean_names) %>%
  rename_with_tax_table(janitor::make_clean_names) %>%
  transform_sample_counts(~. / sum(.))
all.equal(
  tse %>% rownames %>% str_subset('Streptococcus'),
  ps %>% taxa_names %>% str_subset('Streptococcus')
)
```