### 2022-02-06 notebook/_posts/2022-01-25-explore-gut-data-2/explore-gut-data-2.Rmd

---
title: "Explore gut datasets, 2"
description: |
author:
  - name: Michael R. McLaren
    url: {}
categories:
  - R
date: 2022-01-25
draft: false
output:
  distill::distill_article:
    self_contained: false
    dev: svg
---

Purpose: Examine phylum-level variation w/ clinical condition in different studies.

```{r}
library(here)
library(tidyverse)
library(speedyseq)
library(metacal)

library(cowplot)
theme_set(theme_cowplot())
library(ggbeeswarm)
```

# Studies

```{r}
fs::dir_ls(here('notebook/_data/gut'))

study <- c(
  'HMP_2019_ibdmdb',
  # 'HMP_2019_t2d',
  'KieserS_2018',
  # 'QinJ_2012',
  'RaymondF_2016',
  'ThomasAM_2018b'
)
tbl <- tibble(study = study) %>%
  mutate(
    file = str_glue('2021-03-31.{study}.relative_abundance.phyloseq.rds'),
    path = here('notebook/_data/gut/', file),
    ps = map(path, readRDS)
  )
```

Aim: For each study, get a tibble with the processed phyloseq object and the relevant clinical variable.
The clinical var will usually be `study_condition`, but need to check, and also set the factor levels.

Note, processing that will be applied to all ps objects will be done after.

Thomas et al colorectal cancer

```{r}
ps <- tbl %>% filter(study == 'ThomasAM_2018b') %>% pull(ps) %>% .[[1]]
sam <- ps %>% sample_data %>% as_tibble
sam %>% glimpse
sam %>% dplyr::count(body_site, study_condition, disease)
sam %>% dplyr::count(age_category)
sam %>% dplyr::count(gender)
sam %>% dplyr::count(antibiotics_current_use)
```

HERE

```{r}
process_ps <- function(ps) {
  ps %>%
    filter_tax_table(kingdom == 'Bacteria') %>%
    transform_sample_counts(close_elts) %>%
    mutate_tax_table(.otu = species)
}
tbl <- tbl %>%
  mutate(
    ps_proc = map(ps, process_ps)
  )

```

- prep each dataset
- write some functions to do phylum level plots, compute variation, and compute effects on the phyla
- perhaps do the same for mean eff supposing a G+/G- model.

NOTE: We might want to do this case by case; or do the bias both ways (since some datasets may have a high B:F ratio, others low)

## helper functions

# Acute diarrhoea - Kieser et al 2018

```{r}
ps <- tbl %>% filter(study == 'KieserS_2018') %>% pull(ps_proc) %>% .[[1]]
sam <- ps %>% sample_data %>% as_tibble
sam %>% glimpse
sam %>% dplyr::count(body_site, age_category, study_condition, disease)
sam %>% dplyr::count(study_condition, uncurated_metadata)
```

Let's check the phylum proportions in healthy and sick kids,

```{r}
ps_phy <- ps %>% tax_glom('phylum') %>%
  mutate_tax_table(.otu = phylum)
x_phy <- ps_phy %>% as_tibble
```

```{r}
ps_phy %>%
  filter_taxa2(~median(.) > 0) %>%
  as_tibble %>%
  with_groups(c(phylum, study_condition), summarize,
    across(.abundance, median)) %>%
  pivot_wider(names_from = study_condition, values_from = .abundance)
```

```{r}
ps_phy %>%
  filter_taxa2(~median(.) > 0) %>%
  as_tibble %>%
  ggplot(aes(y = study_condition, x = 1e-5 + .abundance, color =
      study_condition)) +
  facet_wrap(~phylum, ncol = 1) +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

Healthy kids have high proportions of the G+ Actinobacteria and Firmicutes, as well as Bacteroidetes; sick are dominated by Proteobacteria.

This finding is consistent with some of the observations of the paper.

After reading the article, I learned that
- many-most of the children were identified as having organisms that are considered gut pathogens / known causes of diarrhoea, which are all in Proteobacteria (E. coli, Vibrio, Campylobacter, Shigella, Aeromonas)
- these samples are a random subset of samples that were sequenced by 16S sequencing.
- 16S qPCR measurements were made to gain some absolute abundance info
- The authors report a fairly large increase in Streptococcus proportion, which is in Firmicutes. This finding does not seem to mesh with what I'm seeing in the shotgun profiles
- The authors report various other DA results

There is still important questions and interesting things I might look into, such as the disagreement between 16s and shotgun results. However, a relatively straightforward story could already be suggested.

Observations
- large phylum+ level differences, and in particular large differences in the G+ vs G- taxa, with a high fraction of controls being the G+ Actino and Firm, and a high fraction of cases being the G- Proteobacteria.
- Proteobacteria is known to contain many gut pathogens
<!--  -->

Things to look into
- the alpha diversity within phyla within patients, to see the effective number of species (and genera) within phyla
- What fraction of comm's within patients by most abundant 1 species - want to understand whether mean eff var would be driven by just one species
  - Either way, I think I can make the point that phylum-level conservation is important when we do a cross-species analysis, since it won't always be the same species (various Proteobacteria), I'm guessing
- Specifically dig into the Proteobacteria - see what species/genera are contributing to the Prot fraction in each individual esp. with diarrhea to see if true that it's not just one species, but various, but all within the phylum
- Fraction of comms made up of G+ versus G- phyla
  - perhaps use this as a proxy for mean efficiency, rather than bothering to simulate
<!--  -->

### Proteobacteria

Question: Which Proteobacteria species are driving the high Proteobacteria proportion in sick kids?

```{r}
prot_spp <- ps %>%
  filter_tax_table(phylum == 'Proteobacteria') %>%
  filter_taxa2(~max(.) > 1e-2) %>%
  as_tibble
prot_genus <- ps %>%
  filter_tax_table(phylum == 'Proteobacteria') %>%
  tax_glom('genus') %>%
  as_tibble
```

```{r}
prot_genus %>%
  mutate(across(genus, fct_reorder, .abundance)) %>%
  ggplot(aes(y = genus, x = 1e-5 + .abundance, color = study_condition)) +
  theme_minimal_hgrid() +
  geom_vline(xintercept = c(1e-2, 0.1, 1), color = 'grey') +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

```{r}
prot_spp %>%
  mutate(across(.otu, fct_reorder, .abundance)) %>%
  ggplot(aes(y = .otu, x = 1e-5 + .abundance, color = study_condition)) +
  theme_minimal_hgrid() +
  geom_vline(xintercept = c(1e-2, 0.1, 1), color = 'grey') +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

We can see that E. coli and Klebsiella spp. are the drivers of the high Proteobacteria proportions, with Vibrio in one patient also being substantial (10%).
Escherichia and Klebsiella are both in the family Enterobacteriaceae.

### Streptococcus

```{r}
strep <- ps %>%
  filter_tax_table(genus == 'Streptococcus') %>%
  tax_glom('genus') %>%
  as_tibble
strep_spp <- ps %>%
  filter_tax_table(genus == 'Streptococcus') %>%
  filter_taxa2(~max(.) > 3e-3) %>%
  as_tibble
```

```{r}
strep %>%
  ggplot(aes(y = study_condition, x = 1e-5 + .abundance, color =
      study_condition)) +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

```{r}
strep_spp %>%
  ggplot(aes(y = study_condition, x = 1e-5 + .abundance, color =
      study_condition)) +
  facet_wrap(~species, ncol = 1) +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

```{r}
ps %>%
  filter_tax_table(family == "Streptococcaceae") %>%
  tax_glom('family') %>%
  as_tibble %>%
  ggplot(aes(y = study_condition, x = 1e-5 + .abundance, color =
      study_condition)) +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

We also don't see the result at the family level.

In short, the Metaphlan3 profiles do not support the Streptococcus associations reported in the study, which were found using 16S sequencing and shotgun sequencing (analyzed with Metaphlan2).

## Check for any clades with large differences between conditions

```{r}
check_diffs <- function(ps, taxrank, fn = mean) {
  ps %>% tax_glom(taxrank) %>% as_tibble %>%
    with_groups(c(phylum:species, study_condition), 
      summarize, across(.abundance, fn)) %>%
    pivot_wider(names_from = study_condition, values_from = .abundance) %>%
    mutate(diff = acute_diarrhoea - control) %>%
    arrange(desc(abs(diff)))
}
```

```{r}
check_diffs(ps, 'species', mean) %>% slice(1:4)
check_diffs(ps, 'genus', mean) %>% slice(1:4)
check_diffs(ps, 'family', mean) %>% slice(1:4)
check_diffs(ps, 'order', mean) %>% slice(1:4)
check_diffs(ps, 'class', mean) %>% slice(1:4)
check_diffs(ps, 'phylum', mean) %>% slice(1:4)
```

# Antibiotics - Raymond et al 2016

```{r}
ps <- tbl %>% filter(study == 'RaymondF_2016') %>% pull(ps_proc) %>% .[[1]]
sam <- ps %>% sample_data %>% as_tibble
sam %>% glimpse
sam %>% count(body_site, age_category, study_condition, disease)
sam %>% count(study_condition, antibiotics_family)
sam %>% count(study_condition, days_from_first_collection)
```

Note, the 'study_condition' varies by timepoint for the same subject, so isn't useful for filtering or coloring the subjects who received the antibiotic.
We therefore need to create a new variable.

```{r}
sam <- sam %>%
  with_groups(subject_id, mutate,
    subject_group = case_when(
      'yes' %in% antibiotics_current_use ~ 'antibiotics',
      TRUE ~ 'control'
    ),
    days_idx = rank(days_from_first_collection) %>% as.factor
  )
sample_data(ps) <- sample_data(sam)
sam %>% count(subject_group, days_from_first_collection, days_idx)
```

Let's check the phylum proportions,

```{r}
ps_phy <- ps %>% tax_glom('phylum') %>%
  mutate_tax_table(.otu = phylum)
x_phy <- ps_phy %>% as_tibble
focal_phyla <- c('Actinobacteria', 'Bacteroidetes', 'Firmicutes',
  'Proteobacteria')
```

```{r}
x_phy %>%
  filter(phylum %in% focal_phyla, subject_group == 'antibiotics') %>%
  with_groups(c(phylum, study_condition), summarize,
    across(.abundance, median)) %>%
  pivot_wider(names_from = study_condition, values_from = .abundance)
```

```{r}
x_phy %>%
  filter(phylum %in% focal_phyla, subject_group == 'antibiotics') %>%
  ggplot(aes(y = study_condition, x = 1e-5 + .abundance, color =
      study_condition)) +
  facet_wrap(~phylum, ncol = 1) +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

```{r}
x_phy %>%
  filter(phylum %in% focal_phyla, subject_group == 'antibiotics') %>%
  ggplot(aes(days_idx, 1e-5 + .abundance, color = phylum)) +
  facet_wrap(~phylum, scales = 'free_y', ncol = 1) +
  scale_y_log10() +
  scale_color_brewer(type = 'qual', palette = 2) +
  geom_line(aes(group = interaction(subject_id, phylum))) +
  geom_point()
```

There seems to be some increase in G- phyla with rx administration, though not a huge effect.
I wonder if the effect would seem larger if we supposed a lysis bias favoring G-?

```{r}
bias <- ps_phy %>% tax_table %>% as_tibble %>%
  mutate(
    efficiency = case_when(
      phylum %in% c('Actinobacteria', 'Firmicutes') ~ 1/10,
      TRUE ~ 1
    )
  ) %>%
  select(.otu, efficiency) %>%
  deframe
ps_phy_cal <- ps_phy %>% calibrate(bias)
```

```{r}
ps_phy_cal %>% 
  as_tibble %>%
  filter(phylum %in% focal_phyla, subject_group == 'antibiotics') %>%
  ggplot(aes(days_idx, .abundance, color = phylum)) +
  facet_wrap(~phylum, scales = 'free_y', ncol = 1) +
  scale_y_log10() +
  scale_color_brewer(type = 'qual', palette = 2) +
  geom_line(aes(group = interaction(subject_id, phylum))) +
  geom_point()
```

```{r}
ps_phy_cal %>% 
  sample_data %>%
  as_tibble %>%
  ggplot(aes(days_idx, .mean_efficiency)) +
  scale_y_log10() +
  expand_limits(y = c(0.1, 1)) + 
  scale_color_brewer(type = 'qual', palette = 2) +
  stat_summary(color = 'darkred', position = position_nudge(x = -0.1)) +
  geom_line(aes(group = subject_id), alpha = 0.5) +
  geom_point(alpha = 0.5)
```

We see sizable variation in the mean efficiency, but just a small change in the mean during and after treatment.

Note, the variation would be larger if we enforced a difference between the average efficiency for the phylum but also allowed for variation within a phylum.

## Check for any clades with large differences between conditions

```{r}
check_diffs <- function(ps, taxrank, fn = mean) {
  ps %>% tax_glom(taxrank) %>% as_tibble %>%
    with_groups(c(phylum:species, study_condition), 
      summarize, across(.abundance, fn)) %>%
    pivot_wider(names_from = study_condition, values_from = .abundance) %>%
    mutate(diff = cephalosporins - control) %>%
    arrange(desc(abs(diff)))
}
```

```{r}
check_diffs(ps, 'species', mean) %>% slice(1:2)
check_diffs(ps, 'genus', mean) %>% slice(1:2)
check_diffs(ps, 'family', mean) %>% slice(1:2)
check_diffs(ps, 'order', mean) %>% slice(1:2)
check_diffs(ps, 'class', mean) %>% slice(1:2)
check_diffs(ps, 'phylum', mean) %>% slice(1:2)
```

# Colorectal cancer

```{r}
ps <- tbl %>% filter(study == 'ThomasAM_2018b') %>% pull(ps_proc) %>% .[[1]]
sam <- ps %>% sample_data %>% as_tibble
sam %>% glimpse
sam %>% count(study_condition, disease)
sam %>% count(body_site, age_category, study_condition)
```

Let's check the phylum proportions,

```{r}
ps_phy <- ps %>% tax_glom('phylum') %>%
  mutate_tax_table(.otu = phylum)
x_phy <- ps_phy %>% as_tibble
focal_phyla <- c('Actinobacteria', 'Bacteroidetes', 'Firmicutes',
  'Proteobacteria')
```

```{r}
x_phy %>%
  filter(phylum %in% focal_phyla) %>%
  with_groups(c(phylum, study_condition), summarize,
    across(.abundance, median)) %>%
  pivot_wider(names_from = study_condition, values_from = .abundance)
```

```{r}
x_phy %>%
  filter(phylum %in% focal_phyla) %>%
  ggplot(aes(y = study_condition, x = 1e-5 + .abundance, color =
      age_category)) +
  facet_wrap(~phylum, ncol = 1) +
  scale_color_brewer(type = 'qual') +
  scale_x_log10() +
  geom_quasirandom(groupOnX = FALSE)
```

Can see that the B:F ratio is somewhat lower in cases.

```{r}
bias <- ps_phy %>% tax_table %>% as_tibble %>%
  mutate(
    efficiency = case_when(
      phylum %in% c('Actinobacteria', 'Firmicutes') ~ 1/10,
      TRUE ~ 1
    )
  ) %>%
  select(.otu, efficiency) %>%
  deframe
ps_phy_cal <- ps_phy %>% calibrate(bias)
sam_cal <- ps_phy_cal %>% 
  sample_data %>%
  as_tibble
```

```{r}
sam_cal %>%
  ggplot(aes(study_condition, .mean_efficiency)) +
  scale_y_log10() +
  expand_limits(y = c(0.1, 0.3)) + 
  scale_color_brewer(type = 'qual', palette = 2) +
  stat_summary(color = 'darkred', position = position_nudge(x = -0.1)) +
  geom_point(alpha = 0.5)
```

```{r}
sam_cal %>%
  lm(data = ., log(.mean_efficiency) ~ study_condition) %>%
  summary
```

## Check for any clades with large differences between conditions

```{r}
check_diffs <- function(ps, taxrank, fn = mean) {
  ps %>% tax_glom(taxrank) %>% as_tibble %>%
    with_groups(c(phylum:species, study_condition), 
      summarize, across(.abundance, fn)) %>%
    pivot_wider(names_from = study_condition, values_from = .abundance) %>%
    mutate(diff = CRC - control) %>%
    arrange(desc(abs(diff)))
}
```

```{r}
check_diffs(ps, 'species', mean) %>% slice(1:5)
check_diffs(ps, 'genus', mean) %>% slice(1:5)
check_diffs(ps, 'family', mean) %>% slice(1:5)
check_diffs(ps, 'order', mean) %>% slice(1:5)
check_diffs(ps, 'class', mean) %>% slice(1:5)
check_diffs(ps, 'phylum', mean) %>% slice(1:5)
```

# Finding available datasets

note, first need to set x to the master sample data table

```{r}
x %>% dplyr::count(treatment) %>% print(n=Inf)
x %>% dplyr::count(antibiotics_current_use)
x %>% dplyr::count(antibiotics_family) %>% print(n=Inf)
```

studies that involve an antibiotics treatment?

```{r}
x %>% 
  dplyr::filter(antibiotics_family == 'cephalosporins') %>%
  dplyr::count(study_name)
x %>% 
  dplyr::filter(study_name == 'RaymondF_2016') %>%
  dplyr::count(antibiotics_current_use, antibiotics_family)
x %>% 
  dplyr::filter(study_name == 'RaymondF_2016')
```

## Diarrhoea

looking at study conditions shows there are two studies with accute diarrhoea

```{r}
ps1 <- curatedMetagenomicData::curatedMetagenomicData(
  "DavidLA_2015.relative_abundance", 
  dryrun = FALSE
) %>% 
  .[[1]] %>%
  mia::makePhyloseqFromTreeSummarizedExperiment(
    abund_values = "relative_abundance"
  ) %>%
  rename_with_sample_data(janitor::make_clean_names) %>%
  rename_with_tax_table(janitor::make_clean_names)

# ps1 %>% sample_data %>% as_tibble %>% glimpse
sam1 <- ps1 %>% sample_data %>% as_tibble
sam1 %>% dplyr::count(study_condition)
sam1 %>% dplyr::count(age_category)
sam1 %>% dplyr::count(body_subsite)
sam1 %>% dplyr::count(age_category, study_condition)
sam1 %>% dplyr::count(age_category, study_condition, body_subsite)
```

# CUT

## Phylum proportions by condition

```{r}
x <- ps %>%
  transform_sample_counts(close_elts) %>%
  tax_glom('phylum') %>%
  as_tibble
```

```{r}
x %>%
  with_groups(phylum, summarize, across(.abundance, median))
```

```{r}
focal_phyla <- c('Actinobacteria', 'Bacteroidetes', 'Firmicutes',
  'Proteobacteria')
x %>%
  filter(phylum %in% focal_phyla) %>%
  ggplot(aes(.abundance, fill = disease)) +
  facet_wrap(~phylum, scales = 'fixed', ncol = 1) +
  scale_x_log10() +
  geom_histogram()
```

```{r}
focal_phyla <- c('Actinobacteria', 'Bacteroidetes', 'Firmicutes',
  'Proteobacteria')
x %>%
  filter(phylum %in% focal_phyla) %>%
  # ggplot(aes(y = .abundance, x = disease)) +
  ggplot(aes(y = .abundance, x = (disease != 'healthy') * 1L)) +
  facet_wrap(~phylum, scales = 'free_y', ncol = 1) +
  scale_y_log10() +
  geom_quasirandom() +
  stat_smooth(method = 'lm')
```

```{r}
y <- x %>%
  filter(phylum %in% focal_phyla) %>%
  with_groups(phylum, nest) %>%
  mutate(
    fit = map(data, ~lm(data = .x, log(.abundance) ~ disease)),
    fit = map(fit, broom::tidy)
  ) %>%
  select(-data) %>%
  unnest(fit)
y %>%
  filter(term != '(Intercept)')
```

## Bias

Let's come up with a way to simulate bias vectors with a particular property.