### 2022-02-06 notebook/_posts/2022-01-30-hmp-bias/hmp.Rmd

# Setup

## code to create ps object (not run)

```{r, eval = FALSE}
# It's helpful to load the Bioconductor cruft first so that it doesn't mask the
# functions we actually want to use.
HMP16SData::V13()
library(speedyseq)

ps_v13 <- HMP16SData::V13() %>% HMP16SData::as_phyloseq()
ps_v35 <- HMP16SData::V35() %>% HMP16SData::as_phyloseq()

ps_v13 %>% sample_data %>% glimpse
ps_v35 %>% sample_data %>% glimpse

# To merge, we first need to drop the phylogenetic tree, and adjust the sample
# names.
intersect(
  sample_names(ps_v13),
  sample_names(ps_v35)
) %>% length

ps_v13 <- ps_v13 %>% 
  mutate_sample_data(
    region = 'V13', 
    specimen_id = .sample,
    .sample = str_glue('X{specimen_id}_{region}')
  )
ps_v35 <- ps_v35 %>% 
  mutate_sample_data(
    region = 'V35', 
    specimen_id = .sample,
    .sample = str_glue('X{specimen_id}_{region}')
  )

ps_merge <- merge_phyloseq(
  ps_v13 %>% {.@phy_tree <- NULL; .}, 
  ps_v35 %>% {.@phy_tree <- NULL; .}
)

# let's clean up names and aggregate to the genus level,

ps <- ps_merge %>%
  rename_with_sample_data(janitor::make_clean_names) %>%
  rename_with_tax_table(janitor::make_clean_names) %>%
  tax_glom('genus')
ps
rm(ps_merge, ps_v13, ps_v35)
dir.create('_output')
saveRDS(ps, '_output/hmp-16s-genus.rds')
```

```{r}
```

## start here

```{r}
library(here)
library(tidyverse)
library(speedyseq)
library(metacal)

library(cowplot)
theme_set(theme_cowplot())
library(ggbeeswarm)
```

```{r}
ps <- readRDS('_output/hmp-16s-genus.rds') %>%
  mutate_sample_data(., sample_sum = sample_sums(.))
ps
sam <- ps %>% sample_data %>% as_tibble
```

```{r}
sam %>% count(specimen_id) %>% count(n)
control_specimens <- sam %>% count(specimen_id) %>% filter(n>1) %>%
  pull(specimen_id)
```

To estimate bias, we'll work with the specimens with both regions.

TODO: Filter samples to a smaller more workable set, perhaps just gut.

```{r}
sam %>% 
  filter(specimen_id %in% control_specimens) %>%
  count(hmp_body_site)
```

TODO: Filter genera to high prevalence ones.

```{r}
ps1 <- ps %>% 
  filter_sample_data(
    hmp_body_subsite == 'Stool', 
    specimen_id %in% control_specimens
  ) %>%
  filter_taxa2(~sum(.) > 10)
```

```{r}
ps1 %>% sample_sums %>% sort %>% summary
ps1 %>% sample_sums %>% sort %>% head(20)
```

```{r}
ps2 <- ps1 %>% filter_sample_data(sample_sum >= 3000)
control_specimens_filt <- ps2 %>% sample_data %>% as_tibble %>% 
  count(specimen_id) %>% filter(n>1) %>%
  pull(specimen_id)
```

```{r}
x <- ps2 %>%
  transform_sample_counts(close_elts) %>%
  as_tibble %>%
  with_groups(.otu, summarize, 
    across(.abundance, .names = '{.fn}',
      .fns = list(mean = mean, median = median, max = max)
    )
  ) %>%
  arrange(max)
x %>% print(n=10)
x %>% count(max > 1e-2)
focal_taxa <- x %>% filter(max > 1e-2) %>% pull(.otu)
```

```{r}
ps_fit <- ps2 %>%
  prune_taxa(focal_taxa, .)
```

```{r}
ps_list <- list('V13', 'V35') %>% set_names %>%
  map(~filter_sample_data(ps_fit, region == .x)) %>%
  map(filter_sample_data, 
    specimen_id %in% control_specimens_filt
  ) %>%
  map(mutate_sample_data, .sample = str_c('X', specimen_id)) %>%
  map(transform_sample_counts, ~. + 0.3)
bias_fit <- estimate_bias(
  ps_list$V13,
  ps_list$V35,
  boot = TRUE
)
bias <- bias_fit %>% coef
```

```{r}
tax <- ps_fit %>% tax_table %>% as_tibble %>%
  left_join(summary(bias_fit)$coefficients, by = c('.otu' = 'taxon'))
```

```{r}
tax %>%
  ggplot(aes(estimate, fill = phylum)) +
  scale_x_log10() +
  scale_fill_brewer(type = 'qual') +
  geom_histogram()
```

```{r}
tax %>%
  mutate(.otu = fct_reorder(.otu, estimate)) %>%
  ggplot(aes(.otu, estimate, color = phylum,
      ymin = estimate / gm_se^2, ymax = estimate * gm_se^2)) +
  scale_color_brewer(type = 'qual', palette = 2) +
  geom_hline(yintercept = 1, color = "grey") +
  geom_pointrange() +
  scale_y_log10() +
  coord_flip()
```

The CIs are pretty tight, but I wonder how well the model really does?

### check model fit

```{r}
props <- ps_list %>%
  map(transform_sample_counts, close_elts) %>%
  map_dfr(as_tibble)
```

```{r}
observed.fitted <- fitted(mc_fit) %>% otu_table(taxa_are_rows = FALSE)
props.fitted <- bind_rows(
  props, 
  as_tibble(observed.fitted) %>% add_column(type = "Fitted")
)
```

### finding better taxa names

```{r}
tax <- ps2 %>% tax_table %>% as_tibble
tax %>% count(genus) %>% filter(n>1)
tax %>% count(str_c(family, genus, sep = '_')) %>% filter(n>1)
tax %>% filter(is.na(family))
```