### 2022-02-08 notebook/case-studies.Rmd
Create summary page and listings for case studies
Add listing pages for each case study, which contain all the posts for
that case study; and a 'Case studies' page that will provide a summary
with links of all analyses for each case study, but for now just links
to the listing pages (or the single post for the marine case study).

---
title: "Case studies"
description: |
  Overview of case-study analyses.
author:
  - name: Michael R. McLaren
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Foliar fungi experiment

[All posts](case-study-fungi.html)

# Vaginal microbiomes of pregnant women

[All posts](case-study-momspi.html)

# Human gut microbiomes

[All posts](case-study-gut.html)

# Microbial growth in marine sediments

[Case study analysis](posts/2022-01-08-lloyd2020evi-case-study/index.html)