### 2022-02-09 notebook/_posts/2022-01-20-vieirasilva2019quan/_cut.Rmd
Revise and render vieirasilva2019quan analysis
Analysis of bias sensitivity of Spearman correlations on QMP profiles
from vieirasilva2019quan.

### Taxonomy

how can I get taxonomy and/or phylogeny for the genus identifiers?

First, create a file with the RDP 16 taxonomy from the DADA2 training file

```
zcat rdp_train_set_16.fa.gz | grep '^>' | sort | uniq -u | sed -e 's/^>//' -e 's/;$//' - > rdp_train_set_16_headers.txt
```

```{r}
cns <- c('domain', 'phylum', 'class', 'order', 'family', 'genus')
cts <- 'cccccc'
# x <- path('~/data/16s-taxonomy-databases/dada2/rdp-trainset-16',
#   'rdp_train_set_16_headers.txt') %>%
#   read_delim(delim = ';', col_names = cns, col_types = cts)
x <- path('~/data/16s-taxonomy-databases/dada2/rdp-trainset-16',
  'rdp_train_set_16_headers.txt') %>%
  read_lines %>%
  enframe %>%
  select(-name) %>%
  separate(value, into = cns, sep = ';', fill = 'right') %>%
  glimpse
```

```{r}
genera <- taxa_names(ps) %>% 
  str_subset('unclassified', negate = TRUE)
genera %in% x$genus
setdiff(genera, x$genus)

unclassified_taxa <- taxa_names(ps) %>% 
  str_subset('unclassified') %>%
  str_extract('(?<=unclassified_).+')
unclassified_taxa %in% x$family
unclassified_taxa %in% x$order
unclassified_taxa %in% x$phylum
unclassified_taxa %in% x$domain
```

try again

```{r}
rnks <- c('domain', 'phylum', 'class', 'order', 'family', 'genus')
x <- path('~/data/16s-taxonomy-databases/rdp/', 
  'RDPClassifier_16S_trainsetNo16_rawtrainingdata',
  'trainset16_022016_headers.txt') %>%
  read_tsv(col_names = c('id', 'path'))
y <- x %>%
  separate(path, c('root', rnks), sep = ';', fill = 'right')

```

ok, giving up..

hmm - check to see if can get from the rrndb tables?
https://rrndb.umms.med.umich.edu/static/download/

### spearman versus linear modeling

NOTE: The results differ when I first omit the samples with missing calprotectin measurements, or not.
perhaps because the ranks of the abundances differ?

```{r}
z <- ps %>%
  filter_sample_data(!is.na(faecal_calprotectin_mg_g)) %>%
  as_tibble %>%
  with_groups(.sample, mutate, proportion = close_elts(.abundance)) %>%
  with_groups(.otu, mutate, 
    across(c(.abundance, proportion, faecal_calprotectin_mg_g), 
      rank, ties.method = 'average',
      .names = '{.col}_rank')
  ) %>%
  with_groups(.otu, nest) %>%
  mutate(
    spearman = map(data, 
      ~cor.test(data = .x,
        ~ .abundance + faecal_calprotectin_mg_g, 
        method = 'spearman', na.action = na.omit)),
    lm = map(data, 
      ~lm(data = .x, .abundance_rank ~ faecal_calprotectin_mg_g_rank,
        na.action = na.omit)),
  ) %>%
  select(-data) %>%
  pivot_longer(c(spearman, lm), names_to = 'test', values_to = 'fit') %>%
  mutate(across(fit, map, broom::tidy)) %>%
  unnest(fit) %>%
  with_groups(test, mutate, p.bh = p.adjust(p.value, method = 'BH'))
z1 <- z %>% filter(test == 'spearman' | term != '(Intercept)')
```

NOTE: The p values are nearly identical as long as I first drop the samples with missing data.
But not if I don't.

```{r}
z1 %>%
  select(.otu, test, p.value) %>%
  pivot_wider(names_from = test, values_from = p.value) %>%
  ggplot(aes(spearman, lm)) +
  geom_abline(color = 'darkgrey') +
  scale_x_log10() +
  scale_y_log10() +
  geom_point()
```

```{r}
z1 %>%
  select(.otu, test, estimate) %>%
  pivot_wider(names_from = test, values_from = estimate) %>%
  ggplot(aes(spearman, lm)) +
  geom_abline(color = 'darkgrey') +
  # scale_x_log10() +
  # scale_y_log10() +
  geom_point()
```

The estimates are related but differ - I suspect this has to do with ties in the abundance (0s) and in the calprotectin (min values / detection limit). Perhaps the problematic taxa are very low prevalence?

### try all tests

```{r}
y <- ps %>%
  # filter_sample_data(!is.na(faecal_calprotectin_mg_g)) %>%
  # prune_taxa(str_subset(taxa_names(.), 'unclassified', negate = TRUE), .) %>%
  as_tibble %>%
  with_groups(.sample, mutate, proportion = close_elts(.abundance)) %>%
  with_groups(.otu, nest) %>%
  mutate(
    calprotectin_qmp_spearman = map(data, 
      ~cor.test(data = .x,
        ~ .abundance + faecal_calprotectin_mg_g, 
        method = 'spearman')),
    calprotectin_rmp_spearman = map(data, 
      ~cor.test(data = .x,
        ~ proportion + faecal_calprotectin_mg_g, 
        method = 'spearman')),
    moisture_qmp_spearman = map(data, 
      ~cor.test(data = .x,
        ~ .abundance + moisture_content_percent, 
        method = 'spearman')),
    moisture_rmp_spearman = map(data, 
      ~cor.test(data = .x,
        ~ proportion + moisture_content_percent, 
        method = 'spearman')),
  ) %>%
  select(-data) %>%
  pivot_longer(contains('spearman'), names_to = 'test', values_to = 'fit') %>%
  mutate(across(fit, map, broom::tidy)) %>%
  unnest(fit) %>%
  with_groups(test, mutate, p.bh = p.adjust(p.value, method = 'BH'))
```

TODO: check that samples with NAs are handled in the spearman tests

```{r}
y.w <- y %>%
  select(.otu, test, estimate, p.bh) %>%
  pivot_wider(names_from = c(test), values_from = c(estimate, p.bh), 
    names_glue = '{test}_{.value}', names_sort = TRUE) %>%
  select(.otu, sort(tidyselect::peek_vars())) %>%
  mutate(across(where(is.numeric), round, 4)) %>%
  arrange(calprotectin_qmp_spearman_p.bh)
```

Note, I'm not sure what the RMP values used in the original study were.
From the manuscript, it's not obvious if the unclassified taxa were removed before the proportions were computed, or if copy number correction was used.

# saved plotting stuff

Mean efficiency distribution:

```{r}
x <- reps %>%
  slice(1) %>%
  pull(actual) %>%
  .[[1]] %>%
  sample_data %>%
  as_tibble

x %>%
  ggplot(aes(x = mean_efficiency, fill = faecal_calprotectin> 50)) +
  scale_x_log10() +
  geom_histogram()
```

```{r}
x %>%
  ggplot(aes(x = faecal_calprotectin, y = mean_efficiency, color = diagnosis)) +
  scale_y_log10() +
  scale_x_log10() +
  geom_point() +
  stat_smooth(aes(group = 1), method = 'lm')
```

# Checks and notes

```{r}
# check na handling
x = tibble(a = c(1,2,3,NA,4,5), b = c(6,5,4,3,2,NA))
cor.test(~a + b, data = x, method = 'spearman') %>% broom::tidy()
cor.test(~a + b, data = x %>% drop_na, method = 'spearman') %>% broom::tidy()
```

Note, I'm not sure what the RMP values used in the original study were.
From the manuscript, it's not obvious if the unclassified taxa were removed before the proportions were computed, or if copy number correction was used.

## Original bias sens analysis

Maintain the sample sums, but perturb the relative abundances.

```{r}
n_taxa <- ntaxa(ps)
# Standard deviation for the distribution of log efficiencies
sigma.log_eff <- 3
# helper function to draw a single random bias vector
rbias <- function() exp(rnorm(n_taxa, 0, sigma.log_eff))
# helper function to calibrate the phyloseq object by the bias vector
my_cal <- function(bias, ps) {
  ps %>%
    calibrate(bias, norm = 'keep')
}

set.seed(42)
# Number of replicates
R <- 3e2
# data frame with R random bias vectors
reps <- seq(R) %>%
  set_names %>%
  map(~rbias()) %>%
  enframe("id", "bias") %>%
  mutate(
    actual = map(bias, my_cal, ps = ps)
  )
```

```{r}
my_stat <- function(ps, otu) {
  data <- ps %>%
    prune_taxa(otu, .) %>%
    as_tibble
  spearman_test <- cor.test(data = data,
    ~ .abundance + faecal_calprotectin, 
    method = 'spearman', exact = FALSE
  )
  spearman_test %>% broom::tidy()
}

#> focal_otu <- 'Veillonella'
original <- my_stat(ps, focal_otu)
reps_veill <- reps %>%
  mutate(
    .otu = 'Veillonella',
    fit = map2(actual, .otu, my_stat)
  ) %>%
  unnest(fit)
reps_strep <- reps %>%
  mutate(
    .otu = 'Streptococcus',
    fit = map2(actual, .otu, my_stat)
  ) %>%
  unnest(fit)
```

```{r}
plot_bias_sens <- function(reps) {
  est_dist <- ggplot(reps, aes(estimate)) +
    geom_histogram() +
    geom_vline(xintercept = original$estimate, color = "darkred") +
    geom_vline(xintercept = 0, color = "grey") +
    labs(title = "Distribution of point estimates", x = "Estimate", y = "Count")
  pval_dist <- ggplot(reps, aes(p.value)) +
    geom_histogram() +
    geom_vline(xintercept = original$p.value, color = "darkred") +
    geom_vline(xintercept = 0.05, color = "grey") +
    scale_x_log10() +
    labs(title = "Distribution of p-values", x = "p-value", y = "Count")
  est_dist / pval_dist
}
# ci_reps <- ggplot(bias_reps, aes(x = estimate, y = as.integer(id))) +
#   geom_point(size = 0.8) +
#   geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
#   geom_point(data = original, aes(y = R*1.05), 
#     size = 0.8, color = "darkred") +
#   geom_errorbarh(data = original, 
#     aes(xmin = conf.low, xmax = conf.high, y = R*1.05),
#     color = "darkred") +
#   geom_vline(xintercept = 0, color = "grey") +
#   theme(axis.ticks.y = element_blank(), axis.text.y = element_blank()) +
#   labs(title = "Replicate 95% CIs", x = "Estimate", 
#     y = "Replicate")
# plot_grid(
#     plot_grid(est_dist, pval_dist, ncol = 1, labels = c("A", "B")),
#     ci_reps,
#     labels = c("", "C")
#     )
```

```{r}
plot_bias_sens(reps_veill)
```

```{r}
plot_bias_sens(reps_strep)
```