### 2022-02-09 notebook/_posts/2022-01-20-vieirasilva2019quan/vieirasilva2019quan.Rmd

---
title: "Explore bias sensitivity of Vieira-Silva et al (2019) QMP analysis"
description: |
author:
  - name: Michael R. McLaren
    url: {}
categories:
  - ref:vieirasilva2019quan
  - bias sensitivity
  - absolute abundance
date: 2022-01-20
draft: false
bibliography: ../../../main.bib
output:
  distill::distill_article:
    self_contained: false
    dev: svg
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  cache = TRUE,
  autodep = TRUE,
  cache.comments = FALSE,
  dpi = 300
)
```

This document reanalyzes the 'Quantitative Microbiome Profiles' of @vieirasilva2019quan.

# Setup

## R setup

```{r libraries}
library(here)
# Tools for microbiome data
library(speedyseq)
# Tools for general purpose data manipulation and plotting
library(tidyverse)
library(fs)
# ggplot helpers
library(ggbeeswarm)
library(ggridges)
library(cowplot)
library(patchwork)
library(scales)
# stats helpers
# library(broom)

library(metacal); packageVersion("metacal")
```

```{r ggplot}
theme_set(theme_cowplot())
```

## Data setup

The genus-level 'Quantitative Microbiome Profiles' from @vieirasilva2019quan are available for download at http://raeslab.org/software/QMP2/, and the sample metadata is available in the supplemental content at the journal site.

```{r}
data_path <- here('notebook/_data/vieirasilva2019quan')
dir_create(data_path)
# Supplementary tables from the nature website; contains the metadata
fn <- path(data_path, 'supplementary-data-1.xlsx')
if (!file_exists(fn)) {
  download.file(
    'https://static-content.springer.com/esm/art%3A10.1038%2Fs41564-019-0483-9/MediaObjects/41564_2019_483_MOESM3_ESM.xlsx',
    fn
  )
  file_chmod(fn, '-w')
}
# QMP proiles from the Raes lab website
fn <- path(data_path, 'qmp-matrix-genus.tsv')
if (!file_exists(fn)) {
  download.file('http://raeslab.org/software/QMP2/QMP.matrix.tsv', fn)
  file_chmod(fn, '-w')
}
```

I'll read in the metadata and QMP profiles into a phyloseq object.
The metadata is Table S1 of the Excel spreadsheet.
There is an extra value in the BMI column of Table S1 that will be dropped in the import step below.

```{r}
otu <- path(data_path, 'qmp-matrix-genus.tsv') %>%
  read.table() %>%
  as('matrix') %>%
  otu_table(taxa_are_rows = FALSE)
sam <- path(data_path, 'supplementary-data-1.xlsx') %>%
  readxl::read_excel(sheet = 'Table S1', range = 'A2:L174', na = "NA")
original_column_names <- names(sam)
ps <- phyloseq(
  otu, 
  sam %>% janitor::clean_names() %>% sample_data
) %>%
  rename_sample_data(
   faecal_calprotectin = faecal_calprotectin_mg_g,
   moisture_content = moisture_content_percent
  )
stopifnot(setequal(sample_names(ps), sample_names(otu)))
sam <- ps %>% sample_data %>% as_tibble %>% glimpse
```

# Explore data

There are `r nsamples(ps)`, with the following counts by disease status,

```{r}
sam %>% count(diagnosis)
```

The QMP profiles are such that the sum within a sample should equal the reported cell counts per gram in the sample data; let's check.

```{r}
all.equal(sample_sums(ps), sam$average_faecal_cell_count_cells_g)
cor(sample_sums(ps), sam$average_faecal_cell_count_cells_g)
```

This small difference may be down to rounding error, since apparently the entires in the QMP matrix have been rounded to integers:

```{r}
otu %>% c %>% {all.equal(., round(.))}
```

It would be nice to have read counts; however, I don't think it's possible to obtain read counts from the available information here.
Perhaps it would be possible to reverse the QMP calculation if we had the copy-number-variation information.

Since we lack read-count information, we will ignore the sequencing-count noise, which was also done in the original paper.

Our primary covariates of interest are Faecal calprotectin and moisture content.

```{r}
sam %>% select(faecal_calprotectin, moisture_content) %>% summary
```

```{r}
sam %>%
  ggplot(aes(x = faecal_calprotectin)) +
  scale_x_log10() +
  geom_histogram()
```

Note that a large fraction of calprotectin measurements are precisely at 30, suggesting this is a value assigned to samples that were below the detection limit of the assay.
Also, the values vary over multiple orders of magnitude.

```{r}
sam %>%
  ggplot(aes(x = moisture_content)) +
  geom_histogram()
```

In contrast, the moisture varies relatively continuously

# Reproduce QMP Spearman analysis

Let's try to reproduce the results from the paper.
We'll focus on the Spearman correlations of the genera absolute abundances with faecal calprotectin and stool moisture content.
The relevant manuscript text is

> Applying QMP on the combined PSC/IBD/mHC data set, we identified 11 genera as significantly correlat-ing  with  faecal  calprotectin  concentrations  (Spearman  correlation,  n= 149, FDR < 0.05; Fig. 3a and Supplementary Table 6). Stool mois-ture content had 45 associated taxa (Spearman correlation, n= 143, FDR <    0.05;  Supplementary  Table  6).  Combining  both  variables,  only  6  out  of  11  positive(+)  or  negative(−)  calprotectin  associations  remained significant when controlling for faecal water content vari-ation: Anaerostipes(−), Escherichia(+), Fusobacterium(+), Gemmiger(−), Streptococcus(+)  and  Veillonella(+)  (nested  linear  model  on  QMP  rank  abundances,  n= 133,  FDR < 0.05;  Fig. 3a  and  Supplementary  Table  6).

A summary table with the Spearman results are Table S6 of the supplementary data Excel sheet,

```{r}
study_res <- path(data_path, 'supplementary-data-1.xlsx') %>%
  readxl::read_excel(sheet = 'Table S6', na = "NA")
```

```{r}
qmp_spearman <- ps %>%
  prune_taxa(str_subset(taxa_names(.), 'unclassified', negate = TRUE), .) %>%
  as_tibble %>%
  with_groups(.otu, nest) %>%
  mutate(
    qmp_calprotectin_spearman = map(data, 
      ~cor.test(data = .x,
        ~ .abundance + faecal_calprotectin, 
        method = 'spearman', exact = FALSE, continuity = TRUE)),
    qmp_moisture_spearman = map(data, 
      ~cor.test(data = .x,
        ~ .abundance + moisture_content, 
        method = 'spearman', exact = FALSE, continuity = TRUE)),
  ) %>%
  select(-data) %>%
  pivot_longer(contains('spearman'), names_to = 'test', values_to = 'fit') %>%
  mutate(across(fit, map, broom::tidy)) %>%
  unnest(fit) %>%
  with_groups(test, mutate, p.bh = p.adjust(p.value, method = 'BH'))
```

Compare Calprotectin results,

```{r}
study_res %>%
  select(Genera, starts_with('Calprotectin') & ends_with('QMP')) %>%
  arrange(`Calprotectin  Spearman.P QMP`) %>%
  slice_head(n=12)
qmp_spearman %>%
  filter(test == 'qmp_calprotectin_spearman') %>%
  arrange(p.value) %>% 
  select(.otu, estimate, p.value, p.bh) %>%
  slice_head(n=12)
```

The results are similar but not identical.

Compare moisture results,

```{r}
study_res %>%
  select(Genera, starts_with('Moisture') & ends_with('QMP')) %>%
  arrange(`Moisture Spearman.P QMP`) %>%
  slice_head(n=12)
qmp_spearman %>%
  filter(test == 'qmp_moisture_spearman') %>%
  arrange(p.value) %>% 
  select(.otu, estimate, p.value, p.bh) %>%
  slice_head(n=12)
```

Here the estimate and p-values seem identical up to rounding, but the FDR-corrected p values differ somewhat.

Note, I filtered the OTUs that lack genus classifications prior to doing the testing, which I suspect is what was done in the original study.
If these OTUs are kept, the FDR-corrected p-values are larger, due to the larger number of tests.
But the study results have smaller p-values and only list the OTUs with genus names in the table.

## Visually inspect variation for top hits

### Calprotectin

```{r}
top_hits <- qmp_spearman %>% 
  filter(test == 'qmp_calprotectin_spearman') %>%
  slice_min(p.bh, n = 9)
x <- ps %>%
  filter_sample_data(!is.na(faecal_calprotectin)) %>%
  prune_taxa(top_hits$.otu, .) %>%
  as_tibble %>%
  mutate(across(.otu, factor, levels = top_hits$.otu))
```

To visualize the spearman correlation, plot and perform a linear regression on the rank-transformed values,

```{r, fig.dim = c(8,8)}
x %>%
  with_groups(.otu, mutate,
    across(c(faecal_calprotectin, .abundance), rank),
  ) %>%
  ggplot(aes(faecal_calprotectin, .abundance)) +
  facet_wrap(~.otu) +
  coord_fixed() +
  geom_point() +
  stat_smooth(method = 'lm') +
  labs(y = 'rank(abundance)', x = 'rank(faecal calprotectin)')
```

Many of the top hits have relatively low pravalence, and it appears that the correlation is being driven by differential prevalence in the two groups of samples where calprotectin has its min value, or is above its min value.
For these taxa, we might expect bias to have less impact on the associations.

### Moisture

```{r}
top_hits <- qmp_spearman %>% 
  filter(test == 'qmp_moisture_spearman') %>%
  slice_min(p.bh, n = 9)
x <- ps %>%
  filter_sample_data(!is.na(moisture_content)) %>%
  prune_taxa(top_hits$.otu, .) %>%
  as_tibble %>%
  mutate(across(.otu, factor, levels = top_hits$.otu))
```

```{r, fig.dim = c(8,8)}
x %>%
  with_groups(.otu, mutate,
    across(c(moisture_content, .abundance), rank),
  ) %>%
  ggplot(aes(moisture_content, .abundance)) +
  facet_wrap(~.otu) +
  coord_fixed() +
  geom_point() +
  stat_smooth(method = 'lm') +
  labs(y = 'rank(abundance)', x = 'rank(moisture content)')
```

Here we instead see that top taxa as being more prevalent.

## Check if hits are driven by differential prevalence

One way we can do this is to also do Spearman tests after converting abundances to presence/absence (0 if absent, 1 if present).
If the original result is driven by prevalence, then the estimates should be similar to when the QMP abundance was used.

```{r}
qmp_presence <- ps %>%
  prune_taxa(str_subset(taxa_names(.), 'unclassified', negate = TRUE), .) %>%
  as_tibble %>%
  mutate(
    presence = (.abundance > 0) * 1
  ) %>%
  with_groups(.otu, nest) %>%
  mutate(
    qmp_calprotectin = map(data, 
      ~cor.test(data = .x,
        ~ .abundance + faecal_calprotectin, 
        method = 'spearman', exact = FALSE, continuity = TRUE)),
    presence_calprotectin = map(data, 
      ~cor.test(data = .x,
        ~ presence + faecal_calprotectin, 
        method = 'spearman', exact = FALSE, continuity = TRUE)),
    # cond.qmp_calprotectin = map(data,
    #   ~cor.test(data = .x %>% filter(.abundance > 0),
    #     ~ .abundance + faecal_calprotectin, 
    #     method = 'spearman', exact = FALSE, continuity = TRUE)),
    qmp_moisture = map(data, 
      ~cor.test(data = .x,
        ~ .abundance + moisture_content, 
        method = 'spearman', exact = FALSE, continuity = TRUE)),
    presence_moisture = map(data, 
      ~cor.test(data = .x,
        ~ presence + moisture_content, 
        method = 'spearman', exact = FALSE, continuity = TRUE)),
    # cond.qmp_moisture = map(data,
    #   ~cor.test(data = .x %>% filter(.abundance > 0),
    #     ~ presence + moisture_content, 
    #     method = 'spearman', exact = FALSE, continuity = TRUE)),
  ) %>%
  select(-data) %>%
  pivot_longer(contains(c('qmp', 'presence')), 
    names_to = 'test', values_to = 'fit') %>%
  mutate(across(fit, map, broom::tidy)) %>%
  unnest(fit) %>%
  with_groups(test, mutate, p.bh = p.adjust(p.value, method = 'BH'))
```

```{r}
qmp_presence_wide <- qmp_presence %>%
  select(.otu, test, estimate, p.value, p.bh) %>%
  separate(test, into = c('response', 'covariate'), sep = '_') %>%
  pivot_wider(names_from = response, values_from = c(estimate, p.value, p.bh), 
    names_glue = '{response}_{.value}', names_sort = TRUE) %>%
  select(.otu, covariate, starts_with('qmp'), starts_with('presence'))
#> select(.otu, sort(tidyselect::peek_vars()))
```

```{r}
p1 <- qmp_presence_wide %>%
  ggplot(aes(presence_estimate, qmp_estimate)) +
  facet_wrap(~covariate) +
  coord_fixed() +
  geom_point()
p1
```

```{r}
p2 <- qmp_presence_wide %>%
  ggplot(aes(presence_p.value, qmp_p.value)) +
  facet_wrap(~covariate) +
  coord_fixed() +
  scale_x_log10() +
  scale_y_log10() +
  geom_point()
p2
```

Overall, can see that most of the QMP correlations are highly similar to the presence correlations.

```{r, echo = FALSE}
print_table <- function(x, covariate) {
  x %>%
    filter(covariate == !!covariate) %>%
    select(-covariate) %>%
    rename(genus = .otu) %>%
    slice_min(qmp_p.value, n = 10) %>%
    mutate(across(where(is.numeric), signif, 2)) %>%
    knitr::kable()
}
```

**Faecal calprotectin**

```{r, echo = FALSE}
print_table(qmp_presence_wide, 'calprotectin')
```

**Moisture content**

```{r, echo = FALSE}
print_table(qmp_presence_wide, 'moisture')
```

Consider the calprotectin results.
For some taxa (like Veillonella, EscherichiaShigella) the two estimates are very similar;for others like Streptococcus they are quite different.

# Bias sensitivity analysis

Hypothesis: Correlations that are driven by differential prevalence will be less sensitive to bias than correlations that are not.

For the purposes of this analysis, zero observations will be treated as true zeros, which I expect to reduce the apparent impact of bias perturbation or calibration on the DA results relative to if we assumed zeros as indicating small positive abundances.

## Analysis

In order to perform a bias sensitivity analysis on the QMP profiles, we will calibrate (inverse-perturb) the relative abundances by the sampled bias vector, while keeping the total sum for the same constant, so that the total abundance of the same remains given by the fecal cell count.

```{r}
# Returns a matrix with taxa as columns
sample_bias <- function(n, gsd, vcv, taxa_as = 'cols') {
  stopifnot(isSymmetric(vcv))
  stopifnot(n > 1)
  n_taxa <- nrow(vcv)
  gsd_log <- log(gsd)
  bias <- MASS::mvrnorm(n, mu = rep(0, n_taxa), Sigma = vcv) %>%
    t %>%
    scale %>%
    {. * gsd_log} %>%
    exp
  if (taxa_as == 'cols')
    bias <- t(bias)
  bias
}

my_stat <- function(ps, otus) {
  ps %>%
    prune_taxa(otus, .) %>%
    as_tibble %>%
    with_groups(.otu, nest) %>%
    mutate(
      fit = map(data, ~cor.test(data = .x,
          ~ .abundance + faecal_calprotectin, 
          method = 'spearman', exact = FALSE, continuity = TRUE)),
      fit = map(fit, broom::tidy)
    ) %>%
    select(-data) %>%
    unnest(fit)
}
#> my_stat(ps, 'Veillonella')
```

```{r}
set.seed(42)
R <- 300
gsd <- 3
vcv <- diag(ntaxa(ps))
reps <- sample_bias(n = R, gsd = gsd, vcv = vcv) %>%
  split(., row(.)) %>%
  enframe(".idx", "bias")
```

```{r}
taxa_to_test <- c(
  #  low prevalence
  'Veillonella', 'EscherichiaShigella', 'Fusobacterium',
  # high prevalence
  'Streptococcus', 'Anaerostipes', 'Fusicatenibacter'
  )
reps_fit <- xfun::cache_rds({
  reps %>%
    mutate(
      fits = map(bias, 
        ~my_stat(
          ps = calibrate(ps, .x, norm = 'keep'),
          otus = taxa_to_test
        ))
    ) %>%
    unnest(fits)
}, dir = '_cache/', file = 'reps_fit', 
hash = list(reps, taxa_to_test))
```

```{r}
plot_bias_sens <- function(reps, original, .otu) {
  est_dist <- ggplot(reps, aes(estimate)) +
    geom_histogram() +
    geom_vline(xintercept = original$estimate, color = "darkred") +
    geom_vline(xintercept = 0, color = "grey") +
    labs(title = "Distribution of point estimates", x = "Estimate", y = "Count")
  pval_dist <- ggplot(reps, aes(p.value)) +
    geom_histogram() +
    geom_vline(xintercept = original$p.value, color = "darkred") +
    geom_vline(xintercept = 0.05, color = "grey") +
    scale_x_log10() +
    labs(title = "Distribution of p-values", x = "p-value", y = "Count")
  est_dist / pval_dist +
    plot_annotation(title = .otu)
}

original <- qmp_presence %>%
  filter(.otu %in% taxa_to_test, test == 'qmp_calprotectin') %>%
  with_groups(.otu, nest) %>%
  rename(original = data)

reps_plots <- reps_fit %>%
  with_groups(.otu, nest) %>%
  rename(reps = data) %>%
  left_join(original, by = ".otu") %>%
  mutate(.,
    plot = pmap(., plot_bias_sens)
  ) %>%
  select(.otu, plot) %>%
  deframe
```

## Top hits with low prevalence

```{r, echo = F}
fig.dim_sens <- c(6,6)
```

```{r, echo = FALSE, fig.dim = fig.dim_sens}
reps_plots$Veillonella
```

```{r, echo = FALSE, fig.dim = fig.dim_sens}
reps_plots$EscherichiaShigella
```

```{r, echo = FALSE, fig.dim = fig.dim_sens}
reps_plots$Fusobacterium
```

## Top hits with high prevalence

```{r, echo = FALSE, fig.dim = fig.dim_sens}
reps_plots$Streptococcus
```

```{r, echo = FALSE, fig.dim = fig.dim_sens}
reps_plots$Anaerostipes
```

```{r, echo = FALSE, fig.dim = fig.dim_sens}
reps_plots$Fusicatenibacter
```

## Summary

It isn't clear from these plots that there is a difference in bias sensitivity between the two sets.