### 2022-03-05 differential-abundance.Rmd
Revise differential-abundance section
- Add paragraph on rank-based analyses
- Revise for clarity and succinctness

# DA methods vary in their robustness to bias {#differential-abundance}

<!-- # Differential-abundance methods vary in their robustness to taxonomic bias  -->

We now consider how the error in MGS-derived abundance measurements affect DA analyses.
We focus on multiplicative DA analyses, or those that measure the fold change (FC) in the relative or absolute abundance of a taxon.
Such analyses are widely used in microbiome studies and have more direct ecological interpretations than other types via the multiplicative processes of exponential growth and decay.
In addition, we briefly consider non-parametric rank-based analyses, which are common in microbiome-wide association studies.

## Fold change between a pair of samples

The building blocks of a multiplicative DA analysis are the FCs in relative or absolute abundance between pairs of individual samples.
If the fold error (FE) in a given abundance measure is constant, then it will not affect the FCs among samples.
<!-- Ratios -->
We saw, for example, that consistent bias creates constant FE in the ratio between two species $i$ and $j$ equal to the ratio in their efficiencies (Equation \@ref(eq:ratio-error)).
This error completely cancels in the FC from sample $a$ to $b$,
\begin{align}
  (\#eq:ratio-fc-error)
\underbrace{\frac{\widehat{\text{ratio}}_{i/j}(b)}{\widehat{\text{ratio}}_{i/j}(a)}} _\text{measured FC}
  &= \frac
    {\text{ratio}_{i/j}(b) \cdot \cancel{\text{efficiency}_{i} / {\text{efficiency}_j}}}
    {\text{ratio}_{i/j}(a) \cdot \cancel{\text{efficiency}_{i} / {\text{efficiency}_j}}}
\\[0.5ex]
  &=
  \underbrace{\frac{\text{ratio}_{i/j}(b)}{\text{ratio}_{i/j}(a)}}_\text{actual FC}
  .
\end{align}
<!-- Proportions -->
In contrast, the variable FE in the proportion of a species (Equation \@ref(eq:prop-error)) does not completely cancel,
\begin{align}
  (\#eq:prop-fc-error)
\underbrace{\frac{\widehat{\text{prop}}_{i}(b)}{\widehat{\text{prop}}_{i}(a)}} _\text{measured FC}
  &= \frac
    {\text{prop}_{i}(b) \cdot \cancel{\text{efficiency}_{i}} / {\text{efficiency}_S(b)}}
    {\text{prop}_{i}(a) \cdot \cancel{\text{efficiency}_{i}} / {\text{efficiency}_S(a)}}
\\[0.5ex]
  &=
  \underbrace{\frac{\text{prop}_{i}(b)}{\text{prop}_{i}(a)}}_\text{actual FC}
  \cdot
  \underbrace{\left[\frac{\text{efficiency}_S(b)}{\text{efficiency}_S(a)}\right]^{-1}}_\text{fold error}
  .
\end{align}
Although the constant species' efficiency cancels, the varying mean efficiency does not, leaving a FE in the FC that is the same for every species and is equal to the inverse FC in the mean efficiency.

Figure \@ref(fig:error-proportions) (bottom row) illustrates these different effects of bias on FCs in proportions versus ratios.
In this hypothetical example, the mean efficiency decreases by a factor of 2.6 from Sample 1 to Sample 2 (FC of 0.4X), causing the FC in the proportion of each species to appear 2.6X larger than its true value.
Though the fold error for all species is the same, the implications depend on the actual FC and correspond to three distinct types of error: an increase in magnitude, a decrease in magnitude, and a change in direction; we refer to the latter as a _sign error_ in reference to the corresponding log fold change (LFC).
We can see each type of error in Figure \@ref(fig:error-proportions).
For Species 1, which increases and thus moves in the opposite direction of the mean efficiency, we see an increase in magnitude of the measured FC (actual FC: 2.3X, measured FC: 6.5X).
For Species 2, which decreases and thus moves in the same direction as the mean efficiency but by a larger factor, we see a decrease in magnitude (actual FC: 0.15X, measured FC: 0.44X).
For Species 2, which decreases by a smaller factor than the mean efficiency, we see a change in direction (actual FC: 0.6X, measured FC: 1.8X), such that the species actually appears to increase (a sign error).
In contrast, the ratios between species change by the same factors in the actual and observed taxonomic profiles.
For example, the ratio of Species 2 to Species 3 changes from $1/1$ to $1/4$ in the actual profiles and from $3$ to $3/4$ in the observed profiles, a decrease by 4X either way.

<!-- The ratios in sample 1 are Actual: 1:1:1, Observed: 1:18:6 -->
<!-- The ratios in sample 2 are Actual: 15:1:4, Observed: 5:6:8 -->

We can similarly understand the impact of bias on the measured FCs in a species' absolute abundance.
To the extent that the FE in the abundance measurement is constant across samples, it will cancel and lead to accurate FCs.
For the various measurement methods described in Section \@ref(absolute-abundance), the FE in a species' abundance can be expressed as a species-specific constant multiplied by a species-independent factor that, depending on experimental details, may vary with sample composition.
Thus the FE in the FC between a pair of samples is the same for all species.
Further, it can be reduced by experimental designs that minimize cross-sample variation in the species-independent factor, regardless of the magnitude of bias (variation in efficiency among species) in the MGS measurement.

## Regression analysis of many samples

DA analysis can be framed as a regression analysis, the simplest of which uses the simple linear regression model
\begin{align}
  (\#eq:regression)
  y(a) = \alpha + \beta x(a) + \varepsilon(a),
\end{align}
describing how a microbial _response variable_ $y$ relates to a _covariate_ $x$ and _residual (or unexplained) error_ $\varepsilon$.
For example, $y$ may be the log absolute abundance of species $i$ and $x$ may be a continuous variable, such as pH, or a binary variable, such as $x=1$ for treated patients and $x=0$ for untreated controls.
The regression coefficients $\alpha$ and $\beta$ are parameters to be estimated from the data.
A DA analysis is primarily interested in the slope coefficient $\beta$, which determines how the average or expected value of $y$ increases with $x$.
For a binary covariate, $\beta$ equals the average difference in $y$ between samples with $x=1$ versus those with $x=0$.
Taking the response $y$ to be the logarithm of a species' relative or absolute abundance makes the DA analysis multiplicative, with the coefficient $\beta$ corresponding to the average log fold change (LFC) in abundance between conditions or for a unit increase in $x$.

Appendix \@ref(appendix-regression) derives mathematical results for the effect that measurement error in the response caused by taxonomic bias has on LFCs estimated via simple linear regression.
Here we illustrate these results with a example.
Suppose that we wish to estimate the average LFC in the absolute abundance of every species between samples from two conditions, such as treated and untreated patients.
Further suppose that we have measured absolute abundances by multiplying the proportions from MGS with an accurate measurement of total-community abundance.
For each species $i$, we set $y(a)$ in Equation \@ref(eq:regression) equal to $\log \widehat{\text{abun}}_{i}(a)$, and use either Ordinary Least Squares (OLS) or Maximum Likelihood Estimation (MLE) to estimate the coefficients $\alpha$ and $\beta$.
From Equation \@ref(eq:density-prop-error), we know that the FE in the abundance of each species equals its efficiency relative to the mean efficiency of the sample.

How does this error affect the estimate of $\beta$ for species $i$?
There are three possibilities: 
In the first scenario, the mean efficiency is constant across samples, such that the FE in $\widehat{\text{abun}}_{i}(a)$ is constant.
This constant FE induces a constant additive error in $y$ that affects the estimate of the intercept $\alpha$, but not the slope $\beta$.
In the second scenario, the mean efficiency varies but is not systematically higher or lower in samples from treated versus untreated patients.
This unsystematic variation in the mean efficiency induces additive error in $y$ that is uncorrelated with $x$ and is statistically equivalent to an increase in the residual variation ($\epsilon$ in the regression model).
The estimate of $\beta$ will be less precise (i.e., have a larger standard error) than in Scenario 1, but will not be systematically distorted.
In the third scenario, the mean efficiency is systematically different in treated versus untreated patients.
If the mean efficiency is higher in treated patients, then the additive error in $y$ will tend to be smaller (or more negative) in this condition, leading to a downward shift in the estimate of $\beta$.
More generally, the portion of additive variation in the log mean efficiency that is correlated with $x$ creates an equal but opposite error in the estimate of $\beta$ that is the same for each species, whereas the portion that is uncorrelated with $x$ affects the standard errors of $\beta$.

Figure \@ref(fig:regression-example) shows a simulated example for which the mean efficiency is systematically higher in the condition $x=1$, leading to a systematic decrease in the estimated LFC for each species.
As we saw for the FC between individual samples, the biological significance varies by species, despite the absolute error in the LFC being shared.
We see decreases in magnitude (Species 9, 10, and 1 in Figure \@ref(fig:regression-example)), changes in sign (Species 5), or increases in magnitude (remaining species) depending the value of the actual LFC relative to the error.
The portion of the variation in the log mean efficiency that is uncorrelated with $x$ increases the standard errors for most species (Figure \@ref(fig:regression-example) D).
The exception is Species 9, which is the primary driver of the the mean efficiency dynamics in this example.
Decreased magnitudes and increased standard errors can both cause associations to be missed that would otherwise have been detected (Species 10 and 1), while increased magnitudes can turn weak or statistically insignificant associations into strong and statistically significant ones (Species 7, 6 and 4).

<!-- Figure: Regression example -->

```{r regression-example, fig.cap = '(ref:regression-example)'}
here::here(
  "notebook/_posts/2021-08-03-simulate-regression-example/",
  "simulate-regression-example_files/figure-html5",
  "main-figure-1.svg"
  ) %>%
  include_svg
```

(ref:regression-example) **Taxonomic bias distorts multi-sample differential abundance inference when the mean efficiency of samples is associated with the covariate of interest.** This figure shows the results of a regression analysis of simulated microbiomes consisting of 50 samples and 10 species from two environmental conditions indexed by $x=0$ and $x=1$. In this simulation, the species with the largest efficiency (Species 9) also has the largest positive LFC, which drives the positive association of the log mean efficiency with the condition (shown in Panels A and B). This positive LFC in the log mean efficiency induces a systematic negative shift in the estimated LFCs of all species (Panels C and D). Panel D shows the mean LFC (points) with 95% confidence intervals (CIs), for each species estimated from either the actual or the measured densities. The error (difference in LFC estimates on measured and actual) equals the negative LFC of the mean efficiency (shown in Panel B).

<!-- end Figure -->

## Rank-based analyses

Researchers often prefer to use nonparametric, rank-based DA methods to determine how species vary across sample conditions.
Rather than estimating LFCs, these methods estimate the average change in the cross-sample rank of a species' relative or absolute abundance.
Popular rank-based methods in microbiome DA analysis include the Wilcoxon-Mann-Whitney and Kruskal-Wallis tests when the covariate $x$ is discrete covariate and Spearman's rank correlation when $x$ is continuous.
Rank-based DA methods are unaffected by a constant FE in the response prior to rank transformation, since such an error does not affect the cross-sample ranks.
Hence consistent taxonomic bias does not affect rank-based analyses of species ratios.
However, variable FE in the untransformed response (as seen in species proportions) can change the cross-sample ranks.
If associated with the covariate $x$, the resulting error in the ranks will create a systematic error in the DA result.