### 2022-05-27 abundance-measurement.Rmd

# How bias affects abundance measurements {#abundance-measurement}

This section extends the theoretical results of @mclaren2019cons to describe the effect that consistent taxonomic bias within an MGS experiment affects the relative and absolute abundances measured for various microbial species.
We show that some approaches to quantifying species abundance yield constant fold errors (FEs), while others yield FEs that depend on overall community composition and thus can vary across samples.

## Model of MGS measurement

Our primary tool for understanding the impact of taxonomic bias on MGS measurement is the theoretical model of MGS measurement developed and empirically validated by @mclaren2019cons.
This model describes the mathematical relationship between the read counts obtained by MGS and the (actual) abundances of the various species in a sample.
Here we extend the model as first described @mclaren2019cons, which considers only relative abundances, to also consider absolute abundances.
For concreteness, we will consider _absolute abundance_, or simply _abundance_, to refer to the number of cells per unit volume in a sample (cell concentration).
That said, our results equally apply to other definitions of absolute abundance, such as the total number of cells in a sample or ecosystem and other abundance units such as biomass or genome copy number.

This model is the simplest that respects the multiplicative nature of taxonomic bias and the *compositional* nature of MGS measurements.
The actual abundance of a species in a given sample, multiplied by its measurement efficiency---its rate of conversion from cells to taxonomically assigned sequencing reads---determines the species' read count in that sample.
Taxonomic bias presents as variation in the measurement efficiencies among species within an MGS experiment.
The read counts further depend on sample-specific experimental factors that are typically unknown, such that they are best interpreted as only providing relative abundances (such data is said to be _compositional_) (@gloor2017micr).

We consider a set of microbiome samples measured by a specific MGS protocol that extracts, sequences, and taxonomically assigns reads to a set of microbial species $S$.
We make several simplifying assumptions to facilitate our analysis and presentation.
First, we consider only species-level assignment, and suppose that reads that cannot be uniquely assigned to a single species in $S$ are discarded.
Second, we ignore the possibility that reads are misassigned to the wrong species or the wrong sample.
Third, we suppose that taxonomic bias acts consistently across samples at the level of species---that is, a given species is always measured more efficiently than another to the same degree, regardless of variation in changes in sub-species-level composition or the sample matrix.
Finally, unless otherwise stated, we treat sequencing measurements as deterministic, ignoring the 'random' variation in read counts that arise from the sampling of sequencing reads and other aspects of the MGS process.
These assumptions, though unrealistic descriptions of most MGS experiments, serve the purpose of clearly demonstrating when and why consistent taxonomic bias creates errors in DA analysis.

Our model stipulates that the taxonomically-assigned read count for a species $i$ in a sample $a$ equals its abundance multiplied by species-specific and sample-specific factors,
\begin{align}
  (\#eq:mgs-model)
  \text{reads}_{i}(a)
  = \text{abun}_{i}(a) \quad \cdot
    \underbrace{\text{efficiency}_{i}}_{\substack{\text{species specific,} \\  \text{sample independent}}}
    \cdot \quad
    \underbrace{\text{effort}(a)}_{\substack{\text{species independent,} \\  \text{sample specific}}}.
%  \\ \\
%  M_i^{(a)} = A_i^{(a)} B_i F^{(a)}.
\end{align}
The species-specific factor, $\text{efficiency}_{i}$, equals the *relative measurement efficiency* (or simply *efficiency*) of the species relative to an arbitrary baseline species (@mclaren2019cons).
The variation in efficiency among species corresponds to the taxonomic bias of the MGS protocol.
The sample-specific factor, $\text{effort}(a)$, quantifies the *effective experimental effort* per unit abundance in measuring the sample.
Equation \@ref(eq:mgs-model) implies that the total number of assigned reads in sample $a$ equals
\begin{align}
  (\#eq:total-reads)
  \text{reads}_S(a)
    = \text{abun}_S(a) \cdot \text{efficiency}_S(a) \cdot \text{effort}(a),
    \\ \\
  \sum_i M_i^{(a)}
    = \left(\sum_i A_i^{(a)}\right) \cdot \bar B^{(a)} \cdot F^{(a)},
\end{align}
where $\text{abun}_{S}(a) \equiv \sum_{j\in S}\text{abun}_j(a)$ is the total abundance of species in $S$ and 
\begin{align}
  (\#eq:mean-efficiency)
  \text{efficiency}_S(a) 
    \equiv \frac{\sum_{j\in S}\text{abun}_j(a) \cdot \text{efficiency}_j}{\text{abun}_S(a)}
%    \\
%  \bar B^{(a)}
%    \equiv \frac{\sum_{i} A_i^{(a)} B_i}{\sum_i A_i^{(a)}}
\end{align}
is the _sample mean efficiency_, the mean efficiency of all species in sample $a$ weighted by their abundance.

## Relative abundance {#relative-abundance}

We distinguish between two types of species-level *relative abundances* within a sample.
The *proportion* of species $i$ in sample $a$ equals its abundance relative to the total abundance of all species in $S$,
\begin{align}
  (\#eq:prop)
  \text{prop}_{i}(a) 
    &\equiv \frac{\text{abun}_i(a)}{\text{abun}_S(a)}.
%  \\&= \frac{\text{abun}_i(a)}{\sum_{i \in S}\text{abun}_i(a)}.
%  P_{i}^{(a)}
%    &\equiv \frac{A_i^{(a)}}{\sum A_i^{(a)}}.
\end{align}
The *ratio* between two species $i$ and $j$ equals the abundance of $i$ relative to that of $j$,
\begin{align}
  (\#eq:ratio)
  \text{ratio}_{i/j}(a) \equiv \frac{\text{abun}_i(a)}{\text{abun}_j(a)}.
%  R_{i/j}^{(a)} \equiv \frac{A_i^{(a)}}{A_j^{(a)}}.
\end{align}

The measured proportion of a species is given by its proportion of all the assigned reads in a sample,
\begin{align}
  (\#eq:prop-meas)
  \widehat{\text{prop}}_{i}(a) = \frac{\text{reads}_i(a)}{\text{reads}_S(a)}.
\end{align}
We can rewrite the right-hand side (using Equations \@ref(eq:mgs-model), \@ref(eq:total-reads), and \@ref(eq:prop-meas)) to find
\begin{align}
  (\#eq:prop-error)
  \widehat{\text{prop}}_{i}(a)
  &= \text{prop}_{i}(a) \cdot \underbrace{\frac{\text{efficiency}_{i}}{\text{efficiency}_S(a)}}_{\substack{\text{variable} \\ \text{fold error}}}.
\end{align}
Taxonomic bias creates a multiplicative error in the species' proportion equal to its efficiency divided by the mean efficiency in the sample.
Since the mean efficiency varies across samples, so does the FE in a species' proportion.
This phenomenon can be seen for Species 3 in the two hypothetical communities in Figure \@ref(fig:error-proportions). 
Species 3, which has an efficiency of 6, is under-measured in Sample 1 (FE < 1) but over-measured (FE > 1) in Sample 2.
This difference occurs because the even distribution of species Sample 1 yields a mean efficiency of 8.33; in contrast, the lopsided distribution in Sample 2, which is dominated by the low-efficiency Species 1, has a mean efficiency of just 3.15.
A demonstration in bacterial mock communities is shown in [Figure 3C](https://doi.org/10.7554/eLife.46923.004) of @mclaren2019cons.

The measured ratio between species $i$ and $j$ is given by the ratio of their read counts, 
\begin{align}
  (\#eq:ratio-meas)
  \widehat{\text{ratio}}_{i/j}(a) = \frac{\text{reads}_i(a)}{\text{reads}_j(a)}.
\end{align}
From Equations \@ref(eq:mgs-model) and \@ref(eq:ratio-meas), it follows that 
\begin{align}
  (\#eq:ratio-error)
  \widehat{\text{ratio}}_{i/j}(a)
%  &= \frac{\text{abun}_{i}(a)}{\text{abun}_{j}(a)} 
  &= \text{ratio}_{i/j}(a) \cdot \underbrace{\frac{\text{efficiency}_{i}}{\text{efficiency}_{j}}}_{\substack{\text{constant} \\ \text{fold error}}}.
\end{align}
Taxonomic bias creates FE in the ratio measurement that is equal to the ratio in the species' efficiencies; the FE is therefore constant across samples.
For instance, in Figure \@ref(fig:error-proportions), the ratio of Species 3 (with an efficiency of 6) to Species 1 (with an efficiency of 1) is over-estimated by a factor of 6 in both communities despite their varying compositions.
A demonstration in bacterial mock communities is shown in [Figure 3D](https://doi.org/10.7554/eLife.46923.004) of @mclaren2019cons.

<!-- begin figure -->

```{r error-proportions, out.width = '90%', fig.cap = '(ref:cap-error-proportions)', cache = TRUE}
"figures/illustrations/error-proportions.svg" %>%
  include_svg
```

(ref:cap-error-proportions) **Taxonomic bias creates fold errors in species proportions that vary across samples and lead to inaccurate fold differences between samples.** Top row: Error in proportions measured by MGS in two hypothetical microbiome samples that contain different relative abundances of three species. Bottom row: Error in the measured fold difference in the third species that is derived from these measurements. Species' proportions may be measured as too high or too low depending on sample composition. For instance, Species 3 has an efficiency of 6 and is under-measured in Sample 1 (which has a mean efficiency of 8.33) but over-measured in Sample 2 (which has a mean efficiency of 3.15).

<!-- end figure -->

**Higher-order taxa:**
We can consider a higher-order taxon $I$, such as a genus or phylum, as a set of species, $\{i \in I\}$.
The abundance of taxon $I$ in sample $a$ is the sum of the abundances of its constituent species, $\text{abun}_{I}(a) = \sum_{i \in I}\text{abun}_{i}(a)$.
Similarly, the read count of taxon $I$ is the sum $\text{reads}_{I}(a) = \sum_{i \in I}\text{reads}_{i}(a)$.
The efficiency of taxon $I$ equals the abundance-weighted average of the efficiencies of its constituent species,
\begin{align}
  (\#eq:efficiency-general)
  \text{efficiency}_I(a) 
    \equiv \frac{\sum_{i\in I}\text{abun}_i(a)\cdot \text{efficiency}_i}{\text{abun}_I(a)}.
\end{align}
If the constituent species have different efficiencies, then the efficiency of the higher-order taxon will vary across samples depending on the relative abundances of its constituents (@mclaren2019cons).
As an example, suppose that Species 1 and Species 2 in Figure \@ref(fig:error-proportions) were in the same phylum.
The efficiency of the phylum would then be $\tfrac{1}{2} \cdot 1 + \tfrac{1}{2} \cdot 18 = 9.5$ in Sample 1 and $\tfrac{15}{16} \cdot 1 + \tfrac{1}{16} \cdot 18 \approx 2.1$ in Sample 2.
Equations \@ref(eq:prop-error) and \@ref(eq:ratio-error) continue to describe the measurement error in proportions and ratios involving higher-order taxa, so long as the sample-dependent taxon efficiency defined by Equation \@ref(eq:efficiency-general) is used.
In this way, we can see that both proportions and ratios among higher-order taxa may have inconsistent fold errors.

## Absolute abundance {#absolute-abundance}

Several extensions of the standard MGS experiment make it possible to measure absolute species abundances.
These extensions fall into two general approaches.
The first approach leverages information about the abundance of the total community; for example, @vandeputte2017quan measured total-community abundance using flow cytometry and multiplied this number by MGS genus proportions to obtain the absolute abundances of individual genera (@vandeputte2017quan).
A second approach leverages information about the abundance of one or more individual species; for example, a researcher might 'spike in' a known, fixed amount of an extraneous species to all samples prior to MGS, and normalize the read counts of all species to the spike-in species (@harrison2021theq).
We consider each approach in detail to determine how taxonomic bias affects the resulting absolute-abundance measurements.

### Leveraging information about total-community abundance

Suppose that the abundance of the total community in each sample has been measured by a non-MGS method.
The absolute abundances of individual species can be obtained by multiplying the species proportions (obtained via MGS) by this total-abundance measurement,
\begin{align}
  (\#eq:density-prop-meas)
  \widehat{\text{abun}}_{i}(a) 
  &= \widehat{\text{prop}}_{i}(a) \cdot \widehat{\text{abun}}_S(a).
%  \\ \\
%  \hat A_i^{(a)} &= \hat P_i^{(a)} \cdot \widehat{\sum_i A_i^{(a)}}.
%  \\ \\
%  \hat A_i^{(a)} &= \hat P_i^{(a)} \cdot \widetilde{\sum_i A_i^{(a)}}.
\end{align}
Total-abundance measurements recently used for this purpose include counting cells with microscopy (@lloyd2020evid) or flow cytometry (@props2017abso, @vandeputte2017quan, @galazzo2020howt), measuring the concentration of a marker-gene with qPCR or ddPCR (@zhang2017soil, @barlow2020aqau, @galazzo2020howt, @tettamantiboshier2020comp), and measuring bulk DNA concentration with a florescence-based DNA quantification method (@contijoch2019gutm).

Importantly, these methods of measuring total-community abundance are themselves subject to taxonomic bias.
Flow cytometry may, for example, yield lower cell counts for species whose cells tend to clump together or are prone to lysis during steps involved in sample collection, storage, and preparation.
Marker-gene concentrations measured by qPCR are affected by variation among species in extraction efficiency, marker-gene copy number, and PCR binding and amplification efficiency.
We can easily understand the impact of taxonomic bias on total-abundance measurement under simplifying assumptions analogous to those in our MGS model.
Suppose that each species has an _absolute efficiency_ for the total-abundance measurement that is constant across samples.
Neglecting other sources of random and systematic error, the total-abundance measurement equals
\begin{align}
  (\#eq:total-density-error)
  \widehat{\text{abun}}_S(a) 
  &= \sum_{i\in S} \text{abun}_i(a) \cdot \text{efficiency}^{\text{tot}}_i
\\&= \text{abun}_S(a) \cdot \text{efficiency}^{\text{tot}}_S(a),
\end{align}
where $\text{efficiency}_{i}^{\text{tot}}(a)$ is the absolute measurement efficiency of species $i$ for the total-abundance measurement and $\text{efficiency}^{\text{tot}}_S(a)$ is the mean efficiency of the total-abundance measurement in the sample.
<!-- Note, we have assumed that the total-abundance and MGS measurements are sensitive to the same species -->

The resulting absolute-abundance measurements of individual species are affected by taxonomic bias in both the MGS and total-abundance measurement.
We can compute this effect by substituting Equations \@ref(eq:prop-error) and \@ref(eq:total-density-error) into Equation \@ref(eq:density-prop-meas), yielding
\begin{align}
  (\#eq:density-prop-error)
  \widehat{\text{abun}}_{i}(a) 
  = \text{abun}_{i}(a) \cdot \frac{\text{efficiency}_{i} \cdot \text{efficiency}^{\text{tot}}_S(a)}{\text{efficiency}_S(a)}.
%  \\ \\
%  \widetilde{\sum A_i^{(a)}}
%  = \left(\sum A_i^{(a)}\right) \cdot \frac{B_i \cdot \bar B^{\text{tot},(a)}}{\bar B^{(a)}}.
\end{align}
Equation \@ref(eq:density-prop-error) indicates that the FE in the measured absolute abundance of a species equals its MGS efficiency relative to the mean MGS efficiency in the sample, multiplied by the mean efficiency of the total measurement.
As in the case of proportions (Equation \@ref(eq:prop-error)), the FE depends on sample composition through the two mean efficiency terms and so will vary across samples.
Notably, if the mean efficiency of the total-abundance measurement mirrors that of the MGS measurement, then the two offset each other and lead to more stable FEs.
We discuss how this possibility might be exploited in real experimental workflows in Section \@ref(solutions).

### Leveraging information about a reference species

Suppose that the absolute abundance of a _reference species_ $r$ is known, either by experimental design or by direct measurement using a non-MGS method. 
The absolute abundances of each species can be obtained by multiplying its read count by the abundance-to-reads ratio for species $r$,
\begin{align}
  (\#eq:density-ratio-meas)
  \widehat{\text{abun}}_{i}(a) = \text{reads}_{i}(a) \cdot \frac{\widehat{\text{abun}}_{r}(a)}{\text{reads}_{r}(a)}.
\end{align}

A common application of this approach involves adding a 'spike-in' (as described above) in a known (and typically constant) abundance across samples (@stammler2016adju, @ji2019quan, @harrison2021theq, REFs).
In this case, the reference abundance $\widehat{\text{abun}}_{r}(a)$ is determined from the concentration of the spike-in stock multiplied by the ratio of the spike-in to sample volumes.
(In practice, researchers in spike-in experiments have typically used a more indirect calculation to determine species abundances, but which yields identical results to \@ref(eq:density-ratio-meas); Appendix REF).

Others have instead sought to determine naturally-occurring species that are thought to be constant across samples; we refer to such species as _housekeeping species_ by analogy with the housekeeping genes used for absolute-abundance conversion in gene-expression studies (REF).
Housekeeping species can sometimes be identified using prior scientific knowledge; for example, in shotgun sequencing experiments, researchers have used sequencing reads from the plant or animal host as a reference (REFs).
A related approach involves computationally identifying species that are constant between pairs of samples (@david2014host) or between sample conditions (@mandal2015anal, @kumar2018anal).
The abundance of a housekeeping species is typically unknown; therefore, to estimate the abundances of other species, we simply set $\widehat{\text{abun}}_{r}(a)$ to 1 in Equation \@ref(eq:density-ratio-meas).
The resulting abundance measurements have unknown but fixed units, which is sufficient for measuring fold changes across samples.

We suggest an additional way of using the reference-species strategy even in the absence of a spike-in or constant species:
Performing targeted measurements of the absolute abundance of one or more naturally occurring species.
These species can then be used as reference species in Equation \@ref(eq:density-ratio-meas) to measure the absolute abundances of all species.
The most common form of targeted measurement involves using qPCR or ddPCR to measure the concentration of a marker-gene in the extracted DNA.
It is also possible to directly measure cell concentration by performing ddPCR prior to DNA extraction (REFs), flow cytometry with species-specific florescent probes, or CFU counting on selective media.
Appendix [REF] describes how using multiple reference species and/or statistical modeling can address the fact that any one native reference species is unlikely to be found in all samples.

The effect of taxonomic bias on absolute abundance measurements derived from a reference species can be determined by substituting Equation \@ref(eq:ratio-error) into Equation \@ref(eq:density-ratio-meas), which indicates that the measured abundance of species $i$ is 
\begin{align}
  (\#eq:density-ratio-error)
  \widehat{\text{abun}}_{i}(a) 
  = \text{abun}_{i}(a) \cdot \frac{\text{efficiency}_{i}}{\text{efficiency}_{r}}
  \cdot \underbrace{\frac{\widehat{\text{abun}_r}(a)}{\text{abun}_r(a)}}_{\substack{\text{fold error in} \\ \text{ref. species}}}.
  % \cdot \begin{array}{c} \text{fold error in} \\ \widehat{\text{abun}_r}(a) 
\end{align}
The FE in $\widehat{\text{abun}}_{i}(a)$ consists of two terms: the relative efficiency of species $i$ to species $r$ in the MGS measurement, and the FE in the reference species' abundance.
If bias is consistent across samples and the reference species can be spiked in or measured with a constant FE, then the FE in the $\widehat{\text{abun}}_{i}(a)$ will be constant.