### 2022-06-05 appendix-diversity.Rmd
Clean up appendix
- Remove old appendix and notes
- Put diversity derivation in its own section
- Create stubs for some missing sections

# Alpha diversity and variation in the mean efficiency {#diversity-and-mean-efficiency}

The diversity of order $q$ is given by
\begin{align}
  ^qD = \left(\sum_{i=1}^n p_i^q\right)^{1 /(1-q)},
\end{align}
where $^qD$ is understood to be given by the limit of the RHS when $q = 1$ (and so the given expression is undefined).
The order-2 diversity is therefore
\begin{align}
  ^2D = \frac{1}{\sum_{i=1}^n {p_i^2}},
\end{align}
which is equivalent to the Inverse Simpson Index (REF Jost).

Consider an infinite pool of species.
Let $\sigma^{2}$ denote the variance in the relative efficiency among species in the pool.
Consider a community that is assembled by randomly choosing $K \ge 1$ species from the pool and setting the abundances of the $K$ species in a manner that is independent of their efficiencies.

**Claim:** 
Let $\rho_{k}$ for $1 \le k \le K$ denote the proportions of the $K$ species in the community; this new notation serves as a reminder that the subscript $k$ indexes a random species specific to this particular community.
Conditional on the order-2 diversity $^2D$ of a community, the arithmetic variance in the mean efficiency is
\begin{align}
  Var[\bar B \mid ^2D] 
% = \sigma^2 \sum_{k=1}^K \rho_k^2
  = \frac{\sigma^2}{^2D}.
\end{align}

**Proof:** Let $\beta_{k}$ denote the efficiency of the $k$-th sampled species.
The mean efficiency is therefore $\bar B = \sum_k \rho_k \beta_k$.
Conditional on the proportions $\{\rho_{k}\}$, the variance in the mean efficiency is
\begin{align}
  Var[\bar B \mid \{\rho_{k}\}]
  &= \sum_{k=1}^K \rho_k^2 Var[\beta_k]
  \\&= \left(\sum_{k=1}^K p_k^2\right) \sigma^2.
\end{align}
The first line follows from the fact that the $\beta_{k}$ are independent.
The summation in the final line is equal to $1/2^D$, thus proving the result.

**Note:** 
We have showed that the arithmetic (additive) variance of $\bar B$ decreases with $^2D$; however, the geometric (multiplicative) variance is most relevant for understanding the effect of bias on DA.
As $2^D$ increases, the distribution of $\bar B$ will converge (by the central limit theorem) to a normal distribution, and both the arithmetic and geometric variance will decrease.